### 2019-02-15 manuscript/MS_trophToxNTO.tex
init Manuscript

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%\RequirePackage{lmodern}	% font set: Latin Modern
\RequirePackage{charter}	% font set: Charter

\usepackage[french,english]{babel}

% Package ADDED
\usepackage{xcolor,graphics,graphicx}

% For citations
\usepackage{csquotes}

% For line spacing
\usepackage{setspace}
\linespread{1.3}

% To rotate a table or a figure
\usepackage{lscape}
% MATHS
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{empheq}

% TAILLE de la page
\usepackage{geometry} % define page margin
\geometry{top=20mm,left=30mm,right=30mm,bottom=20mm} % margin from top, left, right and bottom

% COMPTEUR de ligne
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
\usepackage{caption}
\usepackage{lineno}
\linenumbers

% BIBLIOGRAPHIE
%\usepackage[numbers]{natbib}
\usepackage[authoryear]{natbib}
\usepackage{hyperref}

\begin{document}

\begin{center}
	\Large\textbf{
	Trophic transfer of anticoagulant rodenticides while managing rodent pests: the fine line between predator-prey regulation and pesticide-pest regulation
}\par
\end{center}

\vspace{.5cm}

\begin{center}
	\large\textbf{
		Virgile Baudrot$^{1,2}$ 
		Javier Fernandez-de-Simon$^1,3$
		Michael Coeurdassier$^1$,
		Geoffroy Couval$^1,4$,
		Patrick Giraudoux$^1$,
		Xavier Lambin$^5,6$		
	}\par	
\end{center}

\begin{center}
	$^1$ Université Bourgogne Franche-Comté - UMR CNRS 6249 Laboratoire Chrono-environnement, 25030 Besançon, France\\
	$^2$ BioSP, INRA, 84000 Avignon, France\\
	$^3$ IREC. Instituto de Investigación en Recursos Cinegéticos\\
	$^4$ FREDON Franche-Comté, Espace Valentin Est, 12, Rue de Franche-Comté - Bât E, 25480 Ecole-Valentin, France\\
	$^5$ School of Biological Sciences, University of Aberdeen, Zoology Building, Tillydrone Avenue, Aberdeen AB24 2TZ, Scotland, UK\\
	$^*$ Corresponding authors: virgile.baudrot@posteo.net or x.lambin@abdn.ac.uk
\end{center}

\paragraph{Running Title} Bufferscapes to preserve Non-Target Organism / Ecosystem services

\vspace{.5cm}

\begin{abstract} ~
	\begin{enumerate}
		\item Understanding pesticide impacts on populations of target/non-target species and communities is a challenge to applied ecology. When predators that otherwise regulate pest densities ingest prey contaminated with pesticides, this can suppress predator populations as contaminated prey act as a super-predator, with pesticides controlling both pest and predators. It is, however, unknown how species relationships and protocols of rodenticide treatments, i.e. farmer functional responses, interact to affect pest regulation.
		\item Here we used linked differential equations to model a heuristic non-spatialized system including montane water voles, specialist vole predators (stoats, weasels), and a generalist predator (red fox) which consume voles, mustelids and other prey. By considering anticoagulant rodenticide toxicokinetic and toxicodynamic equations, we explored the impact of 5 farmer functional responses (defined by both rodenticide quantity and threshold vole density above which rodenticide spreading is prohibited) on predator-prey interactions, rodenticide transfer across the trophic chain and population effects.
		\item Spreading low quantities of rodenticide while maintaining sufficient voles as prey resources led to less rodenticide being applied and extended periods without rodenticide in the environment, benefitting predators while avoiding episodes with high vole density. This may meet farm production interests while minimizing the impact on small mustelid and fox populations.
		Spreading rodenticide at low vole densities suppressed mustelid and fox populations, leading to vole population dynamics being entirely regulated by rodenticide use. This vole-eradication treatment regime inhibited predator ecosystem services and promoted pesticide dependence.
		\item Farmer functional response with high or intermediate densities thresholds triggering treatment produced vole population fluctuations caused by rodenticide that could be followed by periods during which mustelids regulated vole populations. These alternative phases of mustelids and farmer regulation highlight the benefit of intraguild relationship where mustelids may rescue foxes from poisoning.
	\end{enumerate}

\paragraph{Synthesis and applications} Different farmer functional responses lead to a rich variety of population dynamics in predator-prey systems. That such a pesticide-tri-trophic system may cause a variety of population dynamics responses to pesticide use in agro-ecosystems is a novel insight. Our model reveals the need for maintaining refuges with sufficient non-poisoned voles for specialist mustelids, to conserve predator community, given the super-predator role of rodenticides. We suggest that long periods without pesticide treatment are essential to maintain predator populations, and that practices of pesticides use that attempt to permanently eradicate a pest over a large scale are counterproductive.

\end{abstract}

\paragraph{Keywords} Biodiversity conservation; bromadiolone transfer; cyclic fluctuations; pesticides impact; secondary poisoning; super-predation; cascade effects

\section{Introduction}

Since the “green revolution” following the 1950s, pesticides use has increased to control pests damaging properties, public health or crops \citep{Tilman2002}. Pesticides usage is varyingly triggered by the perception/estimation of pest densities. Natural enemies (e.g., predators, parasites, competitors) also reduce pest densities and hence may preclude the need for using pesticides \citep{Michalko2017}. Natural enemies are, however, also affected by pesticides, either by direct exposure, through ingestion of contaminated prey \citep{Berny2007} or indirectly by cascade effects of resource depletion \citep{Halstead2014}.  Thus, under some regimes of pesticides use, pest populations only become regulated by pesticides once predators have collapsed. To preserve ecosystem health and the services of predators through regulation of pest densities, we need to assess the feasibility and benefit of pesticide treatment regimes in their ability to control pest species with minimal damage on predators \citep{Halstead2014}. It is however empirically challenging to assess the overall impact of pesticide treatments on the dynamics of species linked by trophic interactions. In this context, process-based models describing simplified scenarios are powerful tools to reveal hidden patterns by disentangling processes emerging from pesticide impacts on predator-prey system (e.g., \citet{Baudrot2018}).

Voles and other grassland rodent species undergo multi-annual population cycles (e.g.,  \cite{Krebs2013}. At their peaks, they may attain extremely high densities, causing substantial damage to grass/cultivated crops and forestry and conflicts with humans \citep{Delattre2009}. Farmers worldwide expand financial resources to purchase and spread anticoagulant rodenticides (hereafter AR), hoping to reduce vole populations and damages and increase profits despite the investment required \citep{Stenseth2003}. They do so according to protocols, equivalent to farmer functional responses (hereafter FFR), that involve varying amounts of AR spread in response to different thresholds in vole densities.

Voles and many small rodents are perceived as pests, but they are also keystone species, crucial to the functioning of grassland ecosystems, as well as being the prey of numerous predators, including species of conservation concern \citep{Delibes-Mateos2011, Coeurdassier2014}. Their population cycles create pulses of resources crucial to the viability of a wide range of resident predators and the aggregation of mobile avian vole predators \citep{Korpimaki1991}. The smallest mustelids (e.g. weasels Mustela nivalis) are specialist vole predators. Their numerical response has been shown theoretically to be necessary for generating predator-prey cycles \citep{Hanski1991}. They are said to be responsible for driving 3-to-5-year vole cycles in Fennoscandia (specialist predation hypothesis) \citet{Hanski1991}. Generalist resident predators like foxes (Vulpes vulpes) are expected to have regulatory and limiting effects on voles, owing to dietary plasticity that slows down vole population increase at low density \citet{Hanski1991}. Foxes do not show numerical responses to vole abundance \citep{Weber2002} but they influence the food chain through occasional killing and consumption of bite-sized mustelids. Mustelids form a small proportion of fox diet (0-10\%) but their offtake could represent a significant portion of the population (reviewed in \citet{Lambin2017}).

Anticoagulant rodenticides are non-selective toxicants with deleterious effects on non-target fauna (e.g. \cite{Coeurdassier2014}. Despite AR being exclusively licensed for rodent control, a large number of predator species are secondarily exposed to AR \citep{Sanchez-Barbudo2012}. Repeated consumption of dead and sub-lethally intoxicated voles reduced fox abundance in farmland in eastern France \citep{Jacquot2013} and ARs caused short-term declines in stoats in New Zealand \citep{Alterio1996}. Rodent-eating mustelid populations are affected by ARs given the pervasive levels of contamination reported \citep{McDonald1998}. Thus, there is little doubt ARs use inadvertently depresses predator populations. As they likely limit vole populations, it is essential to understand when AR use becomes counterproductive by altering the pest population dynamics, producing more frequent outbreaks and high residual vole abundance.

With the aim of understanding the potentially complex interactions between prey that are perceived as pest predators and farmers spreading rodenticide in response to vole abundance and their functional responses, we studied a simplified system inspired by cyclically fluctuating montane water voles (\textit{Arvicola scherman}), small mustelids (stoats, weasels) that mostly eat voles (specialists), and foxes (generalists), with voles and mustelids as food items. We used a process-based model using differential equations to explore 5 FFR types of AR spread, combining population dynamics, predator-prey interactions and rodenticide transfer across the trophic chains. Model parameters and FFR were inspired by farming systems in the Jura Mountains, Franche-Comté (France), the region of Comté cheese production. In Franche-Comté, farmers shifted from polyculture to almost exclusively grass production for milk used to produce cheese from the early 1970s \citep{Giraudoux1997}. Due to recurrent vole outbreaks and damages to grasslands, massive rodenticide treatments were implemented from the early 80s with consequences on non-target wildlife. Practices developed technically under pressure from public opinion, farmer unions and farmer technical organizations collaborating with researchers to find treatment regimes with less harmful consequences for biodiversity \citep{Delattre2009}.

Hence, our main objective was to explore the properties of FFR in relation to varying vole density, to vole outbreak frequencies, to the guild of interacting predators and the effects on the global tri-trophic population dynamics. Additionally, we sought to establish under what regime of use the ARs deliver benefits to farmers without imposing damages to the ecosystem.

\section{Methods}

We specified and parameterized the tri-trophic system of voles-mustelids-foxes with use of AR by farmers in response to vole density. We considered different FFR to assess AR upward transfer through the trophic web, considering the coupled dynamics of pulsed of AR spread and population dynamics. In all cases, parameterization units are on hectare-1 and day-1.

\subsection{Model for the tri-trophic dynamic}

We considered a tri-trophic system described by Fig. 1 and equations (1-5); parameterization is provided in Supporting Information 1. Voles, denoted V, were the primary prey, Mustelids, M, intermediate predators, and Foxes, F, were top predators, consuming voles and mustelids. For each species, the instantaneous variation of population size over time is:

\begin{equation}
\begin{array}{l l}
\dfrac{dV}{dt} & = V r_V \left( 1- \dfrac{V}{K_V}\right) - \Phi_{V,M}(V)M - \Phi_{V,F}(V,M)F\\[.3cm]
%
\dfrac{dM}{dt} & = \varepsilon_M \Phi_{V,M}(V)M - m_M M - \Phi_{M,F}(V,M)F\\[.3cm]
%
\dfrac{dF}{dt} & = F r_F \left( 1- \dfrac{F}{K_F}\right)
\end{array}
\end{equation}

The vole population followed a logistic growth rate, with rV  the maximal reproduction rate, fixed at $r_V = \ln(2 \times 600)/365$ per day, since montane water vole populations can increase from 0  to  600 individuals ha-1 or more \citep{Giraudoux1997} resulting in the equilibrium density being fixed at $K_V = 600$ individuals. 
%
The vole population was preyed upon by mustelid and fox populations. The vole consumption rate at different vole densities was described by functional responses ($\Phi_{V,M}$ for mustelids, $\Phi_{V,F}$  for foxes), see equations \eqref{eq:functionResponse}.
%
We treated small mustelids as vole specialist predators \citep{King2006}, assuming a Holling Type 2 functional response with attack rate $a_M$ and handling time $h_M$ (equation \eqref{eq:functionResponse}). We then represented foxes feeding on voles and mustelids by a multi-species functional response derived from Holling Type 3, referring to generalist feeding behaviour \citep{Baudrot2016}. For that, we denoted $a_{VF}$ and $a_{MF}$ the fox attack rate on voles and mustelids respectively. The parameter hF was the handling time for foxes.

\begin{equation}
\begin{array}{l l}
 \Phi_{V,M}(V)M = \dfrac{a_M V}{1 + h_M a_M V} \\[.3cm]
 %
 \Phi_{V,F} = \dfrac{a_{VF} V}{a_{VF}V + a_{MF}F} \times \dfrac{(a_{VF}V + a_{MF} M)^2}{1 + h_F(a_{VF}V + a_{MF} M)^2}\\[.3cm]
 %
  \Phi_{M,F}(V,M)F = \dfrac{a_{MF}V}{a_{VF}V + a_{MF}F} \times \dfrac{(a_{VF}V + a_{MF}M)^2}{1+h_F(a_{VF}V + a_{MF}M)^2}
\end{array}
eq:functionResponse
\end{equation}

Parameterization of functional responses was estimated to fit the daily satiation level of predators for handling times, and the observed 5-6 year vole cycles for attack rates (Supporting Information 1). We assumed foxes spent longer searching for voles than mustelids, based on each species diet and daily number of individuals captured. Therefore, $a_{VF}$ was considered larger than $a_{MF}$ and selected to produce 6-year vole cycles without AR.

Since we assumed small mustelids behave as specialist predators, we considered a numerical response linearly dependent on the functional response with parameter $\varepsilon_M$ (dimensionless) as conversion efficiency of prey into newborn predator (Supporting Information 2). The mustelid background mortality rate (i.e. all other reasons of death: ageing, disease, etc.) was $m_M$ (Supporting Information 2). 
We assumed foxes had a logistic growth rate function, parameterized with maximal growth rate $r_F = ln(3)/365$ and equilibrium density $K_F = 0.03$ [individuals.ha$^{-1}$] \citep{Ruette2003} without numerical response \citep{Weber2002}.

\subsection{Model with rodenticide}

Fig. 1 represents the whole study system. Rodenticide is spread in grasslands during treatments, denoted $T_{Broma}(V)$, triggered by vole density $V$ (FFR described in Table 1).
%
Firstly, baits (50 mg.kg$^{-1}$  of bromadiolone, hereafter AR) were spread in grasslands at quantity 7.5 to 20 kg.ha$^{-1}$.
%
Such quantity, $C$, was available for voles, and a proportion disappeared in the environment at rate $k_0$ (set at $k_0 = 0.0815$) \citep{Sage2008}. The proportion consumed per vole, with rate $\kappa(C)$, was assumed to be an increasing function. The function $\kappa(C)$ was characterized by a maximum ingestion $M_{in}$, and a half-saturation constant for ingestion $D_{in}$ in [mg.kg$^{-1}$]:

\begin{equation}
\kappa_C= \dfrac{M_{in}\times C}{D_{in} + C}\dfrac{M_{in} \times C}{D_{in} + C} 
\end{equation}

For the toxicokinetics of AR (i.e. internal compound dynamics) leading to AR concentrations in animal body (voles, mustelids and foxes), we considered an uptake without biotransformation and time-regulated distribution, (i.e. AR concentration in the body of animals was instantly homogeneous) and that the whole body was consumed or scavenged without selection/rejection of tissues-organs. We also assumed disappearance including excretion of the parent compound and metabolisation, and that metabolites were non-toxic and/or excreted in the scats. 
For the toxicokinetics of AR ingested by voles, a fraction $C_V$ was assumed to remain active, stored mainly in vole livers and available to predators ingesting voles. The absorption rate of ARs ($\eta$) exceeds 50\% in less than 24h \citep{Jacquot2013}. The excretion rate from voles, $k_{out}$, $V$ was 0.4 day$^{-1}$ \citep{Sage2008}. The mortality rate through ARs was $\mu(C_V)$. Death through poisoning created a dead vole population ($V_d$) with AR concentration $C_V$. Dead voles could either be scavenged by mustelids/foxes or decompose at rate $d$. We assumed AR in dead voles disappeared from the system when voles decomposed. 
Mustelids could feed on live voles $V$, or non-decomposed dead voles Vd and we assumed a Type 2 functional response adapted for a multi-species functional response \citep{Baudrot2016}. Mustelids ingested AR with absorption rate $\eta_M$ (ratio between biomasses of voles, $B_V$, and mustelids, $B_M$) and the total of ingested voles (alive $V$ and dead $V_d$) was defined by function $\Theta_M (V,V_d=$ (see Supporting Information 2), in equation (4) which provides the rate of AR ingested by a mustelid:

\begin{equation}
C_V \times \eta_M \times \Theta_M(V,V_d)
\end{equation}

A fraction of AR ingested was accumulated in weasels while the rest was excreted with rate $k_{out,M}$. AR ingestion induced weasel lethal poisoning at rate $\mu_M(C_M)$, additive to natural mortality rate $m_M$.
AR was ingested by foxes with a rate proportional to the functional response of foxes to voles, dead voles and mustelids. Foxes also accumulated AR available in their prey, resulting in upward AR transfer in the trophic chain. Foxes accumulated AR in concentration $C_F$. A fraction of AR was excreted by foxes at rate $k_{out,F}$ \citep{Sage2008}. AR caused fox mortality at rate $\mu_F$ ($C_F$) \citep{Sage2010}.
We used log-logistic equations for describing dose-dependent mortality of animals exposed to AR. Vole and predator mortality rates due to AR $\mu_X(C_X)$ ($X$ referring to the considered species) were expressed by equation (5):

\begin{equation}
\mu_X(C_X) = \dfrac{1}{\text{period of time}} \times \dfrac{1}{1+ (LD_{50}/C_X)^H}
\end{equation}

where LD50 was the daily median lethal dose (50\% of population dying) with different values for voles (2 mg kg$^{-1}$), mustelids (2.1 mg kg$^{-1}$) and foxes (0.5 mg kg$^{-1}$, Supporting Information 1). $H$ is the Hill's coefficient, modulating the curve steepness and was estimated to fit the sparse data we have (Supporting Information 1). Mortality rates considered the duration of AR toxicity from experiments of up to 6 days \citep{Sage2010}.

\subsection{The farmer functional responses explored through simulation}

We considered a range of realistic FFR spanning treatments during vole outbreaks only as a precautionary approach; in which treatments only takes place at intermediate or low vole density threshold. These scenarios are inspired by historic and contemporary protocols of bromadiolone use to control montane water voles in Franche-Comté, but also representative of practice globally (Table 1) \citep{Delattre2009}. To check the influence of foxes and intraguild predation on the system’s dynamics, we simulated scenarios with and without foxes.

Our simulations tracked the linked vole-mustelid-fox dynamics for 40 years, after a “burn-in” period of 10 years to reduce dependency of results upon initial conditions, to observe several vole cycles and to characterise AR effects on these species population dynamics. This burn-in period also had AR treatment triggered at specific vole densities and with a given rodenticide quantity for each FFR (see Table 1). The burn-in phase was selected according to a set of simulations with different initial conditions. Those simulations showed that in a given FFR (i.e., same threshold of vole density and amount of AR spread), the dynamics of the population were converging toward a similar pattern. 
In order to assess the impact of FFR on both agriculture and conservation interests, we estimated the following cost functions: (i) Number of treatment events per FFR ; (ii) Cumulative amount of AR (kg) ; (iii) Proportion of time when the AR-induced mortality of mustelids higher than 50\% (i.e. lethal exposure profile killing 50\% of mustelid population) ; (iv) Proportion of time when the mortality of mustelids was higher than 50\% due to natural mortality; (v) Proportion of time when the vole density was below 50 voles ha-1, as a proxy for time when forage grass grows with low herbivore influence; (vi) Mean vole, mustelid and fox densities.

\section{Results}

\subsection{Population dynamics}

Allowing for mortality by predators ingesting AR-poisoned voles changed the outcome of predator-prey dynamics involving vole, mustelid and fox populations. Secondary poisoning led to a rich spectrum of emergent dynamics according to the FFR to vole abundance.  
Without AR (FFR a), vole dynamics were regulated by mustelid predation that gave rise to 6-year cycles (Figs. 2 and 3; maximal vole and mustelid densities were around 600 ind.ha-1 and 2 ind.ha-1 respectively). The fox population remained at its carrying capacity (i.e., 0.03 ind. ha-1, see Supporting Information 1).

Under FFR b (high vole density threshold triggering treatment and high AR amount per treatment) and FFR d (intermediate threshold and low AR amount), vole dynamics were sequentially regulated by either AR treatments, which we refer to as farmer-regulated phase (FR in Figures 2 \& 3), or by mustelids, mustelids-regulated phase (MR), alternately (MR in Figures 2 \& 3). Farmer-regulated periods started when densities of living voles triggered treatments. This produced sudden declines of live voles followed by increases of dead voles. However, the vole population re-grew quickly which triggered frequent further treatments and pulses of availability of contaminated (both live and dead) voles. In FFR d (intermediate threshold and low AR), vole declines were not as deep as when pulses of AR amount were high (FFR b), owing to the reduced rodenticide amount per treatment. Mustelids and foxes also experienced AR-induced declines during this period. Under FFRs b (high threshold density, high AR) and d (intermediate threshold, low AR), mustelid-regulated periods started when mustelid numbers grew slowly to a peak, which depressed vole density, precluding rodenticide treatments and releasing the fox population from secondary poisoning, such that its abundance rebounded. Vole depletion by mustelids and subsequent mustelid declines allowed the vole population to grow again up to threshold densities and initiated a new period of regulation by farmers.
Vole dynamics were permanently regulated by AR treatment under FFR c (intermediate threshold, high AR, Figs. 2c, 3c). Populations of live and dead voles experienced high frequency fluctuations (around 2 peaks every year) driven by AR. As AR treatments were frequent, being triggered by voles peaks, contaminated dead voles were always abundant (peaks at 90 ind.ha-1), and mustelid and fox densities remained low (mustelids, converging to 0 ind. ha-1; foxes, 0.003-0.006 ind. ha$^{-1}$) with only short-term fluctuations (2 peaks every year, following vole cycles).
With FFR e (low threshold, low AR), vole populations were maintained by farmers at around 50 voles ha-1 (Figs. 2e and 3e). The population of live voles was buffered by treatments. Additionally, whenever voles reached densities triggering treatment, predator populations experienced strong declines. However, fox densities (0.01-0.015 ind. ha-1) were higher compared to FFR c (intermediate threshold, high AR), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities. The maximum numbers of dead voles under this FFR e was relatively low (highest around 15 ind.ha-1) but, due to frequent treatments, there was a steady replenishment of contaminated dead voles. This, in turn, induced mustelid and fox mortality and population declines (Supporting Information 3). Additionally, low availability of live voles triggered small mustelids mortality through starvation down to abundances similar to those resulting from AR use (Supporting Information 3).

\subsection{Comparison between farmer functional responses in terms of cost functions}

As expected, spreading AR generally reduced vole densities (Table 1) but the extent to which it also affected predators and the costs and benefits for conservation and farmer’s interests varied widely. 
The FFR b (high threshold, high AR) implied the lowest number of treatments (mean 0.87 year-1), while FFR d (intermediate threshold, low AR) resulted in the lowest amount of AR used (mean 14.5 kg ha-1 year-1). The FFR e (low threshold, low AR) had the lowest mustelid mortality due to AR (no instance in 40 years) but maximised starvation-induced mortality of mustelids (Table 1). Additionally, FFR e (low threshold, low AR) had the highest proportion of time with <50 voles ha-1, (96 % of study period). When considering the effectiveness in reducing mean vole densities, FFR e (low threshold, low AR) ranked highest followed by FFR c (intermediate threshold, high AR). The FFR d (intermediate threshold, low AR) and b (high threshold, high AR) allowed the highest mean densities of mustelids and foxes respectively. 
Comparison of these protocols indicates that FFR d (intermediate threshold and low AR) delivered the best compromise between farm production and conservation interests (Table 1). It had the lowest AR quantity applied (mean 14.5 kg ha-1 year-1) and the highest mean densities of mustelids (0.47 ha-1) and foxes (0.02 ha-1). The number of treatments was among the lowest values (1.6 treatment year-1). Its drawback was that it only suppressed voles at low densities (<50 voles ha-1 for 21% of time).  Conversely, emergent mean vole densities were low to intermediate (122 voles ha-1). 
Other FFR regimes were poorer compromises either because mustelids were kept at low densities (FFR  c, intermediate threshold, high AR, and FFR e, low threshold, low AR) or had relatively high mean vole densities (229 voles ha-1, FFR b, high threshold, high AR). In summary, FFR d (intermediate threshold, low AR) was the best compromise because it minimized the impact over predator populations contributing to the natural reduction of vole densities with the lowest quantity of AR applied for farm production interests.

\subsection{Influence of intra-guild predation}

Fig. 4 shows the system dynamics under FFR d, where successive farmer-regulated and mustelids-regulated phases occured. This simulation shows that the removal of foxes did not eliminate the successions of mustelids-regulated and farmer-regulated phases. However, the mustelid regulated period allowed short-term peaks of voles, suggesting the emergence of a classical one-predator - one-prey cycles interrupted by a farmer-regulated period. Without foxes, population dynamics of mustelids presented a more chaotic behaviour, while it presented regular cyclic pattern with fox occurrence. Therefore, this simple model suggests a stabilizing role of a generalist predator (foxes in this system) during the mustelid-regulated period. At the end of the mustelid-regulated period, foxes strongly contributed to vole mortality and, to a lesser degree, mustelid mortality. Consequently, the removal of foxes implied less predation over voles during the mustelid regulated period, and short-term vole releases from mustelid predation. The 2-year rolling mean of vole density (blue lines in Fig. 4) illustrates the change of regime from farmer-regulated to mustelids-regulated phases. Indeed, the amplitude of averaged vole densities (i.e., the amplitude of vole cycles for the 2-year rolling mean) was relatively stable at the beginning of farmer regulation period and then suddenly decreased to become minimal before sharply increasing, announcing the regime change. These changes in density amplitude may be used as an early-warning signal of the regime transition.

\section{Discussion} 
Considering that rodenticide kills not only voles but also their predators through secondary poisoning, our models show that AR profoundly changes the outcome of predator-prey dynamics involving vole, mustelid and fox populations beyond what mere intuition could elucidate. Our study reveals how the dual influences of the amount of pesticide spread and the vole density threshold triggering AR spread drives (i) pesticide spreading frequency, (ii) predation ecosystem service, and subsequently (iii) the control of pest outbreaks. Two types of a rich spectrum of emergent dynamics, including farmer or mustelid regulation deviating from classical predator prey dynamics arose because poisoned voles acted as “super-predators”.

\subsection{Modelling farmer regulation into a classical predator-prey system}

The threshold functional response of farmers deciding when to apply varying amounts of rodenticides according to prevailing vole density was crucial in selecting the emergent ecosystem dynamics, resulting in much variability in ecosystem and conservation and farming production interests. In the ecosystem our models depict, farmers spreading rodenticide not only depleted vole prey exploited by specialist and generalist predators but also created pulses of lethally or sub-lethally poisoned voles that became super-predators by poisoning their predators.  Arguably this set of ecological interactions has similarities with circumstances where a pathogen affecting prey species also infects predators, as in the case with the flea vectored plague (\textit{Yersinia pestis}) infecting prairie dogs (\textit{Cynomys} spp.) and black footed ferrets (Mustela nigripes) in central US \citep{Matchett2010}. However, to our knowledge, the behaviour of such tri-trophic model with multiple reciprocal interactions has not been explored. This is despite the obvious relevance to the management of the globally widespread circumstances where keystone small mammals are poisoned and may secondarily poison their predators \citep{Delibes-Mateos2011}.  
Under the “reference” scenario without AR spreading (FFR a), we assumed a predator-prey cycle which is a plausible pattern thoroughly explored theoretically \citep{Hanski2001}. There is no controversy on the role of small mustelids tracking vole dynamics, though it is not yet well understood whether predation may drive steep declines \citep{King2006}. Parameters of the reference scenario for our predator-prey model were realistic and tuned to generate population fluctuations similar to those observed in the studied cyclic system \citep{Delattre2009}.  The addition of pulses of rodenticide and their toxicokinetics in vole and predators are based on previous experiments with bromadiolone, a widely used AR, ensuring biologically realistic functional forms and their parameterization. Irrespective of the FFR considered, the frequency of vole cycles dramatically increased compared to the reference scenario, except during mustelid-regulated phases emerging under some FFR scenarios.

\subsection{How specialist predators may protect generalists from poisoning}

An interesting model behaviour was seen with FFR b (high vole density threshold and high AR) and FFR d (intermediate vole density threshold and low AR) with farmer- and mustelid-regulated phases alternating with low frequency. Such flipping between alternative states in population dynamics has been previously described in predator-prey model where weasels rely on a primary prey and entrain the dynamics of secondary prey \citep{Hanski1996} but not for the kind of indirect interaction we explore here. It further demonstrates that adding biological realistic complexity to simple models may drastically change the emergent properties of trophic interactions.
From these scenarios, we understand that the emergence of successive farmer and mustelid phases is neither driven by vole density threshold alone nor by AR amount, but instead by a subtle combination of both. The modelling description of these patterns uncovered the dual key roles of mustelids on fox dynamics, as intraguild competitors and as a vector for poisoning. This led to a surprising form of facilitation for foxes: mustelids protect foxes from collapses. The establishment of such a response can be described in 3 steps. Firstly, low mustelid densities inhibit their regulation of voles and contribute to farmer AR use. In line with empirical evidence, the latter directly impacts foxes by poisoning \citep{Jacquot2013}. Secondly, fox predation on mustelids is reduced, and with an intermediate AR amount, this allows mustelids to slowly recover. Vole outbreaks and subsequently farmer treatments are then gradually delayed, benefitting mustelids recovery. This is the point of transition from farmer to mustelids regulation, starting the third step: mustelids increase faster, suppressing vole densities and precluding the need for AR treatments, and eventually indirectly allowing fox population growth. Our finding that complexities in trophic interaction, induced by the poisoning of predator by poisoned prey, may cause the system to flip between alternative states is novel and robust. However, given we only explored deterministic versions of our models, any inference on the frequency of flipping between states should be cautious given the inherent stochastic nature of natural and farmland environments. If such dynamics occur within real farming systems, flipping between states is unlikely to emerge with regularity (~40 years) where many other factors impact population dynamics. Moreover, under our heuristic modelling framework, we have shown that the removal of fox population induced chaotic dynamics of voles and mustelids (Fig. 4). While generalists are known to have stabilising effect \citep{Hanski1991}, the benefit of specialist predators imparted to generalist predator and resulting increase in the prevalence of intraguild predation would be difficult to detect in empirical studies. Nevertheless, other generalist predators such as the endangered red kite (\textit{Milvus milvus}) which feed on voles opportunistically, occupy areas with bromadiolone treatments and are also affected by rodenticides \citep{Coeurdassier2014} and may therefore also benefit from the presence of mustelids in the ecosystem.

\subsection{How region-wide vole extirpation may inhibit ecosystem services}

Under FFR c (intermediate vole density threshold and high AR) and FFR e (low vole density threshold and low AR), the whole system was solely driven by farmer regulation, whereby the chronic use of AR completely suppressed the pest-regulation ecosystem service of predators. It has previously been shown empirically that repeated rodenticide treatments are highly detrimental to the populations of predators and reduce their densities  
\citep{Jacquot2013}. Secondary poisoning of predators is an established reality \citep{Berny2007}.  Through modelling, we formalised the insight that some poison deployment protocols, including those presently used in the empirical system which motivated our study, are counterproductive if employed on a large scale, suppressing natural predator regulation of pest rodents. It has been long known that poisoning rodents with AR permeates the food chain at peak abundance and achieves little in terms of protecting crops and may have strong deleterious impact (FFR b and c) \citep{Olea2009}. In Franche-Comté, a change in treatment protocols, from controlling voles at high densities to low-intermediate densities, has reduced the mortality of non-target species (including foxes) \citep{Jacquot2013}. Nevertheless, deployment regimes of pesticides that can contaminate the food chain should also enable periods of time which allow predator populations to rebound and avoid extirpation from the ecosystem. We have shown that, over time, farmers who strictly maintain voles at low density thresholds would suppress predation services provided by vole predators and instigate pesticide dependence. In addition, small mammals like voles certainly have ecosystem functions. Our results also suggest that the presence of small mustelids in ecosystems is beneficial for biodiversity conservation (see above) and agriculture interests. Given the importance of vole cycles and their trophic interactions, it is desirable to maintain vole population fluctuations of sufficient amplitude to maintain ecosystem processes.

\subsection{Managing rodents and ecosystems}

Presently, in Franche-Comté, farmers relying on bromadiolone alone can only treat pre-emptively when voles are at low densities (FFR e) whereas those using alternative methods without pesticide are allowed to spread AR in low quantity up to intermediate vole densities (FFR d). Spreading AR in low quantity seems superficially desirable, but our heuristic model, assuming an idealised homogenous landscape, show this is associated with frequent treatments. Consequently, it would induce a near permanent availability of a small number of intoxicated voles which, combined with low availability of non-contaminated voles, would reduce predator populations. Therefore, the extreme situation of using a low vole density threshold (FFR e) at a large scale is undesirable because it depletes the prey resources of foxes and mustelids and their populations. Triggering treatment at intermediate vole density with a low amount of AR (FFR d) allowed for temporal refuges, i.e. longer periods free of rodenticide necessary for predator densities to rebound while simultaneously avoiding episodes with high vole density, as required by farm production interests. Under a landscape management approach, such temporal refuges could be spatial refuges, with parts of the landscape free of pesticides where predator populations can recover.

Our key result and the basis for management prescriptions is that allowing for refuges where voles are not poisoned and allowed to persist at medium-high densities such that they can be exploited by mustelids is crucial for predator population recovery and preserving the ecosystem services mustelids deliver.  Treatment regimes allowing refuges seems compatible with both conservation and farming interests. A critical insight is to avoid potential side effects of chronic low-dose AR prescription (e.g., depletion of community services, stimulation of resistances), as is well known with antibiotics, by demanding regularly long-term period without treatment. However, combining chronic treatments and long periods free of AR may be difficult to achieve in real systems. Our model only considers temporal refuges, and the conceptualization of untreated areas as equivalent to triggering treatment at intermediate vole density cannot provide guidance on the size of these spatial refuges. Nevertheless, while management of voles is implemented at the scale of fields, mustelids and foxes roams over much larger areas  \citep{King2006}, such that large refuges with medium-high vole densities voles would be required while allowing to maintain low-voles density at very local scale.

\subsection{Conclusion and perspectives}

Our process-based model revealed pesticides that permeate the food chain upward can lead to diverse population dynamics with alternative states regulated by predators and farmers. It also shows that the practice currently promoted to use low-dose AR treatments at low vole density could be limited by the undesirable side-effects of leading to chronic application of AR on a large scale and the depletion of the vole predator community. This emerging question would benefit from a landscape modelling approach to characterize spatial refuges. We have also uncovered a counterintuitive mechanism whereby, owing to intraguild predation, mustelids could rescue foxes from poisoning. This suggest that contemporary Environmental Risk Assessment of pesticides that mostly consider one-species - one-compound experiments fail to capture the impact of pesticides on trophic links.  Assessing risk at the ecosystem level is empirically challenging such that process-based modelling can play a critical role.

\section*{Authors’ contributions}

XL conceived the initial idea; all authors developed the concept; VB and JF developed the models and led manuscript writing; VB implemented the model and ran simulations. GC contributed treatment protocols; JF, VB and MC explored model parameters; XL, PG and MC contributed critically to drafts; all authors gave final approval for publication.

\section*{Acknowledgments}

JF benefited from a Marie Skłodowska-Curie fellowship (European Commission, project "VOLES", 660718). VB was employed with this project funds. We are very grateful to Deon Roos for reviewing drafts. We thank Alessandro Massolo, Thibault Moulin and Francis Raoul for helpful suggestions. This work benefited from long-term data collected at Zone atelier (ILTER) Arc jurassien (http://zaaj.univ-fcomte.fr) and its financial support.

\section*{References}

%\bibliographystyle{elsarticle-harv}
\bibliographystyle{plainnat}
%\bibliographystyle{unsrt}
\bibliography{trophToxNTO.bib}

\end{document}

### 2019-02-20 trophToxNTO/R/mortality_predation.R
up package done for paper

@@ -1,0 +1,218 @@
+#' Repartition of the ingestion in mustelid and fox populations
+#'
+#' Take the output of the ode integration with deSolve and return a list of data.frame
+#' @param table_df A data.frame of ode system
+#' @param params Parameter of the system
+#' @param group.num Group number to apply the computation
+#'
+#' @return A list of data.frame
+#'
+#' @export
+#'
+#'
+
+Nbr.ingest=function(table_df,params,group.num=1){
+  # -------------------- Munipulation Table ---------------
+  df.mortality = dplyr::filter(table_df,group==group.num)
+
+  with(c(params,df.mortality),{
+
+    Nbr.V.ingest.M =  a.M * D.V/(a.M * D.V + a.Md * D.Vd) * (a.M * D.V + a.Md * D.Vd)/(1 + h.M * (a.M * D.V + a.Md * D.Vd)) * D.M
+
+    Nbr.V.ingest.F = a.F * D.V/(a.F * D.V + a.Fd * D.Vd + a.FM * D.M) * (a.F * D.V + a.Fd * D.Vd + a.FM * D.M)^2/(1 + h.F * (a.F * D.V + a.Fd * D.Vd + a.FM * D.M)^2) * D.F
+
+    Nbr.Vd.ingest.M = a.M * D.Vd/(a.M * D.V + a.Md * D.Vd) * (a.M * D.V + a.Md * D.Vd)/(1 + h.M * (a.M * D.V + a.Md * D.Vd)) * D.M
+
+    Nbr.Vd.ingest.F = a.F * D.Vd/(a.F * D.V + a.Fd * D.Vd + a.FM * D.M) * (a.F * D.V + a.Fd * D.Vd + a.FM * D.M)^2/(1 + h.F * (a.F * D.V + a.Fd * D.Vd + a.FM * D.M)^2) * D.F
+
+    Nbr.M.ingest.F = a.FM * D.M/(a.F * D.V + a.Fd * D.Vd + a.FM * D.M) * (a.F * D.V + a.Fd *  D.Vd + a.FM * D.M)^2/(1 + h.F * (a.F * D.V + a.Fd *D.Vd + a.FM * D.M)^2) * D.F
+
+    sum.ingest.M = Nbr.V.ingest.M+Nbr.Vd.ingest.M
+    sum.ingest.F=Nbr.V.ingest.F+Nbr.Vd.ingest.F+Nbr.M.ingest.F
+
+    return(data.frame(time,
+                      Nbr.V.ingest.M,
[…]

### 2019-02-20 trophToxNTO/R/ode_solver.R

@@ -1,6 +1,6 @@
 #' @title General Solver for Ordinary Differential Equations
 #'
-#' @description function from package \link{deDolve}.
+#' @description function from package \link{deSolve}.
 #' Solves a system of ordinary differential equations;
 #'  a wrapper around the implemented ODE solvers
 #'
@@ -11,22 +11,32 @@
 #' @importFrom deSolve ode
 NULL
 
-
+#' @title Arrange multiple grobs on a page
+#'
+#' @description function from package \link{gridExtra}.
+#' Set up a gtable layout to place multiple grobs on a page.
+#'
+#' @name grid.arrange
+#' @rdname grid.arrange
+#' @keywords internal
+#' @export
+#' @importFrom gridExtra grid.arrange
+NULL
 
 #' Dose-response curves for mortality
 #'
-#' Take a vecor containing : (i) the concentration in the body, (ii) the Lethal Dose of e species, (iii) the hill coeficient of the curve and (iv) the number of day to die
+#' Take a vecor containing : (i) the concentration in the body, (ii) the Lethal Dose of e species,
+#'  (iii) the hill coeficient of the curve and (iv) the number of day to die
+#'
 #' @param C A numeric vector of the concentration in the individual
-#' @param LD50 A numeric value of the Lethal Dose 50 (50% of the individual die)
-#' @param Hill.coef A numeric value coresponding to the Hill coefficient of the dose-response curve
+#' @param LD.50 A numeric value of the Lethal Dose 50 (50% of the individual die)
+#' @param H.coef A numeric value coresponding to the Hill coefficient of the dose-response curve
 #' @param days.death The number of days to die according to the LD50
 #'
-#' @return The Dose-Response curve
 #'
 #' @export
 #'
 #'
-
 dose.resp.mort=function(C,LD.50,H.coef,days.death){
   return(
     1/days.death*(1-1/(1+(C/LD.50)^H.coef))
@@ -57,6 +67,9 @@ intake.V=function(C,max.intake.C,D50.intake.C){
 #'
 #' Take Time, State, Pars as input, return derivative list
 #'
+#' @param Time
+#' @param State
+#' @param Pars
 #'
 #' @return A derivative list to integrate with deSolve package
 #'
@@ -213,7 +226,7 @@ pars  <- function(
 #'
 #' Take all parameters a vector and return a list
 #'
-#' @param pars.init List of parameters for the differential equations
+#' @param n.simu number of simulation to take into account
[…]

### 2019-02-20 trophToxNTO/R/plot.R

@@ -28,9 +28,6 @@ pkgLoad <- function(x){
 #'
 #'
 table_df = function(model.output){
-  pkgLoad("dplyr")
-  pkgLoad("tidyr")
-  pkgLoad("stringr")
 
   df=model.output%>%
     as.data.frame()%>%
@@ -54,11 +51,8 @@ table_df = function(model.output){
 #'
 #'
 tableau = function(model.output){
-  pkgLoad("dplyr")
-  pkgLoad("tidyr")
-  pkgLoad("stringr")
 
-  df.1=table_df(model.output)
+    df.1=table_df(model.output)
 
   df.g.2 = df.1 %>%
     gather(compartiment,value,-c(time,group))%>%

### 2019-02-20 trophToxNTO/R/sensitivity.R

@@ -1,0 +1,275 @@
+#' @title Table of sensitivity for set, min, max values
+#'
+#' @descritpiont Take the output of the ode integration with deSolve and return a data.frame
+#' with a set of measure on
+#' @param model_output the output of the ode solver
+#' @param params Parameter of the system
+#' @param key specific name of the parameters to consider
+#'
+#' @return A list of data.frame
+#'
+#' @importFrom zoo rollmean
+#'
+#' @export
+#'
+#'
+tab_sensitivity <- function(model_output, params, tab_ref){
+
+  tab_set <- table_df(model_output)
+  df_params <- as.data.frame(params)
+  df_params$Mean_V_set <- mean(tab_set$D.V)
+
+  df_params$Mean_M_set <- mean(tab_set$D.M)
+
+  df_params$Mean_F_set <- mean(tab_set$D.F)
+
+  df_params$V_inf50_set <- nrow(tab_set[tab_set$D.V <50,]) / nrow(tab_set)
+
+  df_params$nrb_AR_set <- nbr_spike(tab_set$C.soil)
+
+  tab_mortality_set <- Nbr.mortality(tab_set,params,group.num=1)
+
+  df_params$M_50AR_set <- sum(tab_mortality_set$Percent.M.mort.B > 0.5) / nrow(tab_mortality_set)
+
+  df_params$M_50natural_set <- sum(tab_mortality_set$Percent.M.mort.Nm > 0.5) / nrow(tab_mortality_set)
+
+  return(df_params)
+}
+
+
+
+tab_SSE <- function(model_output, params, tab_ref){
+  tab_set <- table_df(model_output)
+  mat_diff <- tab_ref[, c("C.F", "C.M", "C.V", "D.F", "D.M", "D.V", "D.Vd")] - tab_set[, c("C.F", "C.M", "C.V", "D.F", "D.M", "D.V", "D.Vd")]
+  params$SSE <- sum(mat_diff^2)/(dim(mat_diff)[1] * dim(mat_diff)[2])
+  return(params)
+}
+
+#' @title Table of sensitivity for set, min, max values
+#'
+#' @descritpiont Take the output of the ode integration with deSolve and return a data.frame
+#' with a set of measure on
[…]

### 2019-02-20 trophToxNTO/R/utils.R

@@ -1,0 +1,148 @@
+#' @title ggplot2 dual axes
+#'
+#' @description ggplot2 dual axis
+#'
+#' @details see https://gist.github.com/jslefche/e4c0e9f57f0af49fca87
+#'
+#' @importFrom gtable gtable_add_grob
+#' @importFrom grid grid.newpage
+#' @import ggplot2
+#'
+#' @examples
+#' # Create fake data.frame
+#' data.add.x = data.frame(
+#'   y1 = runif(100, 0, 100),
+#'   x1 = runif(100, 0, 100)
+#' )
+#' # Add second x-axis that scales with first
+#' data.add.x$x2 = (data.add.x$x1 + 50)^0.75
+#'
+#' # Create plots
+#' plot1.x = qplot(y = y1, x = x1, data = data.add.x)
+#' plot2.x = qplot(y = y1, x = x2, data = data.add.x)
+#'
+#' # Run function
+#' ggplot_dual_axis(plot1.x, plot2.x, "x")
+#'
+#' @export
+#'
+ggplot_dual_axis = function(plot1, plot2, which.axis = "x") {
+
+  plot2 <- plot2 + theme(axis.title.y = element_text(angle = 270))
+  # Update plot with transparent panel
+  #plot2 = plot2 + theme(panel.background = element_rect(fill = NA))
+
+  grid.newpage()
+
+  # Increase right margin if which.axis == "y"
+  if(which.axis == "y") plot1 = plot1 + theme(plot.margin = unit(c(0.7, 1.5, 0.4, 0.4), "cm"))
+
+  # Extract gtable
+  g1 = ggplot_gtable(ggplot_build(plot1))
+
+  g2 = ggplot_gtable(ggplot_build(plot2))
+
+  # Overlap the panel of the second plot on that of the first
+  pp = c(subset(g1$layout, name == "panel", se = t:r))
+
+  g = gtable_add_grob(g1, g2$grobs[[which(g2$layout$name=="panel")]], pp$t, pp$l, pp$b, pp$l)
+
+  # Steal axis from second plot and modify
+  axis.lab = ifelse(which.axis == "x", "axis-b", "axis-l")
+
+  ia = which(g2$layout$name == axis.lab)
+
+  ga = g2$grobs[[ia]]
+
+  ax = ga$children[[2]]
+
+  # Switch position of ticks and labels
[…]

### 2019-02-20 manuscript/MS_trophToxNTO.tex
update manuscript

@@
 \usepackage{lscape}
+
+% To manage cells in table
+\newcommand{\tabcell}[2][c]{%
+	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
+
 % MATHS
@@
 \usepackage{geometry} % define page margin
-\geometry{top=20mm,left=30mm,right=30mm,bottom=20mm} % margin from top, left, right and bottom
+\geometry{top=20mm,left=20mm,right=30mm,bottom=20mm} % margin from top, left, right and bottom
 
@@
 		Virgile Baudrot$^{1,2}$ 
-		Javier Fernandez-de-Simon$^1,3$
+		Javier Fernandez-de-Simon$^{1,3}$
 		Michael Coeurdassier$^1$,
-		Geoffroy Couval$^1,4$,
+		Geoffroy Couval$^{1,4}$,
 		Patrick Giraudoux$^1$,
-		Xavier Lambin$^5,6$		
+		Xavier Lambin$^{5,6}$		
 	}\par	
 \end{center}
-
-\begin{center}
+~\\
 $^1$ Université Bourgogne Franche-Comté - UMR CNRS 6249 Laboratoire Chrono-environnement, 25030 Besançon, France\\
@@
 $^5$ School of Biological Sciences, University of Aberdeen, Zoology Building, Tillydrone Avenue, Aberdeen AB24 2TZ, Scotland, UK\\
-	$^*$ Corresponding authors: virgile.baudrot@posteo.net or x.lambin@abdn.ac.uk
-\end{center}
+~\\
+$^*$ Corresponding authors: \textcolor{red}{virgile.baudrot@posteo.net or x.lambin@abdn.ac.uk}
 
-\paragraph{Running Title} Bufferscapes to preserve Non-Target Organism / Ecosystem services
+\paragraph{Running Title} \textcolor{red}{Bufferscapes to preserve Non-Target Organism / Ecosystem services}
 
@@
 We considered different FFR to assess AR upward transfer through the trophic web, considering the coupled dynamics of pulsed of AR spread and population dynamics.
-In all cases, parameterization units are on hectare-1 and day-1.
+In all cases, parameterization units are on hectare$^{-1}$ and day$^{-1}$.
 
@@
 
-We considered a tri-trophic system described by Fig.
-1 and equations (1-5); parameterization is provided in Supporting Information 1.
+We considered a tri-trophic system described by Figure~\ref{fig:scheme} and equations (1-5); parameterization is provided in Table~\ref{tab:parameterization}.
 Voles, denoted V, were the primary prey, Mustelids, M, intermediate predators, and Foxes, F, were top predators, consuming voles and mustelids.
@@
 \begin{equation}
+\left\{
 \begin{array}{l l}
@@
 \end{array}
+\right.
 \end{equation}
 
-The vole population followed a logistic growth rate, with rV  the maximal reproduction rate, fixed at $r_V = \ln(2 \times 600)/365$ per day, since montane water vole populations can increase from 0  to  600 individuals ha-1 or more \citep{Giraudoux1997} resulting in the equilibrium density being fixed at $K_V = 600$ individuals.
+The vole population followed a logistic growth rate, with $r_V$  the maximal reproduction rate, fixed at $r_V = \ln(2 \times 600)/365$ per day, since montane water vole populations can increase from 0  to  600 individuals ha$^{-1}$ or more \citep{Giraudoux1997} resulting in the equilibrium density being fixed at $K_V = 600$ individuals.
 %
@@
 \end{array}
-eq:functionResponse
+\label{eq:functionResponse}
 \end{equation}
 
-Parameterization of functional responses was estimated to fit the daily satiation level of predators for handling times, and the observed 5-6 year vole cycles for attack rates (Supporting Information 1).
+Parameterization of functional responses was estimated to fit the daily satiation level of predators for handling times, and the observed 5-6 year vole cycles for attack rates (Table~\ref{tab:parameterization}).
 We assumed foxes spent longer searching for voles than mustelids, based on each species diet and daily number of individuals captured.
@@
 
-Since we assumed small mustelids behave as specialist predators, we considered a numerical response linearly dependent on the functional response with parameter $\varepsilon_M$ (dimensionless) as conversion efficiency of prey into newborn predator (Supporting Information 2).
+Since we assumed small mustelids behave as specialist predators, we considered a numerical response linearly dependent on the functional response with parameter $\varepsilon_M$ (dimensionless) as conversion efficiency of prey into newborn predator (Supporting Information).
 The mustelid background mortality rate (i.e.
-all other reasons of death: ageing, disease, etc.) was $m_M$ (Supporting Information 2).
+all other reasons of death: ageing, disease, etc.) was $m_M$ parameterized has the inverse of life expectancy (Table~\ref{tab:parameterization}).
 We assumed foxes had a logistic growth rate function, parameterized with maximal growth rate $r_F = ln(3)/365$ and equilibrium density $K_F = 0.03$ [individuals.ha$^{-1}$] \citep{Ruette2003} without numerical response \citep{Weber2002}.
@@
 
-Fig. 1 represents the whole study system. Rodenticide is spread in grasslands during treatments, denoted $T_{Broma}(V)$, triggered by vole density $V$ (FFR described in Table 1).
+Figure~\ref{fig:scheme} represents the whole study system.
+Rodenticide is spread in grasslands during treatments, denoted $T_{Broma}(V)$, triggered by vole density $V$ (FFR described in Table~\ref{tab:results}).
 %
@@
 \begin{equation}
-\kappa_C= \dfrac{M_{in}\times C}{D_{in} + C}\dfrac{M_{in} \times C}{D_{in} + C} 
+\kappa_C= \dfrac{M_{in}\times C}{D_{in} + C}
 \end{equation}
@@
 Mustelids could feed on live voles $V$, or non-decomposed dead voles Vd and we assumed a Type 2 functional response adapted for a multi-species functional response \citep{Baudrot2016}.
-Mustelids ingested AR with absorption rate $\eta_M$ (ratio between biomasses of voles, $B_V$, and mustelids, $B_M$) and the total of ingested voles (alive $V$ and dead $V_d$) was defined by function $\Theta_M (V,V_d=$ (see Supporting Information 2), in equation (4) which provides the rate of AR ingested by a mustelid:
+Mustelids ingested AR with absorption rate $\eta_M$ (ratio between biomasses of voles, $B_V$, and mustelids, $B_M$) and the total of ingested voles (alive $V$ and dead $V_d$) was defined by function $\Theta_M (V,V_d=$ (Table~\ref{tab:parameterization}), in equation (4) which provides the rate of AR ingested by a mustelid:
 
 \begin{equation}
-C_V \times \eta_M \times \Theta_M(V,V_d)
+\text{Contaminant Intake Rate} = C_V \times \eta_M \times \Theta_M(V,V_d)
 \end{equation}
@@
 \begin{equation}
-\mu_X(C_X) = \dfrac{1}{\text{period of time}} \times \dfrac{1}{1+ (LD_{50}/C_X)^H}
+\mu_X(C_X) = \dfrac{1}{\text{experiment duration}} \times \dfrac{1}{1+ (LD_{50}/C_X)^H}
 \end{equation}
 
-where LD50 was the daily median lethal dose (50\% of population dying) with different values for voles (2 mg kg$^{-1}$), mustelids (2.1 mg kg$^{-1}$) and foxes (0.5 mg kg$^{-1}$, Supporting Information 1).
-$H$ is the Hill's coefficient, modulating the curve steepness and was estimated to fit the sparse data we have (Supporting Information 1).
+where $LD_{50}$ was the daily median lethal dose (50\% of population dying) with different values for voles (2 mg kg$^{-1}$), mustelids (2.1 mg kg$^{-1}$) and foxes (0.5 mg kg$^{-1}$, Table~\ref{tab:parameterization}).
+$H$ is the Hill's coefficient, modulating the curve steepness and was estimated to fit the sparse data we have (Table~\ref{tab:parameterization}).
 Mortality rates considered the duration of AR toxicity from experiments of up to 6 days \citep{Sage2010}.
@@
 We considered a range of realistic FFR spanning treatments during vole outbreaks only as a precautionary approach; in which treatments only takes place at intermediate or low vole density threshold.
-These scenarios are inspired by historic and contemporary protocols of bromadiolone use to control montane water voles in Franche-Comté, but also representative of practice globally (Table 1) \citep{Delattre2009}.
+These scenarios are inspired by historic and contemporary protocols of bromadiolone use to control montane water voles in Franche-Comté, but also representative of practice globally (Table~\ref{tab:results}) \citep{Delattre2009}.
 To check the influence of foxes and intraguild predation on the system’s dynamics, we simulated scenarios with and without foxes.
@@
 Our simulations tracked the linked vole-mustelid-fox dynamics for 40 years, after a “burn-in” period of 10 years to reduce dependency of results upon initial conditions, to observe several vole cycles and to characterise AR effects on these species population dynamics.
-This burn-in period also had AR treatment triggered at specific vole densities and with a given rodenticide quantity for each FFR (see Table 1).
+This burn-in period also had AR treatment triggered at specific vole densities and with a given rodenticide quantity for each FFR (see Table~\ref{tab:results}).
 The burn-in phase was selected according to a set of simulations with different initial conditions.
@@
 In order to assess the impact of FFR on both agriculture and conservation interests, we estimated the following cost functions: (i) Number of treatment events per FFR ; (ii) Cumulative amount of AR (kg) ; (iii) Proportion of time when the AR-induced mortality of mustelids higher than 50\% (i.e.
-lethal exposure profile killing 50\% of mustelid population) ; (iv) Proportion of time when the mortality of mustelids was higher than 50\% due to natural mortality; (v) Proportion of time when the vole density was below 50 voles ha-1, as a proxy for time when forage grass grows with low herbivore influence; (vi) Mean vole, mustelid and fox densities.
+lethal exposure profile killing 50\% of mustelid population) ; (iv) Proportion of time when the mortality of mustelids was higher than 50\% due to natural mortality; (v) Proportion of time when the vole density was below 50 voles ha$^{-1}$, as a proxy for time when forage grass grows with low herbivore influence; (vi) Mean vole, mustelid and fox densities.
 
@@
 Without AR (FFR a), vole dynamics were regulated by mustelid predation that gave rise to 6-year cycles (Figs.
-2 and 3; maximal vole and mustelid densities were around 600 ind.ha-1 and 2 ind.ha-1 respectively).
+2 and 3; maximal vole and mustelid densities were around 600 ind.ha$^{-1}$ and 2 ind.ha$^{-1}$ respectively).
 The fox population remained at its carrying capacity (i.e., 0.03 ind.
-ha-1, see Supporting Information 1).
+ha$^{-1}$, see Table~\ref{tab:parameterization}).
 
@@
 Populations of live and dead voles experienced high frequency fluctuations (around 2 peaks every year) driven by AR.
-As AR treatments were frequent, being triggered by voles peaks, contaminated dead voles were always abundant (peaks at 90 ind.ha-1), and mustelid and fox densities remained low (mustelids, converging to 0 ind.
-ha-1; foxes, 0.003-0.006 ind.
+As AR treatments were frequent, being triggered by voles peaks, contaminated dead voles were always abundant (peaks at 90 ind.ha$^{-1}$), and mustelid and fox densities remained low (mustelids, converging to 0 ind.ha$^{-1}$; foxes, 0.003-0.006 ind.
 ha$^{-1}$) with only short-term fluctuations (2 peaks every year, following vole cycles).
-With FFR e (low threshold, low AR), vole populations were maintained by farmers at around 50 voles ha-1 (Figs.
+With FFR e (low threshold, low AR), vole populations were maintained by farmers at around 50 voles ha$^{-1}$ (Figs.
 2e and 3e).
@@
 However, fox densities (0.01-0.015 ind.
-ha-1) were higher compared to FFR c (intermediate threshold, high AR), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities.
-The maximum numbers of dead voles under this FFR e was relatively low (highest around 15 ind.ha-1) but, due to frequent treatments, there was a steady replenishment of contaminated dead voles.
-This, in turn, induced mustelid and fox mortality and population declines (Supporting Information 3).
-Additionally, low availability of live voles triggered small mustelids mortality through starvation down to abundances similar to those resulting from AR use (Supporting Information 3).
+ha$^{-1}$) were higher compared to FFR c (intermediate threshold, high AR), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities.
+The maximum numbers of dead voles under this FFR e was relatively low (highest around 15 ind.ha$^{-1}$) but, due to frequent treatments, there was a steady replenishment of contaminated dead voles.
+This, in turn, induced mustelid and fox mortality and population declines (Supporting Information).
+Additionally, low availability of live voles triggered small mustelids mortality through starvation down to abundances similar to those resulting from AR use (Supporting Information).
 
@@
 
-As expected, spreading AR generally reduced vole densities (Table 1) but the extent to which it also affected predators and the costs and benefits for conservation and farmer’s interests varied widely.
-The FFR b (high threshold, high AR) implied the lowest number of treatments (mean 0.87 year-1), while FFR d (intermediate threshold, low AR) resulted in the lowest amount of AR used (mean 14.5 kg ha-1 year-1).
-The FFR e (low threshold, low AR) had the lowest mustelid mortality due to AR (no instance in 40 years) but maximised starvation-induced mortality of mustelids (Table 1).
-Additionally, FFR e (low threshold, low AR) had the highest proportion of time with <50 voles ha-1, (96 % of study period).
+As expected, spreading AR generally reduced vole densities (Table~\ref{tab:results}) but the extent to which it also affected predators and the costs and benefits for conservation and farmer’s interests varied widely.
+The FFR b (high threshold, high AR) implied the lowest number of treatments (mean 0.87 year$^{-1}$), while FFR d (intermediate threshold, low AR) resulted in the lowest amount of AR used (mean 14.5 kg.ha$^{-1}$.
+year$^{-1}$).
+The FFR e (low threshold, low AR) had the lowest mustelid mortality due to AR (no instance in 40 years) but maximised starvation-induced mortality of mustelids (Table~\ref{tab:results}).
+Additionally, FFR e (low threshold, low AR) had the highest proportion of time with <50 voles ha$^{-1}$, (96\% of study period).
 When considering the effectiveness in reducing mean vole densities, FFR e (low threshold, low AR) ranked highest followed by FFR c (intermediate threshold, high AR).
 The FFR d (intermediate threshold, low AR) and b (high threshold, high AR) allowed the highest mean densities of mustelids and foxes respectively.
-Comparison of these protocols indicates that FFR d (intermediate threshold and low AR) delivered the best compromise between farm production and conservation interests (Table 1).
-It had the lowest AR quantity applied (mean 14.5 kg ha-1 year-1) and the highest mean densities of mustelids (0.47 ha-1) and foxes (0.02 ha-1).
-The number of treatments was among the lowest values (1.6 treatment year-1).
-Its drawback was that it only suppressed voles at low densities (<50 voles ha-1 for 21% of time).
-Conversely, emergent mean vole densities were low to intermediate (122 voles ha-1).
-Other FFR regimes were poorer compromises either because mustelids were kept at low densities (FFR  c, intermediate threshold, high AR, and FFR e, low threshold, low AR) or had relatively high mean vole densities (229 voles ha-1, FFR b, high threshold, high AR).
+Comparison of these protocols indicates that FFR d (intermediate threshold and low AR) delivered the best compromise between farm production and conservation interests (Table~\ref{tab:results}).
+It had the lowest AR quantity applied (mean 14.5 kg.ha$^{-1}$.year$^{-1}$) and the highest mean densities of mustelids (0.47 ha$^{-1}$) and foxes (0.02 ha$^{-1}$).
+The number of treatments was among the lowest values (1.6 treatment year$^{-1}$).
+Its drawback was that it only suppressed voles at low densities (<50 voles ha$^{-1}$ for 21\% of time).
+Conversely, emergent mean vole densities were low to intermediate (122 voles ha$^{-1}$).
+Other FFR regimes were poorer compromises either because mustelids were kept at low densities (FFR  c, intermediate threshold, high AR, and FFR e, low threshold, low AR) or had relatively high mean vole densities (229 voles ha$^{-1}$, FFR b, high threshold, high AR).
 In summary, FFR d (intermediate threshold, low AR) was the best compromise because it minimized the impact over predator populations contributing to the natural reduction of vole densities with the lowest quantity of AR applied for farm production interests.
@@
 
-Fig.
-4 shows the system dynamics under FFR d, where successive farmer-regulated and mustelids-regulated phases occured.
+Figure~\ref{fig:intraguild} shows the system dynamics under FFR d, where successive farmer-regulated and mustelids-regulated phases occured.
 This simulation shows that the removal of foxes did not eliminate the successions of mustelids-regulated and farmer-regulated phases.
@@
 Consequently, the removal of foxes implied less predation over voles during the mustelid regulated period, and short-term vole releases from mustelid predation.
-The 2-year rolling mean of vole density (blue lines in Fig.
-4) illustrates the change of regime from farmer-regulated to mustelids-regulated phases.
+The 2-year rolling mean of vole density (blue lines in Figure~\ref{fig:intraguild}) illustrates the change of regime from farmer-regulated to mustelids-regulated phases.
 Indeed, the amplitude of averaged vole densities (i.e., the amplitude of vole cycles for the 2-year rolling mean) was relatively stable at the beginning of farmer regulation period and then suddenly decreased to become minimal before sharply increasing, announcing the regime change.
@@
 If such dynamics occur within real farming systems, flipping between states is unlikely to emerge with regularity (~40 years) where many other factors impact population dynamics.
-Moreover, under our heuristic modelling framework, we have shown that the removal of fox population induced chaotic dynamics of voles and mustelids (Fig.
-4).
+Moreover, under our heuristic modelling framework, we have shown that the removal of fox population induced chaotic dynamics of voles and mustelids (Figure~\ref{fig:intraguild}).
 While generalists are known to have stabilising effect \citep{Hanski1991}, the benefit of specialist predators imparted to generalist predator and resulting increase in the prevalence of intraguild predation would be difficult to detect in empirical studies.
@@
 
-\section*{References}
+% BIBLIOGRAPHY
+% \section*{References}
 
@@
 
+\newpage
+
+% TABLE 1: PARAMETERS DESCRIPTION
+
+\begin{table}
+\caption{Populations dynamics and toxicological parameters used in the model simulations.
+Notation ppm stands for parts-per-million [mg.kg$^{-1}$] and n.d denotes a dimensionless parameter.
+See description in Methods.}
+\label{tab:parameterization}
+\begin{tabular}{l p{7cm} l p{3cm}}
+\hline
+Parameters & Definitions & Units & Value\\
+\hline
+\multicolumn{2}{l}{Population dynamics parameters}  & & \\
+\hline
+%------- Vole dynamic
+%
+$r_V$ & Maximal growth rate of voles & ind.day$^{-1}$ & $\ln(2\times600)/365$ \\
+$K_V$ & Carrying capacity of voles & ind.ha$^{-1}$ & $600$  \\
+%
+% --------- Functional responses
+%
+$a_M$ & Attack rate of mustelids on voles & day$^{-1}$ & $1/30$ \\
+$h_M$ &  Handling time of mustelids on voles & day & $1/3.5$\\
+$a_{VF}$ & Attack rate of foxes on voles & day$^{-1}$ & $1/10$ \\
+$a_{MF}$ & Attack rate of foxes on mustelids & day$^{-1}$ & $1/10 \times 80/300$ \\
+$h_F$ &  Handling time of foxes on voles & day & $1/6$\\
+%
+% ------------ Mustelids dynamics
+%
+$\varepsilon_M$ & Conversion efficiency of ingested food & & $0.0025$ \\
+$m_M$ & Natural mortality rate of mustelids & day$^{-1}$ & $1/(0.8 \times 365)$\\
+%
+% --------------- Fox dynamic 
+%
+$r_F$ & Maximal growth rate of foxes & ind.day$^{-1}$ & $\ln(3)/365$ \\
+$K_F$ & Carrying capacity of foxes & ind.ha$^{-1}$ & $0.03$ \\
+%
+% -------- Population biomasses
+%
+$B_V$ & Mean biomass of a vole individual & g & $80$ \\
+$B_M$ & Mean biomass of a mustelid individual & g & $300$ \\
+$B_F$ & Mean biomass of a fox individual & g & $5800$ \\
+%
+$d$ & Degradation rate of dead vole & ind.day$^{-1}$ & $1/7$ \\
+%
+\hline
+\multicolumn{2}{l}{Toxicological parameters}  & & \\
+\hline
+%
+$T_{broma}(V)$ & Farmer input AR function of vole density & kg.ha$^{-1}$ & Scenarios \\
+$k_0$ & Persistence of AR in the field & day$^{-1}$ & $(0.106+0.057)/2$ \\
+%
+%
+$M_{in}$ & Maximal intake rate of vole & day$^{-1}$ & $6\times 1/5$ \\
+%
+$D_{in}$ & Half saturation intake rate of vole & & $100$\\
+%
+$LD_{50,V}$ & Lethal Dose for 50\% of voles & ppm & $2$ \\
+$H_V$ & Hill's coefficient dose-response curve &  n.d. & $4$ \\
+%
+$LD_{50,M}$ & Lethal Dose for 50\% of mustelids & ppm & $2.1$ \\
+$H_M$ & Hill's coefficient dose-response curve & n.d. & $4$  \\
+%
+$LD_{50,F}$ & Lethal Dose for 50\% of foxes & ppm & $0.5$\\
+$H_F$ & Hill's coefficient dose-response curve & n.d. & $2$ \\
+%
+$\eta_M$ & AR uptake rate in mustelids & n.d. & $0.5$ \\
+$\eta_F$ & AR uptake rate in foxes & n.d.& $0.5$\\
+%
+$k_{out,V}$ & Excretion rate of AR by voles & day$^{-1}$ & $0.4$\\
+$k_{out,M}$ & Excretion rate of AR by mutelids & day$^{-1}$ & $0.6$\\
+$k_{out,F}$ & Excretion rate of AR by foxes & day$^{-1}$ & $0.6$\\
+%
+\hline
+\end{tabular}
+\end{table}
+
+% TABLE 2: PARAMETERS DESCRIPTION
+
+\begin{table}
+\caption{Results on the cost functions for the comparison of treatment protocols.
+Column \textit{"FFR"} (farmer functional response) refers to the protocol denotation (a to e) and in brackets respectively: the vole density threshold, and the amount of AR/treatment in kg.
+The column \textit{"Nbr Treat."} refers to the number of treatment in simulation over the 40 years after the burn-in phase.
+\textit{"Total AR"} is the total use of anticoagulant rodenticides (bromadiolone in kg).
+\textit{"M, 50\% AR death"} is the proportion of time with mortality by AR over 50\% in mustelids population.
+\textit{"M, 50\% Nat.
+death"} is the proportion of time with mortality by natural reason (i.e., all except AR) over 50\% in mustelids population.
+\textit{"V < 50 ha$^{-1}$"} is the proportion of time with less than 50 voles ha$^{-1}$.
+Finally, \textit{"Mean V density"}, \textit{"Mean M density"} and \textit{"Mean F density"} are mean densities of voles, mustelids and foxes respectively.}
+\label{tab:results}
+\begin{tabular}{p{2.1cm} c c c c c c c c}
+\hline
+\tabcell{FFR \\ (dose, thresh.)} &
+\tabcell{Nbr. \\ Treat.} &
+\tabcell{Total  \\ AR} &
+\tabcell{M, 50\% \\ AR death} &
+\tabcell{M, 50\% \\ Nat. death} &
+\tabcell{V < \\ 50 ha$^{-1}$} &
+\tabcell{Mean V \\ density} & 
+\tabcell{Mean M \\ density} & 
+\tabcell{Mean F \\ density} \\
+\hline
+%%
+a (0,0) & 0& 0& 0& 1& 0.48 & 211.27 & 0.61 &0.03\\
+%
+b (500,20) & 35& 700& 0.06 &0.94 & 0.14 & 229.1 & 0.33 & 0.02\\
+%
+c (250, 20) & 81 & 1620 &0.11 & 0.89 & 0.28 & 105.49 & <0.01 & <0.01\\
+%
+d (250, 7.5)& 64 & 480 & 0.06 & 0.94 & 0.21 & 122.35 & 0.47 & 0.02\\
+%
+e (50,7.5) & 169 & 1267.5 & 0& 1 & 0.96 & 31.42 & <0.01 & 0.01 \\
+\hline
+\end{tabular}
+\end{table}
+
+% GRAPHIC SCHEME
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=.8\linewidth]{img/system_scheme.png}
+		\caption{Dynamics of mustelids (bold grey line) and red foxes (bold black line) after 50 years of simulation (10 years of burn-in period, shaded area, and 40 years considered here for the compute of results, estimation of
+			cost functions, etc.).
+Letters at the top-right part of each sub-graph corresponds to the results obtained from each farmer functional response.
+FR (dashed line) and MR (dotted line) indicate the farmer-regulated and mustelid-regulated periods of vole population dynamics.}
+		\label{fig:scheme}
+	\end{center}
+\end{figure}
+
+% GRAPHIC POPULATION DYANMICS
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/plot_join.png}
+		\caption{Rate of variation of population densities simulated over 50 years (10 years of burn-in period in grey area): (Left) Dynamics of mustelids (grey line) and red foxes (black line), (Right) Dynamics of live (grey line) and dead (black line) voles.
+Letters at the top-center part of each sub-graph corresponds to the results obtained from each farmer functional response.
+Red area and Green area indicate the farmer-regulated (FR) and mustelid-regulated (MR) periods of vole population dynamics respectively.}
+		\label{fig:populations}
+	\end{center}
+\end{figure}
+
+% GRAPHIC INTRAGUILD WITH AND WITHOUT FOXES
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\linewidth]{img/IGP_Graph.png}
+\caption{Effects of intraguild predation.
+Top graphics: daily densities of voles (in black) and 2-years rolling means of densities (in blue) for the farmer functional response d (i.e., 7.5kg of rodenticide at threshold density of 250 voles) with and without foxes (left and right respectively).
+Bottom graphics: stacked charts of ingestion and mortality proportion.}
+\label{fig:intraguild}
+\end{center}
+\end{figure}
+
 \end{document}

### 2019-02-21 trophToxNTO/R/scriptTOrun.R
add script to run for sensitivity analysis

@@ -1,0 +1,33 @@
+library(trophToxNTO)
+## --------------------------- TIME ------------------------
+times <- seq(from=0, to = 50*365, length = 1e4) ##  30*5 = 5 months
+## --------------------------- EVENEMENTS -------------------
+root.event=condition(Vole.Level=500,Broma.Level=20*50)[[1]]
+event.func=condition(Vole.Level=500,Broma.Level=20*50)[[2]]
+## ------------------------- RUN
+params = pars(n=1)
+init.p <- init(params$n)
+# # Set fix parameters
+# range_param = rbind(unlist(fix_pVar_pars(params,-0.1)), unlist(fix_pVar_pars(params,0.1)))
+# ls_params = as.list(as.data.frame(range_param))
+# ls_params_expand <- expand.grid(ls_params[2:ncol(range_param)])
+
+# # Random parameters
+ls_params_expand <- lapply(1:1e3, function(i) rand_pVar_pars(params))
+
+ls_randSens <- lapply(1:length(ls_params_expand), function(i){
+  tryCatch({
+    model_output = ode(y=init.p,
+                       times,
+                       ode.broma.tritrophic,
+                       parms = ls_params_expand[[i]],
+                       events = list(func = event.func, root = TRUE),
+                       rootfun = root.event)
+    df_temp <- tab_sensitivity(model_output, ls_params_expand[[i]])
+    return(df_temp)},
+    error=function(e) NULL)
+})
+# mat_randSens <- do.call("rbind", ls_randSens)
+df_randSens <- dplyr::bind_rows(ls_randSens)
+# save(df_randSens, file = paste0("data/df_randSens_", Sys.time(), "_", round(runif(1,1e10,1e11)), ".rda"))
+save(df_randSens, file = paste0("df_randSens_", Sys.time(), "_", round(runif(1,1e10,1e11)), ".rda"))

### 2019-03-22 manuscript/MS_trophToxNTO.tex
update files

@@
 ~\\
-$^*$ Corresponding authors: \textcolor{red}{virgile.baudrot@posteo.net or x.lambin@abdn.ac.uk}
+$^*$ Corresponding authors: \textcolor{gray}{virgile.baudrot@posteo.net or x.lambin@abdn.ac.uk}
 
-\paragraph{Running Title} \textcolor{red}{Bufferscapes to preserve Non-Target Organism / Ecosystem services}
+\paragraph{Running Title} \textcolor{gray}{TO CHANGE: Bufferscapes to preserve Non-Target Organism / Ecosystem services}
 
@@
 Foxes do not show numerical responses to vole abundance \citep{Weber2002} but they influence the food chain through occasional killing and consumption of bite-sized mustelids.
-Mustelids form a small proportion of fox diet (0-10\%) but their offtake could represent a significant portion of the population (reviewed in \citet{Lambin2017}).
+Mustelids form a small proportion of fox diet (0-10\%) but their offtake could represent a significant portion of the population (reviewed in \citet{Lambin2018}).
 
@@
 We considered a tri-trophic system described by Figure~\ref{fig:scheme} and equations (1-5); parameterization is provided in Table~\ref{tab:parameterization}.
-Voles, denoted V, were the primary prey, Mustelids, M, intermediate predators, and Foxes, F, were top predators, consuming voles and mustelids.
+Voles, denoted $V$, were the primary prey, Mustelids, $M$, intermediate predators, and Foxes, $F$, were top predators, consuming voles and mustelids.
 For each species, the instantaneous variation of population size over time is:
@@
 For that, we denoted $a_{VF}$ and $a_{MF}$ the fox attack rate on voles and mustelids respectively.
-The parameter hF was the handling time for foxes.
+The parameter $h_F$ was the handling time for foxes.
 
@@
 %
-Firstly, baits (50 mg.kg$^{-1}$  of bromadiolone, hereafter AR) were spread in grasslands at quantity 7.5 to 20 kg.ha$^{-1}$.
+Firstly, baits (50 mg kg$^{-1}$  of bromadiolone, hereafter AR) were spread in grasslands at quantity 7.5 to 20 kg ha$^{-1}$ day$^{-1}$.
+Notation day$^{-1}$ stands because of the daily time resolution.
 %
@@
 The proportion consumed per vole, with rate $\kappa(C)$, was assumed to be an increasing function.
-The function $\kappa(C)$ was characterized by a maximum ingestion $M_{in}$, and a half-saturation constant for ingestion $D_{in}$ in [mg.kg$^{-1}$]:
+The function $\kappa(C)$ was characterized by a maximum ingestion $M_{in}$, and a half-saturation constant for ingestion $D_{in}$ in [mg kg$^{-1}$]:
 
 \begin{equation}
-\kappa_C= \dfrac{M_{in}\times C}{D_{in} + C}
+\kappa(C)= \dfrac{M_{in}\times C}{D_{in} + C}
 \end{equation}
@@
 \begin{equation}
-\mu_X(C_X) = \dfrac{1}{\text{experiment duration}} \times \dfrac{1}{1+ (LD_{50}/C_X)^H}
+\mu_X(C_X) = \dfrac{1}{\text{experiment duration}} \times \left( 1- \dfrac{1}{1+ (LD_{50}/C_X)^H} \right) 
+\label{eq:muX}
 \end{equation}
@@
 
+\subsection{Numerical Simulation and Sensitivity Analysis}
+
+All numerical analysis have been done using the Open language R and particularly the ODE solver package \textit{deSolve} \citep{Soetaert2010}.
+Model implementation and code to run analysis are available on the Github repository \url{https://github.com/virgile-baudrot/trophToxNTO/}.
+%
+From simulations, a sensitivity analysis allowed to study \textit{how uncertainties in the model output are apportioned to different sources of uncertainty in the model input} \citep{Saltelli2019}.
+%
+A first approach consisted in changing parameter within a 10\% range around the original value \textit{one at a time}, and a second global approach consisted in using the first-order Sobol's sensitivity index $S_i$ defined as $S_i = \mathbb{V}(\mathbb{E}_{-i}(y \vert x_i)) / \mathbb{V}(y)$, where $\mathbb{V}(y)$ is the variance of $y$ when all factors are allowed to vary, $\mathbb{E}_{x\sim i}(y \vert x_i))$ the mean of $y$ when one factor is fixed \citep{Sobol1993}.
+We defined the domain of variation with a beta distribution within an hyperspace of plus or minus 10\% from the original value for all parameters.
+
 \section{Results}
@@
 Without AR (FFR a), vole dynamics were regulated by mustelid predation that gave rise to 6-year cycles (Figs.
-2 and 3; maximal vole and mustelid densities were around 600 ind.ha$^{-1}$ and 2 ind.ha$^{-1}$ respectively).
-The fox population remained at its carrying capacity (i.e., 0.03 ind.
-ha$^{-1}$, see Table~\ref{tab:parameterization}).
+2 and 3; maximal vole and mustelid densities were around 600 ind ha$^{-1}$ and 2 ind ha$^{-1}$ respectively).
+The fox population remained at its carrying capacity (i.e., 0.03 ind ha$^{-1}$, see Table~\ref{tab:parameterization}).
 
@@
 Populations of live and dead voles experienced high frequency fluctuations (around 2 peaks every year) driven by AR.
-As AR treatments were frequent, being triggered by voles peaks, contaminated dead voles were always abundant (peaks at 90 ind.ha$^{-1}$), and mustelid and fox densities remained low (mustelids, converging to 0 ind.ha$^{-1}$; foxes, 0.003-0.006 ind.
-ha$^{-1}$) with only short-term fluctuations (2 peaks every year, following vole cycles).
+As AR treatments were frequent, being triggered by voles peaks, contaminated dead voles were always abundant (peaks at 90 ind ha$^{-1}$), and mustelid and fox densities remained low (mustelids, converging to 0 ind ha$^{-1}$; foxes, 0.003-0.006 ind ha$^{-1}$) with only short-term fluctuations (2 peaks every year, following vole cycles).
 With FFR e (low threshold, low AR), vole populations were maintained by farmers at around 50 voles ha$^{-1}$ (Figs.
@@
 Additionally, whenever voles reached densities triggering treatment, predator populations experienced strong declines.
-However, fox densities (0.01-0.015 ind.
-ha$^{-1}$) were higher compared to FFR c (intermediate threshold, high AR), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities.
-The maximum numbers of dead voles under this FFR e was relatively low (highest around 15 ind.ha$^{-1}$) but, due to frequent treatments, there was a steady replenishment of contaminated dead voles.
+However, fox densities (0.01-0.015 ind ha$^{-1}$) were higher compared to FFR c (intermediate threshold, high AR), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities.
+The maximum numbers of dead voles under this FFR e was relatively low (highest around 15 ind ha$^{-1}$) but, due to frequent treatments, there was a steady replenishment of contaminated dead voles.
 This, in turn, induced mustelid and fox mortality and population declines (Supporting Information).
@@
 As expected, spreading AR generally reduced vole densities (Table~\ref{tab:results}) but the extent to which it also affected predators and the costs and benefits for conservation and farmer’s interests varied widely.
-The FFR b (high threshold, high AR) implied the lowest number of treatments (mean 0.87 year$^{-1}$), while FFR d (intermediate threshold, low AR) resulted in the lowest amount of AR used (mean 14.5 kg.ha$^{-1}$.
-year$^{-1}$).
+The FFR b (high threshold, high AR) implied the lowest number of treatments (mean 0.87 year$^{-1}$), while FFR d (intermediate threshold, low AR) resulted in the lowest amount of AR used (mean 14.5 kg ha$^{-1}$ year$^{-1}$).
 The FFR e (low threshold, low AR) had the lowest mustelid mortality due to AR (no instance in 40 years) but maximised starvation-induced mortality of mustelids (Table~\ref{tab:results}).
-Additionally, FFR e (low threshold, low AR) had the highest proportion of time with <50 voles ha$^{-1}$, (96\% of study period).
+Additionally, FFR e (low threshold, low AR) had the highest proportion of time with < 50 voles' ind ha$^{-1}$, (96\% of study period).
 When considering the effectiveness in reducing mean vole densities, FFR e (low threshold, low AR) ranked highest followed by FFR c (intermediate threshold, high AR).
@@
 Comparison of these protocols indicates that FFR d (intermediate threshold and low AR) delivered the best compromise between farm production and conservation interests (Table~\ref{tab:results}).
-It had the lowest AR quantity applied (mean 14.5 kg.ha$^{-1}$.year$^{-1}$) and the highest mean densities of mustelids (0.47 ha$^{-1}$) and foxes (0.02 ha$^{-1}$).
+It had the lowest AR quantity applied (mean 14.5 kg ha$^{-1}$.year$^{-1}$) and the highest mean densities of mustelids (0.47 ha$^{-1}$) and foxes (0.02 ha$^{-1}$).
 The number of treatments was among the lowest values (1.6 treatment year$^{-1}$).
@@
 
+\subsection{Sensitivity of model under this parameterization}
+
+Figure~\ref{fig:sobol_sensitivity} and Figure~\ref{fig:sensitivity} show the sensitivity of each parameter used in the model defined in d (i.e., 7.5kg of AR at threshold density of 250 voles).
+%
+Figure~\ref{fig:sobol_sensitivity} represent the global sensitivity where we computed a first order sensitivity index \citep{Sobol1993, Saltelli2019} given the contribution of each parameter to the variance of the model.
+%
+Figure~\ref{fig:sensitivity} is a one at time analysis \citep{Saltelli2019} where parameters are move +/- 10\% of the fixed value as defined in scenario d (see Table~\ref{tab:parameterization}).
+
+Crossing all cost functions to test the sensitivity of model parameters, the influence of each parameter appears quite homogeneous.
+That is there is no useless parameter or in return one that would lead all.
+%
+Taken cost functions separately, the sensitivity to parameters present some differences.
+%
+Figure~\ref{fig:sobol_sensitivity} shows an homogeneity of parameter sensitivity on the mean of mustelids density, revealing the complexity underlying this latent variable.
+
+% For instance, the vole carrying capacity $K_V$ has a strong influence on density of predators and the number of treatments, but few on the vole density (i.e., mean vole density and period of time with less than 50 vole/ha) reflecting the fact that vole population is totally regulated by predation (i.e., density of predators) or poisoning (i.e.
+number of treatment) but not by resource of voles.
+%
+% On the contrary, lethal dose for 50\% of mustelids, $LD_{50,M}$, has a strong influence on mean vole density and is weaker for other cost functions.
+%
+% From the sensitivity analysis, the drivers of the mean density of mustelids are the mustelids' resources, which is the carrying capacity of vole, $K_V$, and the parameters reflecting intrinsic mustelids parameter as the natural mortality rate $m_M$, and toxicological parameters of mustelids as the Hill's coefficient, the lethal dose for 50\%, and those of voles as the half saturation intake rate of contaminant for voles $D_{in}$ and the vole excretion rate of AR, denoted $k_{out,V}$.
+
 \section{Discussion}
+
 Considering that rodenticide kills not only voles but also their predators through secondary poisoning, our models show that AR profoundly changes the outcome of predator-prey dynamics involving vole, mustelid and fox populations beyond what mere intuition could elucidate.
@@
 \caption{Populations dynamics and toxicological parameters used in the model simulations.
-Notation ppm stands for parts-per-million [mg.kg$^{-1}$] and n.d denotes a dimensionless parameter.
+Notation ppm stands for parts-per-million [mg kg$^{-1}$] and n.d.
+denotes a dimensionless parameter.
 See description in Methods.}
@@
 %
-$r_V$ & Maximal growth rate of voles & ind.day$^{-1}$ & $\ln(2\times600)/365$ \\
-$K_V$ & Carrying capacity of voles & ind.ha$^{-1}$ & $600$  \\
+$r_V$ & Maximal growth rate of voles & day$^{-1}$ & $\ln(2\times600)/365$ \\
+$K_V$ & Carrying capacity of voles & ind ha$^{-1}$ & $600$  \\
 %
@@
 %
-$\varepsilon_M$ & Conversion efficiency of ingested food & & $0.0025$ \\
+$\varepsilon_M$ & Conversion efficiency of ingested food & n.d. & $0.0025$ \\
 $m_M$ & Natural mortality rate of mustelids & day$^{-1}$ & $1/(0.8 \times 365)$\\
@@
 %
-$r_F$ & Maximal growth rate of foxes & ind.day$^{-1}$ & $\ln(3)/365$ \\
-$K_F$ & Carrying capacity of foxes & ind.ha$^{-1}$ & $0.03$ \\
+$r_F$ & Maximal growth rate of foxes & day$^{-1}$ & $\ln(3)/365$ \\
+$K_F$ & Carrying capacity of foxes & ind ha$^{-1}$ & $0.03$ \\
 %
@@
 %
-$T_{broma}(V)$ & Farmer input AR function of vole density & kg.ha$^{-1}$ & Scenarios \\
-$k_0$ & Persistence of AR in the field & day$^{-1}$ & $(0.106+0.057)/2$ \\
+$T_{broma}(V)$ & Farmer input AR function of vole density & mg ha$^{-1}$ day$^{-1}$ & Scenarios \\
+$k_0$ & Disappearance of AR in the field & day$^{-1}$ & $(0.106+0.057)/2$ \\
 %
 %
-$M_{in}$ & Maximal intake rate of vole & day$^{-1}$ & $6\times 1/5$ \\
+$M_{in}$ & Maximal intake rate of vole & ppm day$^{-1}$ & $6\times 1/5$ \\
 %
-$D_{in}$ & Half saturation intake rate of vole & & $100$\\
+$D_{in}$ & Half saturation intake rate of vole & ppm & $100$\\
 %
@@
 
+% GRAPHIC SENSITIVITY ANALYSIS
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/sobol_heatMap.png}
+		\caption{First-order Sobol's sensitivity index (denoted $S_p$ for parameter $p$) providing the global sensitivity of cost functions to parameters of the model \citep{Sobol1993} for the model d (i.e., 7.5kg of AR at threshold density of 250 voles).
+Analysis is based on 4200 simulations.
+Values reflect the expected fractional reduction in the variance for each cost functions that would be achieved if a specific parameter is fixed.
+If $S_p = 1$, the sensitivity is not influenced by the parameter $p$.}
+		\label{fig:sobol_sensitivity}
+	\end{center}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/Sensitivity_pm10_2.png}
+		\caption{Minimal and maximal values of cost functions with variation of + and - 10\% of change for each parameter value with other parameter values fixed, as set in Table~\ref{tab:parameterization}.
+Definition of parameters are in Table~\ref{tab:parameterization} (dd$_X$ are experiment duration as defined by equation \ref{eq:muX}), and cost function are defined in Table~\ref{tab:results}.
+Column SSE stands for Sum of Square Error and is scaled in $[0,1]$.}
+		\label{fig:sensitivity}
+	\end{center}
+\end{figure}
+
 \end{document}

### 2019-03-22 manuscript/SupMat_trophToxNTO.tex
update files

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%\RequirePackage{lmodern}	% font set: Latin Modern
\RequirePackage{charter}	% font set: Charter

\usepackage[french,english]{babel}

% Package ADDED
\usepackage{xcolor,graphics,graphicx}

% For citations
\usepackage{csquotes}

% For line spacing
\usepackage{setspace}
\linespread{1.3}

% To rotate a table or a figure
\usepackage{lscape}
% MATHS
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{empheq}

% TAILLE de la page
\usepackage{geometry} % define page margin
\geometry{top=20mm,left=30mm,right=30mm,bottom=20mm} % margin from top, left, right and bottom

% COMPTEUR de ligne
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
\usepackage{caption}
\usepackage{lineno}
\linenumbers

% BIBLIOGRAPHIE
%\usepackage[numbers]{natbib}
\usepackage[authoryear]{natbib}
\usepackage{hyperref}

\begin{document}

% USED TO ADD "S" in front of Table, Figure, Equation
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}%
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}%
\setcounter{equation}{0}
\renewcommand{\theequation}{S\arabic{equation}}%

% ------------- TITLE PAGE

\begin{center}
	\huge\textbf{
	Supporting Information 1
	}\par
\end{center}

\begin{center}
	\Large\textbf{
		Trophic transfer of anticoagulant rodenticides while managing rodent pests: the fine line between predator-prey regulation and pesticide-pest regulation
	}\par
\end{center}

\vspace{.5cm}

\begin{center}
	\large\textbf{
		Virgile Baudrot$^{1,2}$ 
		Javier Fernandez-de-Simon$^1,3$
		Michael Coeurdassier$^1$,
		Geoffroy Couval$^1,4$,
		Patrick Giraudoux$^1$,
		Xavier Lambin$^5,6$		
	}\par	
\end{center}
~\\
$^1$ Université Bourgogne Franche-Comté - UMR CNRS 6249 Laboratoire Chrono-environnement, 25030 Besançon, France\\
$^2$ BioSP, INRA, 84000 Avignon, France\\
$^3$ IREC. Instituto de Investigación en Recursos Cinegéticos\\
$^4$ FREDON Franche-Comté, Espace Valentin Est, 12, Rue de Franche-Comté - Bât E, 25480 Ecole-Valentin, France\\
$^5$ School of Biological Sciences, University of Aberdeen, Zoology Building, Tillydrone Avenue, Aberdeen AB24 2TZ, Scotland, UK\\
~\\
$^*$ Corresponding authors: virgile.baudrot@posteo.net or x.lambin@abdn.ac.uk

\newpage

% -----------------------------

\section{Model derivation}

% GRAPHIC SCHEME

\begin{figure}
	\begin{center}
		\includegraphics[width=.8\linewidth]{img/system_scheme.png}
		\caption{Dynamics of mustelids (bold grey line) and red foxes (bold black line) after 50 years of simulation (10 years of burn-in period, shaded area, and 40 years considered here for the compute of results, estimation of 
			cost functions, etc.). Letters at the top-right part of each sub-graph corresponds to the results obtained from each farmer functional response. FR (dashed line) and MR (dotted line) indicate the farmer-regulated and mustelid-regulated periods of vole population dynamics.}
		\label{fig:scheme}
	\end{center}
\end{figure}

\subsection{Dynamic of populations for the tri-trophic dynamic}
%
The population dynamics of a tri-trophic system is commonly described as follow:
\begin{equation}
\left\{
\begin{array}{l l}
\dfrac{dV}{dt} & =  r_V V \left( 1- \dfrac{V}{K_V}\right) - \Phi_{V,M}(V)M - \Phi_{V,F}(V,M)F\\[.3cm]
%
\dfrac{dM}{dt} & = \varepsilon_M \Phi_{V,M}(V)M - m_M M - \Phi_{M,F}(V,M)F\\[.3cm]
%
\dfrac{dF}{dt} & = F r_F \left( 1- \dfrac{F}{K_F}\right)
\end{array}
\right.
\label{eq:simpleSystem}
\end{equation}

As described in the manuscript, the vole population, $V$ followed a logistic growth rate, with $r_V$  the maximal reproduction rate, fixed at $r_V = \ln(2 \times 600)/365$ in [day$^{-1}$], since montane water vole populations can increase from 0  to  600 individuals ha$^{-1}$ or more in a year \citep{Giraudoux1997} resulting in the equilibrium density being fixed at $K_V = 600$ individuals.

\subsubsection{About the logistic model}

The idea behind this parametrization is the use of a classical Malthusian equation with $r_V$ in [day$^{-1}$] and $V$ in [ind ha$^{-1}$]:

\begin{equation}
\frac{dV}{dt} = r_V V \quad \Rightarrow \quad
V(t) = V(0) \exp(r_V t) \quad \Rightarrow \quad
r_V = \ln\left( \frac{V(t)}{V(0)} \right) \times \frac{1}{t}
\end{equation}
So to have the maximal growth rate $r_V$, we need the greatest $V(t)$ (measured at 600 ind ha$^{-1}$ but supposed to be more in theoretical optimal condition) obtained in the minimum of time (less than a year). 
%
Then, to add the carrying capacity $K_V$ in [ind$^{-1}$], we use the logistic equation and assumed the carrying capacity to be the one measured in the environment:
\begin{equation}
\frac{dV}{dt} = r_V V \left( 1 - \frac{V}{K_V}\right)
\end{equation}

The estimation of $r$ using the Malthusian function and then applying the carrying capacity $K$ can be seen as non consistent because of the dependency of $r_V$ to $K_V$. 
%
For isntance, in \citet{Ginzburg1992}, the author says that going from $dV/dt = r_V V$ to $dV/dt = r_V V (1-V/K_V)$ would be correct if $r$ was independent of resource limitation $K_V$:

\textit{"Two populations  of a species live in environments  identical in all  respects, that is, they  have the same resource availability  and any  other imaginable characteristics. One of the two is subject to  higher mortality. There is no doubt  that the population growth curves in these two cases will be different. Are the final equilibrium abundances different? Most ecologists will answer that the equilibrium values should be the same and that the higher rate of reproduction just means that the population will ‘get there faster’, but reach the same level nevertheless".}

Then, \citet{Ginzburg1992} states that their is two ways to reflect the extra mortality: either assuming (a) $dV/dt = r_V V(1-V/K_V) - mV$ or (b) $dV/dt = (r_V-m_V)V(1-V/K_V)$.
%
In equation (a), the equilibirum depend on $r$: $V^* = (r_V-m_V) K_V /r_V$, so is wrong with the un-changed equilibrium assumption.
%
And re-writing (b) gives: $dV/dt = rV(1-V/K_V) - m_V V(1-V/K_V)$. Since $- m_V V(1-V/K_V)$ reflects an environmental pressure, it should not be positive when $V>K_V$. Meaning, when the population is over it’s carrying capacity, we are going to growth even faster and to infinity.
%
In the present manuscript, we made assumption (a), and we assume that $r$ is not independent of resource limitation $K_V$. It is for instance well known that vole reproduction rate is linked to the state of resource at the time of reproduction \citep{Pinot2016}.

Many authors respond to Ginzburg \citep{Olson1992, Watkinson1992, Mackenzie1992, Berryman1992} and later \citep{Gabriel2005}.

In his response, \citet{Olson1992} says: \textit{“While the logistic model is often defined as non-mechanistic, it has the attractive feature of describing population growth with only two parameters, $K$ and $r$. $K$ is universally defined as the population-carrying capacity. However, there is much disagreement about the definition of  $r$ and this disagreement has affected conclusions derived from the use of the logistic equation.”}
%
\citet{Olson1992} suggests that the problem is the intuition assigned to the r-independence to equilibirum density $K$.
%
Again in our line, \citet{Watkinson1992} argues that \textit{"Most ecologists, he asserts, would believe that the equilibrium values of these two populations should be the same. In that case I must be one of the minority who does not believe this to be so."}

Another answer by \citet{Berryman1992} proposed to rewrite the model in order to see the link between $r$ and $K$. The approach by \citet{Berryman1992} is interesting as it introduces a relevant effect of considering  the exploitation rate (link to the growth rate) within the carrying capacity, meaning a carrying capacity with its on dynamic (e.g., with its resilience depending on how exploitation happen), and non an extra-system static one.

\subsubsection{Predation of voles, and predator dynamics}

The vole population was preyed upon by mustelid and fox populations, denoted $M$ and $F$ respectively in \eqref{eq:simpleSystem}. The vole consumption rate at different vole densities was described by functional responses ($\Phi_{V,M}$ for mustelids, $\Phi_{V,F}$  for foxes).

\begin{equation}
\begin{array}{l l}
\Phi_{V,M}(V)M = \dfrac{a_M V}{1 + h_M a_M V} \\[.3cm]
%
\Phi_{V,F} = \dfrac{a_{VF} V}{a_{VF}V + a_{MF}F} \times \dfrac{(a_{VF}V + a_{MF} M)^2}{1 + h_F(a_{VF}V + a_{MF} M)^2}\\[.3cm]
%
\Phi_{M,F}(V,M)F = \dfrac{a_{MF}V}{a_{VF}V + a_{MF}F} \times \dfrac{(a_{VF}V + a_{MF}M)^2}{1+h_F(a_{VF}V + a_{MF}M)^2}
\end{array}
\label{eq:functionResponse}
\end{equation}

We treated small mustelids as vole specialist predators \citep{King2006}, assuming a Holling Type 2 functional response with attack rate $a_M$ in [day$^-1$] and handling time $h_M$ in [day] (equation \eqref{eq:functionResponse}). We then represented foxes feeding on voles and mustelids by a multi-species functional response derived from Holling Type 3, referring to generalist feeding behaviour \citep{Baudrot2016}. For that, we denoted $a_{VF}$ and $a_{MF}$ (both in  [day$^-1$]) the fox attack rate on voles and mustelids respectively. The parameter $h_F$ was the handling time for foxes in [day].

Consequently, the functional responses $\Phi_{X,Y}$ are all in [day$^{-1}$].  And the functions:

$\dfrac{dV}{dt}  =  r_V V \left( 1- \dfrac{V}{K_V}\right) - \Phi_{V,M}(V)M - \Phi_{V,F}(V,M)F$ and $\dfrac{dF}{dt} = F r_F \left( 1- \dfrac{F}{K_F}\right)$ are also consistent, all part being in [ind ha$^{-1}$ day$^{-1}$].

\subsubsection{Conversion efficiency and mortality rate in mustelid population}

The parameterization of the conversion efficiency of ingested food into new born, $\varepsilon_M$ in [n.d.] (non dimensional), and the mortality rate, $m_M$ in [day$^{-1}$], in equation \eqref{eq:simpleSystem} is not straightforward since we have to take into account the mortality rate due to starvation.
% 
In the situation of no starvation, ageing, diseases and predation of mustelids are the main reason of death. In the situation of no prey availability at initial condition, then the mortality rate of mustelids should be equal to the mortality rate without food. 
Using the simplest scheme of Dynamic Energy Budget (DEB) theory (seeFigure~\ref{fig:DEB_scheme}), the ingested food is assimilated for a part, we denote $\alpha$ and non-assimilated for another part, $(1- \alpha)$. Then, the assimilated part goes to a reserve (or directly use) in either the somatic maintenance or in the maturity maintenance (to produce reproductive tools). In DEB, a proportion $\kappa$ from the reserve is the part allocated to somatic maintenance [n.d.] and a proportion $1-\kappa$ is addressed to the maturity maintenance.

\begin{figure}
	\begin{center}
		\includegraphics[width=.5\linewidth]{img/DEB_scheme.png}
		\caption{Representation of the standard dynamic energy budget (DEB model \citep{Sousa2010}}
		\label{fig:DEB_scheme}
	\end{center}
\end{figure}

Figure~\ref{fig:DEB_scheme} represents the standard dynamic energy budget (DEB) model \citep{Sousa2010}.
%
Then, assuming the generic functional response $\Phi_{V,M}(V)$, the reproduction rate of the predator is like:
\begin{equation}
\Phi_{V,M}(V) \alpha (1 - \kappa) \varepsilon'
\end{equation}
where $\varepsilon'$ in [n.d.] is the conversion efficiency of daily assimilated food for one individual, dedicated for reproduction into newborn mustelids.

Then, $\alpha \kappa \Phi_{V,M}$ is the part allocated to somatic maintenance structure. Parameters $\alpha$ and $\kappa$ are non-dimensional [n.d.].

We can then take into account the starvation process in the mortality rate, with a food-independent mortality rate $m_{M, -\Phi}$ in [day$^{-1}$] and a part depending on food availability $m_{M, +\Phi} = \sigma \alpha \kappa \Phi_{V,M}$ in [day$^{-1}$], with $\sigma$ [n.d.] a conversion parameter.

Then, the dynamic of Mustelid given in the manuscript equation \eqref{eq:simpleSystem} could be rewritten as:
\begin{equation}
\frac{dM}{dt} = \Phi_{V,M}(V) \alpha (1 - \kappa) \varepsilon' M - \left( m_{M, -\Phi}  + \sigma \alpha \kappa \Phi_{V,M} \right) M - \Phi_{M,F}(V, M) F
\end{equation}

Which is equal to:
\begin{equation}
\frac{dM}{dt} = \left(\alpha (1 - \kappa) \varepsilon' - \sigma \alpha \kappa  \right) \Phi_{V,M}(V)  M - m_{M, -\Phi}  M  - \Phi_{M,F}(V, M) F
\end{equation}

Therefore, if we set  the food-independent mortality rate $m_M = m_{M, -\Phi}$ and the conversion efficiency of ingested food $\varepsilon_M=\alpha ((1 - \kappa) \varepsilon' - \sigma  \kappa ) $ , we get back to the equation provided in model \eqref{eq:simpleSystem} but with a clearer definition of $\varepsilon_M$.

This is why we parameterized $m_M$ in [day$^{-1}$] as the inverse of life expectancy in [day$^{-1}$], and $\varepsilon_M$ in [n.d.] as an unclear parameter derived from the null model (i.e., without AR) to reflect observed population dynamics.

In any case, the dimension of the equation $\dfrac{dM}{dt}  = \varepsilon_M \Phi_{V,M}(V)M - m_M M - \Phi_{M,F}(V,M)F$ is well defined, since $\varepsilon_M \Phi_{V,M}(V)$ is in [day$^{-1}$].

\subsection{Dynamic of populations with AR}

After spreading by farmer, in the model AR is degraded or transferred to voles, which are exposed to the environmental contaminant mainly through ingestion of contaminated foodstuffs (i.e., trophic transfer), and AR accumulates in their tissues and is release with rate $k_{out,V}$.
%
Contaminants in an individual, the body burden, are commonly measured in concentration per mass of individual, [$\mu$g g$^{-1}$ ind$^{-1}$] (equivalent to [mg kg$^{-1}$ ind$^{-1}$] and denoted [ppm ind$^{-1}$]).
%
Predators are only exposed to contaminant through the consumption of voles, a trophic transfer, so as a function of functional response.
\\

To consider both toxicological and population dynamics within the same equation, we convert population dynamics into dynamics of their biomasses as in \citep{Huang2015EcoTox, Baudrot2018}.

We denote $V$, $V_d$ $M$, $F$ the densities of voles, dead voles, mustelids and foxes respectively ; $B_V$, $B_{V_d}$, $B_M$ and $B_F$ the mean biomass of an individual of those species in [g ind$^{-1}$], and $B_{T,V}$, $B_{T,V_d}$, $B_{T,M}$ and $B_{T,F}$ the total biomass of the population in the space unit we used, so in [g ha$^{-1}$]. In other words, $B_{T,X} = B_X \times X$. The notation $C_V$, $C_{V_d}$ $C_M$ and $C_F$ holds for the mean concentration of the contaminant in one individual, commonly called the body burden of the prey in ppm [mg kg$^{-1}$].
%
The growth function of the prey population is a function $g_X(X, C_X)$ in [day$^{-1}$] depending on the population density $X$ and the concentration of the contaminant $C_X$.
%
The dose-response curve $\mu_X(C_X)$ is defined by a log-normal cumulative distribution function as in \cite{Loos2010}.

\paragraph{Variables for the population dynamics}
%
To convert individual dynamics into biomass dynamics, we just multiplied the variable density $X$ in [ind ha$^{-1}$] with a constant $B_X$ in [g ind$^{-1}$], so that the derivative as just to be multiplied by this constant: 
%
\begin{equation}
\frac{d B_{T,X}}{dt} = \frac{d B_X X}{dt} = B_X \frac{d X}{dt} \quad \text{and so we have} \quad  \frac{d X}{dt} = \frac{1}{B_X} \times \frac{d B_{T,X}}{dt}
\label{eq:biomTOpop}
\end{equation}
%

\paragraph{Variable for the toxicant dynamics}
%
For the dynamic of toxicants, we consider an homogeneous population.

We set a new variable $w_X = C_X B_X X$ which is in [$\mu$g g$^{-1}$] $\times$ [g ind$^{-1}$] $\times$ [ind ha$^{-1}$], so that:
%
$w_X$ is in [mg ha$^{-1}$] or [mg population$^{-1}$], which is the quantity of contaminant within the population.

The dynamic of the new variable is defined with:
%
\begin{equation}
w_X = C_X B_X X = C_X B_{T,X} \quad \Rightarrow \quad \frac{dw_X}{dt} = B_X \frac{dX C_X}{dt} = B_X \left( X \frac{dC_X}{dt} + C_X \frac{dX}{dt} \right)
\end{equation}
%

As a consequence, $\dfrac{dw_X}{dt}$ is in [mg ha$^{-1}$ day$^{-1}$].

\subsubsection{Dynamic of the contaminant in the environment}

In the agro-ecotoxico-logical system, we consider an environmental concentration of contaminant denoted by $w_{ext}$ in [mg ha$^{-1}$] in order to reflect the spatially explicit spreading of AR following determined in baits per area.
%
From the Figure~\ref{fig:scheme}, we can directly provide the dynamic of this concentration:

\begin{equation}
\frac{dw_{ext}}{dt} = T_{Broma}(V) - k_0w_{ext} - \kappa(w_{ext})V B_V
\end{equation}

with $T_{Broma}(V)$ the farmer input of AR in [mg ha$^{-1}$ day$^{-1}$] depending on vole density denoted $V$ in [ind ha$^{-1}$].
%
AR concentration in baits is 50 $\mu$g g$^{-1}$ (or mg kg$^{-1}$). So when farmer spread in grasslands at quantity 7.5 to 20 kg ha$^{-1}$ day$^{-1}$, its an amount of 375 to 1000 mg ha$^{-1}$ of AR spread (i.e., mg kg$^{-1}$ $\times$ kg ha$^{-1}$ day$^{-1}$ ).

The disappearance of AR in the field, denoted $k_0$, in [day$^{-1}$], so that $k_0 w_{ext}$ is in [mg ha$^{-1}$ day$^{-1}$].
%
Such quantity, $w_{ext}$, was available in the literature, and a proportion disappeared in the environment at rate $k_0 = 0.0815$ day$^{-1}$ \citep{Sage2008}.

The function $\kappa(w_{ext})$, in [mg kg$^{-1}$ day$^{-1}$], reflects the transfer of contaminant from the environmental compartment to the vole compartment. As a consequence, the term $\kappa(w_{ext})V B_V$ is in [mg ha$^{-1}$ day$^{-1}$] ([mg kg$^{-1}$ day$^{-1}$] $\times$ [kg ha$^{-1}$]).

This rate $\kappa(w_{ext})$ was assumed to be an increasing function characterized by a maximum intake rate $M_{in}$ in [mg kg$^{-1}$ day$^{-1}$], and a half-saturation constant for ingestion $D_{in}$ in [mg ha$^{-1}$] since $w_{ext}$ is in [mg ha$^{-1}$]:

\begin{equation}
\kappa(w_{ext})= M_{in}\times \dfrac{ w_{ext}}{D_{in} + w_{ext}}
\end{equation}

So the dimension of the ingestion rate $\kappa(w_{ext})$ is the same as $M_{in}$ that is [mg kg$^{-1}$ day$^{-1}$].

\subsubsection{Dynamics of contaminant within populations}

With the previous assumption, the general model is described as follow:

\paragraph{Population dynamics}

From the equation \eqref{eq:biomTOpop}, we have: $\dfrac{d B_{T,X}}{dt} = B_X \dfrac{d X}{dt}$

\paragraph{Conversion of $w_V$ into $C_V$}

\begin{equation}
\begin{array}{c l}
\dfrac{d C_V}{dt} & = \dfrac{d w_V / B_{T,V}}{dt} = \dfrac{1}{B_{T,V}^2} \left( \dfrac{d w_V}{dt} B_{T,V} - w_V \dfrac{d B_{T,V}}{dt}\right) \\[3mm]
%
& \\[3mm]
\end{array}
\end{equation}

\begin{equation}
\left\lbrace
\begin{array}{c l}
% OUTSIDE toxic
\dfrac{dw_{ext}}{dt} & = T_{Broma}(V) - k_0w_{ext} - \kappa(w_{ext})B_{T,V} \\[3mm]
% VOLE Dynamic
\dfrac{dB_{T,V}}{dt} & =
B_{T,V} g_V(V, C_V)	- \mu_V(C_V)B_{T,V} - B_V \Phi_{V,M}(V) M - B_V \Phi_{V,F}(V,M) F\\[3mm]
% VOLE toxic
\dfrac{dw_{V}}{dt} & = \kappa(w_{ext}) B_{T,V} - k_{out, V} w_V - w_V \mu(C_V) - C_V B_V \Phi_{V,M}(V)M + \gamma_V w_V g_V(V, C_V)\\[3mm]
%
% DEAD VOLE Dynamic
\dfrac{dB_{T,V_d}}{dt} & = \mu_V(C_V)B_{T,V}\\[3mm]
% DEAD VOLE toxic
\dfrac{dw_{V_d}}{dt} & = \dfrac{dw_{V}}{dt}\\[3mm]
%
% MUSTELID dynamic
\dfrac{dB_{T,M}}{dt} & =
B_{T,M} (\varepsilon_MB_V  \Phi_{V,M}(V)M - m_M - \mu_M(C_M) ) - B_M \Phi_{M,F}(V, M) F\\[3mm]
% MUSTELID toxic
\dfrac{dw_{M}}{dt} & = - k_{out, M} w_M - w_M (m_M + \mu(C_M)) + \eta_M C_M B_M \Phi_{V,M}(V)M - \\[3mm]
& C_M B_M \Phi_{M,F}(V,M)F + \gamma_M w_M g_M(M, C_V)\\[3mm]
%
% FOX dynamic
\dfrac{dB_{T,F}}{dt} & =
B_{T,F} g_F(F, C_F)	- \mu_F(C_F) \\[3mm]
% FOX toxic
\dfrac{dw_{F}}{dt} & = - k_{out, F} w_F - w_F \mu(C_F) + \eta_F C_M B_M \Phi_{M,V} F  + \eta_F C_V B_V \Phi_{V,V} F + \gamma_F w_F g_F(F, C_F)\\[3mm]
\end{array}
\right.
\end{equation}

%

Thus, it is straightforward to express the dynamics as:

\begin{equation}
\left\lbrace
\begin{array}{c l}
% VOLE Dynamic
\dfrac{dV}{dt} & =
\overbrace{	r_V V (1-V/K_V)		}^{\text{growth rate}} -
\overbrace{	\mu_V(C_V)V	}^{\text{poisoning}} -
\overbrace{	\Phi_{V,M}(V,V_d)M - \Phi_{V,F}(V,V_d,M)F	}^{\text{predation}}\\[3mm]
%
% DEAD VOLE Dynamic
\dfrac{dV_d}{dt} & =
\overbrace{ \mu_V(C_V)V }^{\text{new dead voles}} -
\overbrace{	\Phi_{V_d,M}(V,V_d)M - \Phi_{V_d,F}(V,V_d,M)F	}^{\text{predation}}-
\overbrace{d V_d}^{\text{degradation}}
\\[3mm]
%
% MUSTELIDS Dynamic
\dfrac{dM}{dt} & =
\overbrace{	(\varepsilon_M\Theta_M(V,V_d)-m_M) M	}^{\text{growth rate}} - 
\overbrace{\mu_M(C_M)M	}^{\text{poisoning}} -
\overbrace{\Phi_{M,F}(V,V_d,M)F }^{\text{predation}} \\[3mm]
%
% F0X Dynamic
\dfrac{dF}{dt} & =
\overbrace{	r_F F (1-F/K_F)	}^{\text{growth rate}} -
\overbrace{\mu_F(C_F)F	}^{\text{poisoning}}  \\
%
%
% Toxicological dynamics
\dfrac{dC}{d t}  & = 
\overbrace{T_{Broma}(V)}^{\text{treatment}} - 
\overbrace{k_0 C}^{\text{degradation}} -
\overbrace{\kappa(C)}^{\text{consumption}} \\[3mm]
%
% Toxico-kinetics in Voles
\dfrac{dC_V}{dt} &= 
\overbrace{\kappa(C)}^{\text{intake}} - 
\overbrace{C_V\left(k_{out,V} + r_V \left(1 - \dfrac{V}{K_V}\right)\right)}^{\text{excretion}}\\[3mm]
%
% Toxico-kinetics in Mustelids
\dfrac{dC_M}{d t} & =
\overbrace{\eta_M \dfrac{B_V}{B_M} (\Phi_{V,M}(V,V_d,M)+\Phi_{V_d,M}(V,V_d,M)) C_V}^{\text{intake}} -
\overbrace{C_M\left(k_{out,M} + \varepsilon_M \Phi_{V,M}(V,V_d,M) \right)}^{\text{excretion}} \\[3mm]
%
% Toxico-kinetics in Foxes
\dfrac{dC_F}{d t} &=
\overbrace{ \eta_F \dfrac{B_V}{B_F} \left( (\Phi_{V,M}(V,V_d,M)+\Phi_{V_d,M}(V,V_d,M)) C_V + \Phi_{M,F}(V,V_d,M) C_M \right)}^{\text{intake}} - \\[3mm]
&
\overbrace{k_{out,F} + r_F \left(1 - \dfrac{F}{K_F}\right) }^{\text{excretion}} 
\end{array}
\right.
\label{eq:Sys_complet}
\end{equation}

% BIBLIOGRAPHY
% \section*{References}

%\bibliographystyle{elsarticle-harv}
\bibliographystyle{plainnat}
%\bibliographystyle{unsrt}
\bibliography{trophToxNTO.bib}

\end{document}

### 2019-03-22 trophToxNTO/R/sensitivity.R

@@ -1,9 +1,9 @@
 #' @title Table of sensitivity for set, min, max values
 #'
-#' @descritpiont Take the output of the ode integration with deSolve and return a data.frame
+#' @description Take the output of the ode integration with deSolve and return a data.frame
 #' with a set of measure on
 #' @param model_output the output of the ode solver
-#' @param params Parameter of the system
+#' @param parameters Parameter of the system
 #' @param key specific name of the parameters to consider
 #'
 #' @return A list of data.frame
@@ -13,44 +13,44 @@
 #' @export
 #'
 #'
-tab_sensitivity <- function(model_output, params, tab_ref){
+tab_sensitivity <- function(model_output, parameters){
 
   tab_set <- table_df(model_output)
-  df_params <- as.data.frame(params)
-  df_params$Mean_V_set <- mean(tab_set$D.V)
+  df_parameters <- as.data.frame(parameters)
+  df_parameters$Mean_V <- mean(tab_set$D.V, na.rm = T)
 
-  df_params$Mean_M_set <- mean(tab_set$D.M)
+  df_parameters$Mean_M <- mean(tab_set$D.M, na.rm = T)
 
-  df_params$Mean_F_set <- mean(tab_set$D.F)
+  df_parameters$Mean_F <- mean(tab_set$D.F, na.rm = T)
 
-  df_params$V_inf50_set <- nrow(tab_set[tab_set$D.V <50,]) / nrow(tab_set)
+  df_parameters$V_inf50 <- nrow(tab_set[tab_set$D.V <50,]) / nrow(tab_set)
 
-  df_params$nrb_AR_set <- nbr_spike(tab_set$C.soil)
+  df_parameters$nrb_AR <- nbr_spike(tab_set$C.soil)
 
-  tab_mortality_set <- Nbr.mortality(tab_set,params,group.num=1)
+  tab_mortality <- Nbr.mortality(tab_set,parameters,group.num=1)
 
-  df_params$M_50AR_set <- sum(tab_mortality_set$Percent.M.mort.B > 0.5) / nrow(tab_mortality_set)
+  df_parameters$M_50AR <- sum(tab_mortality$Percent.M.mort.B > 0.5) / nrow(tab_mortality)
 
[…]

### 2019-03-22 trophToxNTO/vignettes/getting_started.R

@@ -1,0 +1,76 @@
+## ---- eval=FALSE---------------------------------------------------------
+#  install.packages('devtools') # if not already installed
+#  library('devtools') # load the library
+#  github_install("vbaudrot/trophToxNTO/trophToxNTO") # install from 'master'
+
+## ----loadLibrary---------------------------------------------------------
+library(trophToxNTO)
+
+## ----warning=FALSE-------------------------------------------------------
+vole.level.test = 250
+broma.level.test = 20*50
+
+# Run condition function for scenario
+scenario = condition(Vole.Level=vole.level.test,Broma.Level=broma.level.test)
+
+root.event=scenario[[1]]
+event.func=scenario[[2]]
+
+## ----warning=FALSE-------------------------------------------------------
+# Load package deSolve
+# library(deSolve)
+
+# number of parallel simulation
+params=pars(n=1)
+
+# times vector of integration
+times <- seq(from=0, to=15*365, length.out = 1e3) #  30*5 = 5 months 
+
+# Run the model
+
+model.output    <- ode(times = times,
+                       y = init(params$n),
+                       func = ode.broma.tritrophic,
+                       parms  = params,
+                       events = list(func = event.func, root = TRUE),
+                       rootfun = root.event)
+
+
+## ----warning=FALSE,message=FALSE-----------------------------------------
+graphique(params$n,times,model.output) 
+
+## ----warning=FALSE-------------------------------------------------------
+# number of parallel simulation
+params=pars(n=2)
+
+# Run the model
+model.output    <- ode(times = times,
+                       y = init(params$n),
+                       func = ode.broma.tritrophic,
+                       parms = pars(
+                         n=params$n, # Do not foret to specify this !!!
+                         K.V=seq(200,800,length=params$n)
+                       ),
+                       events = list(func = event.func, root = TRUE),
+                       rootfun = root.event)
+
+# Graphic output
+graphique(params$n,times,model.output) 
+
+## ----warning=FALSE-------------------------------------------------------
+# number of parallel simulation
+params=pars(n=2)
+
+# Run the model
+model.output    <- ode(times = times,
+                       y = init(params$n,
+                                D.F = c(0,0.2)
+                                ),
+                       func = ode.broma.tritrophic,
+                       parms = params,
+                       events = list(func = event.func, root = TRUE),
+                       rootfun = root.event)
+
+# Graphic output
+graphique(params$n,times,model.output) 
+

### 2019-06-03 manuscript/MS_trophToxNTO.tex
update manuscript

@@
 
-\paragraph{Running Title} \textcolor{gray}{TO CHANGE: Bufferscapes to preserve Non-Target Organism / Ecosystem services}
+\paragraph{Running Title} \textcolor{gray}{Bufferscapes to preserve Non-Target Organism / Ecosystem services}
 
@@
 By considering anticoagulant rodenticide toxicokinetic and toxicodynamic equations, we explored the impact of 5 farmer functional responses (defined by both rodenticide quantity and threshold vole density above which rodenticide spreading is prohibited) on predator-prey interactions, rodenticide transfer across the trophic chain and population effects.
-		\item Spreading low quantities of rodenticide while maintaining sufficient voles as prey resources led to less rodenticide being applied and extended periods without rodenticide in the environment, benefitting predators while avoiding episodes with high vole density.
+		\item Spreading rodenticide while maintaining sufficient voles as prey resources led to less rodenticide being applied and extended periods without rodenticide in the environment, benefitting predators while avoiding episodes with high vole density.
 This may meet farm production interests while minimizing the impact on small mustelid and fox populations.
-		Spreading rodenticide at low vole densities suppressed mustelid and fox populations, leading to vole population dynamics being entirely regulated by rodenticide use.
+		Spreading rodenticide to maintain low vole densities suppressed mustelid and fox populations, leading to vole population dynamics being entirely regulated by rodenticide use.
 This vole-eradication treatment regime inhibited predator ecosystem services and promoted pesticide dependence.
@@
 Mustelids could feed on live voles $V$, or non-decomposed dead voles Vd and we assumed a Type 2 functional response adapted for a multi-species functional response \citep{Baudrot2016}.
-Mustelids ingested AR with absorption rate $\eta_M$ (ratio between biomasses of voles, $B_V$, and mustelids, $B_M$) and the total of ingested voles (alive $V$ and dead $V_d$) was defined by function $\Theta_M (V,V_d=$ (Table~\ref{tab:parameterization}), in equation (4) which provides the rate of AR ingested by a mustelid:
+Mustelids ingested AR with absorption rate $\eta_M$ (ratio between biomasses of voles, $B_V$, and mustelids, $B_M$) and the total of ingested voles (alive $V$ and dead $V_d$) was defined by function $\Theta_M (V,V_d=$ (Table~\ref{tab:parameterization}).
+
+A fraction of AR ingested was accumulated in weasels while the rest was excreted with rate $k_{out,M}$.
+%
+AR provided to weasel though vole poisoning, is denoted $D_M$ induces lethal effect at rate $\mu_M(D_M)$, additive to natural mortality rate $m_M$.
+Parameter definition are further detailed in Supporting Information.
+%
+AR was ingested by foxes with a rate proportional to the functional response of foxes to voles, dead voles and mustelids.
+%
+Foxes also accumulated AR available in their prey, resulting in upward AR transfer in the trophic chain.
+Foxes accumulated AR in concentration $C_F$.
+A fraction of AR was excreted by foxes at rate $k_{out,F}$ at a rate between 0.38 and 0.72 [day$^{-1}$]\citep{Sage2010}, and AR caused fox mortality at rate $\mu_F$ ($D_F$).
+
+We used log-logistic equations for describing dose-dependent mortality of animals exposed to AR.
+Vole and predator mortality rates due to AR $\mu_X(\zeta_X)$ ($X$ referring to the considered species, and $\zeta_X$ being the dose eaten $D_X$ or the internal concentration $D_X$) were expressed by equation (5):
 
 \begin{equation}
-\text{Contaminant Intake Rate} = C_V \times \eta_M \times \Theta_M(V,V_d)
-\end{equation}
-
-A fraction of AR ingested was accumulated in weasels while the rest was excreted with rate $k_{out,M}$.
-AR ingestion induced weasel lethal poisoning at rate $\mu_M(C_M)$, additive to natural mortality rate $m_M$.
-AR was ingested by foxes with a rate proportional to the functional response of foxes to voles, dead voles and mustelids.
-Foxes also accumulated AR available in their prey, resulting in upward AR transfer in the trophic chain.
-Foxes accumulated AR in concentration $C_F$.
-A fraction of AR was excreted by foxes at rate $k_{out,F}$ \citep{Sage2008}.
-AR caused fox mortality at rate $\mu_F$ ($C_F$) \citep{Sage2010}.
-We used log-logistic equations for describing dose-dependent mortality of animals exposed to AR.
-Vole and predator mortality rates due to AR $\mu_X(C_X)$ ($X$ referring to the considered species) were expressed by equation (5):
-
-\begin{equation}
-\mu_X(C_X) = \dfrac{1}{\text{experiment duration}} \times \left( 1- \dfrac{1}{1+ (LD_{50}/C_X)^H} \right) 
+\mu_X(\zeta_X) = \dfrac{1}{\text{experiment duration}} \times \left( 1- \dfrac{1}{1+ (LD_{50}/\zeta_X)^H} \right) 
 \label{eq:muX}
@@
 
-where $LD_{50}$ was the daily median lethal dose (50\% of population dying) with different values for voles (2 mg kg$^{-1}$), mustelids (2.1 mg kg$^{-1}$) and foxes (0.5 mg kg$^{-1}$, Table~\ref{tab:parameterization}).
-$H$ is the Hill's coefficient, modulating the curve steepness and was estimated to fit the sparse data we have (Table~\ref{tab:parameterization}).
+where $LD_{50}$ was the daily median lethal dose (50\% of population dying) with a value for voles of 2 mg kg$^{-1}$ as internal concentration, and a dose given as food of  2.1 mg kg$^{-1}$ mustelids \citep{Grolleau1989} and 0.5 mg kg$^{-1}$ for foxes \citep{Sage2010} (see Table~\ref{tab:parameterization}).
+Parameter $H$ is the Hill's coefficient, modulating the curve steepness and was estimated to fit the sparse data we have (Table~\ref{tab:parameterization}).
 Mortality rates considered the duration of AR toxicity from experiments of up to 6 days \citep{Sage2010}.
@@
 %
-A first approach consisted in changing parameter within a 10\% range around the original value \textit{one at a time}, and a second global approach consisted in using the first-order Sobol's sensitivity index $S_i$ defined as $S_i = \mathbb{V}(\mathbb{E}_{-i}(y \vert x_i)) / \mathbb{V}(y)$, where $\mathbb{V}(y)$ is the variance of $y$ when all factors are allowed to vary, $\mathbb{E}_{x\sim i}(y \vert x_i))$ the mean of $y$ when one factor is fixed \citep{Sobol1993}.
+We applied a global approach using the first-order Sobol's sensitivity index $S_i$ defined as $S_i = \mathbb{V}(\mathbb{E}_{-i}(y \vert x_i)) / \mathbb{V}(y)$, where $\mathbb{V}(y)$ is the variance of $y$ when all factors are allowed to vary, $\mathbb{E}_{x\sim i}(y \vert x_i))$ the mean of $y$ when one factor is fixed \citep{Sobol1993}.
 We defined the domain of variation with a beta distribution within an hyperspace of plus or minus 10\% from the original value for all parameters.
@@
 Secondary poisoning led to a rich spectrum of emergent dynamics according to the FFR to vole abundance.
-Without AR (FFR a), vole dynamics were regulated by mustelid predation that gave rise to 6-year cycles (Figs.
-2 and 3; maximal vole and mustelid densities were around 600 ind ha$^{-1}$ and 2 ind ha$^{-1}$ respectively).
+Without AR (FFR a), vole dynamics were regulated by mustelid predation that gave rise to 6-year cycles (Figure~\ref{fig:populations}).
 The fox population remained at its carrying capacity (i.e., 0.03 ind ha$^{-1}$, see Table~\ref{tab:parameterization}).
 
-Under FFR b (high vole density threshold triggering treatment and high AR amount per treatment) and FFR d (intermediate threshold and low AR amount), vole dynamics were sequentially regulated by either AR treatments, which we refer to as farmer-regulated phase (FR in Figures 2 \& 3), or by mustelids, mustelids-regulated phase (MR), alternately (MR in Figures 2 \& 3).
+Under FFR b (high vole density threshold triggering treatment and high AR amount per treatment) and FFR d (intermediate threshold and low AR amount), vole dynamics were sequentially regulated by either AR treatments, which we refer to as farmer-regulated phase (red color sections in Figure~\ref{fig:populations}), or by mustelids, mustelids-regulated phase (blue color sections in Figure~\ref{fig:populations}).
 Farmer-regulated periods started when densities of living voles triggered treatments.
@@
 Vole depletion by mustelids and subsequent mustelid declines allowed the vole population to grow again up to threshold densities and initiated a new period of regulation by farmers.
-Vole dynamics were permanently regulated by AR treatment under FFR c (intermediate threshold, high AR, Figs.
-2c, 3c).
+%
+Vole dynamics were permanently regulated by AR treatment under FFR c (intermediate threshold, high AR, Figure~\ref{fig:populations}-c).
 Populations of live and dead voles experienced high frequency fluctuations (around 2 peaks every year) driven by AR.
-As AR treatments were frequent, being triggered by voles peaks, contaminated dead voles were always abundant (peaks at 90 ind ha$^{-1}$), and mustelid and fox densities remained low (mustelids, converging to 0 ind ha$^{-1}$; foxes, 0.003-0.006 ind ha$^{-1}$) with only short-term fluctuations (2 peaks every year, following vole cycles).
-With FFR e (low threshold, low AR), vole populations were maintained by farmers at around 50 voles ha$^{-1}$ (Figs.
-2e and 3e).
+As AR treatments were frequent, being triggered by voles peaks, contaminated dead voles were always abundant (peaks at 90 ind ha$^{-1}$), and mustelid and fox densities remained low (see Supporting Information: mustelids <0.150 ind ha$^{-1}$ and foxes <0.01 ind ha$^{-1}$) but with a slowly increasing trend on the long term period (Figure~\ref{fig:populations}-c).
+This pattern of an increasing trend appears to be very sensitive to parameterization (i.e.
+a slight change in parameter may lead to a long-term decreasing pattern).
+%
+With FFR e (low threshold, low AR), vole populations were maintained by farmers at around 50 voles ha$^{-1}$ (See Figure~\ref{fig:populations} and Supporting Information).
 The population of live voles was buffered by treatments.
 Additionally, whenever voles reached densities triggering treatment, predator populations experienced strong declines.
-However, fox densities (0.01-0.015 ind ha$^{-1}$) were higher compared to FFR c (intermediate threshold, high AR), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities.
+However, fox densities were higher compared to FFR c (intermediate threshold, high AR) (>0.01 ind ha$^{-1}$, see Supporting Information), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities.
 The maximum numbers of dead voles under this FFR e was relatively low (highest around 15 ind ha$^{-1}$) but, due to frequent treatments, there was a steady replenishment of contaminated dead voles.
@@
 As expected, spreading AR generally reduced vole densities (Table~\ref{tab:results}) but the extent to which it also affected predators and the costs and benefits for conservation and farmer’s interests varied widely.
-The FFR b (high threshold, high AR) implied the lowest number of treatments (mean 0.87 year$^{-1}$), while FFR d (intermediate threshold, low AR) resulted in the lowest amount of AR used (mean 14.5 kg ha$^{-1}$ year$^{-1}$).
+The FFR b (high threshold, high AR) and d (intermediate threshold, low AR) implied the lowest number of treatments (0.5 and 1.8 year$^{-1}$ respectively) and the lowest amount of AR used (mean 10 and 13.5 kg ha$^{-1}$ year$^{-1}$ resp.).
 The FFR e (low threshold, low AR) had the lowest mustelid mortality due to AR (no instance in 40 years) but maximised starvation-induced mortality of mustelids (Table~\ref{tab:results}).
-Additionally, FFR e (low threshold, low AR) had the highest proportion of time with < 50 voles' ind ha$^{-1}$, (96\% of study period).
-When considering the effectiveness in reducing mean vole densities, FFR e (low threshold, low AR) ranked highest followed by FFR c (intermediate threshold, high AR).
-The FFR d (intermediate threshold, low AR) and b (high threshold, high AR) allowed the highest mean densities of mustelids and foxes respectively.
-Comparison of these protocols indicates that FFR d (intermediate threshold and low AR) delivered the best compromise between farm production and conservation interests (Table~\ref{tab:results}).
-It had the lowest AR quantity applied (mean 14.5 kg ha$^{-1}$.year$^{-1}$) and the highest mean densities of mustelids (0.47 ha$^{-1}$) and foxes (0.02 ha$^{-1}$).
-The number of treatments was among the lowest values (1.6 treatment year$^{-1}$).
-Its drawback was that it only suppressed voles at low densities (<50 voles ha$^{-1}$ for 21\% of time).
-Conversely, emergent mean vole densities were low to intermediate (122 voles ha$^{-1}$).
-Other FFR regimes were poorer compromises either because mustelids were kept at low densities (FFR  c, intermediate threshold, high AR, and FFR e, low threshold, low AR) or had relatively high mean vole densities (229 voles ha$^{-1}$, FFR b, high threshold, high AR).
-In summary, FFR d (intermediate threshold, low AR) was the best compromise because it minimized the impact over predator populations contributing to the natural reduction of vole densities with the lowest quantity of AR applied for farm production interests.
+Additionally, FFR e (low threshold, low AR) had the highest proportion of time with < 50 voles' ind ha$^{-1}$, (91\% of study period).
+When considering the effectiveness in reducing mean vole densities over the long-term period, FFR e (low threshold, low AR) ranked highest (only 55 ind ha$^{-1}$ in means, and >118 ind ha$^{-1}$ for other scenarios).
+The FFR b (high threshold, high AR) and d (intermediate threshold, low AR) allowed the highest mean densities of mustelids and foxes.
+%
+Comparison of these protocols indicates that FFR b and d, inducing successive farmer-regulated and predator-regulated phases, were the best scenarios for conservation interests (Table~\ref{tab:results}).
 
@@
 
-Figure~\ref{fig:sobol_sensitivity} and Figure~\ref{fig:sensitivity} show the sensitivity of each parameter used in the model defined in d (i.e., 7.5kg of AR at threshold density of 250 voles).
 %
-Figure~\ref{fig:sobol_sensitivity} represent the global sensitivity where we computed a first order sensitivity index \citep{Sobol1993, Saltelli2019} given the contribution of each parameter to the variance of the model.
+Figure~\ref{fig:sobol_sensitivity} represents the global sensitivity where we computed a first order sensitivity index \citep{Sobol1993, Saltelli2019} given the contribution of each parameter to the variance of the model d (i.e., 7.5kg of AR at threshold density of 250 voles).
 %
-Figure~\ref{fig:sensitivity} is a one at time analysis \citep{Saltelli2019} where parameters are move +/- 10\% of the fixed value as defined in scenario d (see Table~\ref{tab:parameterization}).
-
+Parameters are moved within +/- 10\% interval around the fixed value as defined in scenario d (see Table~\ref{tab:parameterization}).
+%
 Crossing all cost functions to test the sensitivity of model parameters, the influence of each parameter appears quite homogeneous.
-That is there is no useless parameter or in return one that would lead all.
 %
-Taken cost functions separately, the sensitivity to parameters present some differences.
+There is no redundant or conversely any parameters with a dominant influence on model behaviour.
 %
-Figure~\ref{fig:sobol_sensitivity} shows an homogeneity of parameter sensitivity on the mean of mustelids density, revealing the complexity underlying this latent variable.
+Taking cost functions separately, the sensitivity to parameters present some differences.
+%
+This is particularly true for densities of predators on which we focus our attention.  
+For mustelids and foxes variable.
 
@@
 In the ecosystem our models depict, farmers spreading rodenticide not only depleted vole prey exploited by specialist and generalist predators but also created pulses of lethally or sub-lethally poisoned voles that became super-predators by poisoning their predators.
-Arguably this set of ecological interactions has similarities with circumstances where a pathogen affecting prey species also infects predators, as in the case with the flea vectored plague (\textit{Yersinia pestis}) infecting prairie dogs (\textit{Cynomys} spp.) and black footed ferrets (Mustela nigripes) in central US \citep{Matchett2010}.
+Arguably this set of ecological interactions has similarities with circumstances where a pathogen affecting prey species also infects predators, as in the case with the flea vectored plague (\textit{Yersinia pestis}) infecting prairie dogs (\textit{Cynomys} spp.) and black footed ferrets (\textit{Mustela nigripes}) in central US \citep{Matchett2010}.
 However, to our knowledge, the behaviour of such tri-trophic model with multiple reciprocal interactions has not been explored.
@@
 However, given we only explored deterministic versions of our models, any inference on the frequency of flipping between states should be cautious given the inherent stochastic nature of natural and farmland environments.
-If such dynamics occur within real farming systems, flipping between states is unlikely to emerge with regularity (~40 years) where many other factors impact population dynamics.
-Moreover, under our heuristic modelling framework, we have shown that the removal of fox population induced chaotic dynamics of voles and mustelids (Figure~\ref{fig:intraguild}).
+If such dynamics occur within real farming systems, flipping between states is unlikely to emerge with regularity where many other factors impact population dynamics.
 While generalists are known to have stabilising effect \citep{Hanski1991}, the benefit of specialist predators imparted to generalist predator and resulting increase in the prevalence of intraguild predation would be difficult to detect in empirical studies.
@@
 Our model only considers temporal refuges, and the conceptualization of untreated areas as equivalent to triggering treatment at intermediate vole density cannot provide guidance on the size of these spatial refuges.
-Nevertheless, while management of voles is implemented at the scale of fields, mustelids and foxes roams over much larger areas  \citep{King2006}, such that large refuges with medium-high vole densities voles would be required while allowing to maintain low-voles density at very local scale.
+Nevertheless, while management of voles is implemented at the scale of fields, mustelids and foxes roams over much larger areas  \citep{King2006}, such that large refuges with medium-high vole densities voles would be required while maintaining low vole density at local scale.
 
@@
 Our process-based model revealed pesticides that permeate the food chain upward can lead to diverse population dynamics with alternative states regulated by predators and farmers.
-It also shows that the practice currently promoted to use low-dose AR treatments at low vole density could be limited by the undesirable side-effects of leading to chronic application of AR on a large scale and the depletion of the vole predator community.
+It also shows that the practice currently promoted to use low-dose AR treatments at low vole density could have limited by the undesirable side-effects of leading to chronic application of AR on a large scale, in the absence of refuges, and the depletion of the vole predator community.
 This emerging question would benefit from a landscape modelling approach to characterize spatial refuges.
@@
 
-%\bibliographystyle{elsarticle-harv}
-\bibliographystyle{plainnat}
+\bibliographystyle{elsarticle-harv}
+%\bibliographystyle{plainnat}
 %\bibliographystyle{unsrt}
@@
 Column \textit{"FFR"} (farmer functional response) refers to the protocol denotation (a to e) and in brackets respectively: the vole density threshold, and the amount of AR/treatment in kg.
-The column \textit{"Nbr Treat."} refers to the number of treatment in simulation over the 40 years after the burn-in phase.
-\textit{"Total AR"} is the total use of anticoagulant rodenticides (bromadiolone in kg).
+The column \textit{"Nbr treat/yr"} refers to the number of treatment in simulation per year after the burn-in phase.
+\textit{"Total AR/yr"} is the total use of anticoagulant rodenticides per year (bromadiolone in kg).
 \textit{"M, 50\% AR death"} is the proportion of time with mortality by AR over 50\% in mustelids population.
@@
 	\label{tab:results}
-\begin{tabular}{p{2.1cm} c c c c c c c c}
+	\begin{tabular}{p{2cm} c c c c c c c c}
 		\hline
-\tabcell{FFR \\ (dose, thresh.)} &
-\tabcell{Nbr. \\ Treat.} &
-\tabcell{Total  \\ AR} &
-\tabcell{M, 50\% \\ AR death} &
-\tabcell{M, 50\% \\ Nat. death} &
-\tabcell{V < \\ 50 ha$^{-1}$} &
-\tabcell{Mean V \\ density} & 
-\tabcell{Mean M \\ density} & 
-\tabcell{Mean F \\ density} \\
-\hline
+		\tabcell{FFR \\ (dose, limit)} & \tabcell{Nbr.
+\\ treat/yr} & \tabcell{Total  \\ AR/yr} & \tabcell{M, 50\% \\ AR death} & \tabcell{M, 50\% \\ Nat.
+death} & \tabcell{V < \\ 50 ha$^{-1}$} & \tabcell{Mean V \\ density} & \tabcell{Mean M \\ density} & \tabcell{Mean F \\ density} \\ \hline
 		%%
-a (0,0) & 0& 0& 0& 1& 0.48 & 211.27 & 0.61 &0.03\\
+		a (0,0)                   & 0                          & 0                         & 0                             & 1                               & 0.46                          & 220                         & 0.62                        & 0.03                        \\
 		% 
-b (500,20) & 35& 700& 0.06 &0.94 & 0.14 & 229.1 & 0.33 & 0.02\\
+		b (500,20)                & 0.5                        & 10\texttt{}                       & 0.02                          & 0.98                            & 0.39                          & 198                         & 0.52                        & 0.02                        \\
 		%
-c (250, 20) & 81 & 1620 &0.11 & 0.89 & 0.28 & 105.49 & <0.01 & <0.01\\
+		c (250, 20)                & 2.8                        & 56                        & 0.09                          & 0.91                            & 0.15                          & 118                         & 0.08                        & <0.01                       \\
 		%
-d (250, 7.5)& 64 & 480 & 0.06 & 0.94 & 0.21 & 122.35 & 0.47 & 0.02\\
+		d (250, 7.5)               & 1.8                        & 13.5                        & 0.04                          & 0.96                            & 0.29                          & 126                         & 0.55                        & 0.02                        \\
 		%
-e (50,7.5) & 169 & 1267.5 & 0& 1 & 0.96 & 31.42 & <0.01 & 0.01 \\
-\hline
+		e (50,7.5)                 & 5.1                        & 38                        & 0                             & 1                               & 0.91                          & 55                          & 0.07                        & 0.02                        \\ \hline
 	\end{tabular}
@@
 		\includegraphics[width=.8\linewidth]{img/system_scheme.png}
-		\caption{Dynamics of mustelids (bold grey line) and red foxes (bold black line) after 50 years of simulation (10 years of burn-in period, shaded area, and 40 years considered here for the compute of results, estimation of
-			cost functions, etc.).
-Letters at the top-right part of each sub-graph corresponds to the results obtained from each farmer functional response.
-FR (dashed line) and MR (dotted line) indicate the farmer-regulated and mustelid-regulated periods of vole population dynamics.}
+		\caption{Flow diagram of the fate and impacts of anticoagulant rodenticide (AR, bromadiolone) in a tri-trophic food web.
+Black arrows and equations correspond to natural dynamics with trophic interactions, red arrows and equations represent the transfer of AR through the system and its accumulation into the different compartments.
+The arrows and equations in blue correspond to its impact (i.e., death of individuals) on the three species populations.}
 		\label{fig:scheme}
@@
 Letters at the top-center part of each sub-graph corresponds to the results obtained from each farmer functional response.
-Red area and Green area indicate the farmer-regulated (FR) and mustelid-regulated (MR) periods of vole population dynamics respectively.}
+Colored areas indicate the farmer-regulated (red area) and mustelid-regulated (blue area) periods of vole population dynamics respectively.}
 		\label{fig:populations}
@@
 	\begin{center}
-		\includegraphics[width=\linewidth]{img/sobol_heatMap.png}
+		\includegraphics[width=\linewidth]{img/plt_sobolIndex_draw.jpg}
 		\caption{First-order Sobol's sensitivity index (denoted $S_p$ for parameter $p$) providing the global sensitivity of cost functions to parameters of the model \citep{Sobol1993} for the model d (i.e., 7.5kg of AR at threshold density of 250 voles).
@@
 
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\linewidth]{img/Sensitivity_pm10_2.png}
-		\caption{Minimal and maximal values of cost functions with variation of + and - 10\% of change for each parameter value with other parameter values fixed, as set in Table~\ref{tab:parameterization}.
-Definition of parameters are in Table~\ref{tab:parameterization} (dd$_X$ are experiment duration as defined by equation \ref{eq:muX}), and cost function are defined in Table~\ref{tab:results}.
-Column SSE stands for Sum of Square Error and is scaled in $[0,1]$.}
-		\label{fig:sensitivity}
-	\end{center}
-\end{figure}
-
 \end{document}

### 2019-06-03 manuscript/SupMat_trophToxNTO.tex
update manuscript

@@
 % MATHS
+% To manage cells in table
+\newcommand{\tabcell}[2][c]{%
+	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
+
 %% The amssymb package provides various useful mathematical symbols
@@
 	\huge\textbf{
-	Supporting Information 1
+	Supporting Information
 	}\par
@@
 		Virgile Baudrot$^{1,2}$ 
-		Javier Fernandez-de-Simon$^1,3$
+		Javier Fernandez-de-Simon$^{1,3}$
 		Michael Coeurdassier$^1$,
-		Geoffroy Couval$^1,4$,
+		Geoffroy Couval$^{1,4}$,
 		Patrick Giraudoux$^1$,
-		Xavier Lambin$^5,6$		
+		Xavier Lambin$^{5,6}$		
 	}\par	
@@
 
-\newpage
+ \clearpage
 
@@
 
-\begin{figure}
+\begin{figure}[!htb]
 	\begin{center}
@@
 %
-\dfrac{dM}{dt} & = \varepsilon_M \Phi_{V,M}(V)M - m_M M - \Phi_{M,F}(V,M)F\\[.3cm]
+\dfrac{dM}{dt} & = \varepsilon_M B_V \Phi_{V,M}(V)M - m_M M - \Phi_{M,F}(V,M)F\\[.3cm]
 %
@@
 As described in the manuscript, the vole population, $V$ followed a logistic growth rate, with $r_V$  the maximal reproduction rate, fixed at $r_V = \ln(2 \times 600)/365$ in [day$^{-1}$], since montane water vole populations can increase from 0  to  600 individuals ha$^{-1}$ or more in a year \citep{Giraudoux1997} resulting in the equilibrium density being fixed at $K_V = 600$ individuals.
+%
+Other variable and parameters are detailed hereafter.
 
@@
 \begin{array}{l l}
-\Phi_{V,M}(V)M = \dfrac{a_M V}{1 + h_M a_M V} \\[.3cm]
+\Phi_{V,M}(V) = \dfrac{a_M V}{1 + h_M a_M V} \\[.3cm]
 %
-\Phi_{V,F} = \dfrac{a_{VF} V}{a_{VF}V + a_{MF}F} \times \dfrac{(a_{VF}V + a_{MF} M)^2}{1 + h_F(a_{VF}V + a_{MF} M)^2}\\[.3cm]
+\Phi_{V,F}(V) = \dfrac{a_{VF} V}{a_{VF}V + a_{MF}F} \times \dfrac{(a_{VF}V + a_{MF} M)^2}{1 + h_F(a_{VF}V + a_{MF} M)^2}\\[.3cm]
 %
-\Phi_{M,F}(V,M)F = \dfrac{a_{MF}V}{a_{VF}V + a_{MF}F} \times \dfrac{(a_{VF}V + a_{MF}M)^2}{1+h_F(a_{VF}V + a_{MF}M)^2}
+\Phi_{M,F}(V,M) = \dfrac{a_{MF}V}{a_{VF}V + a_{MF}F} \times \dfrac{(a_{VF}V + a_{MF}M)^2}{1+h_F(a_{VF}V + a_{MF}M)^2}
 \end{array}
@@
 
-The parameterization of the conversion efficiency of ingested food into new born, $\varepsilon_M$ in [n.d.] (non dimensional), and the mortality rate, $m_M$ in [day$^{-1}$], in equation \eqref{eq:simpleSystem} is not straightforward since we have to take into account the mortality rate due to starvation.
+The parameterization of the conversion efficiency of ingested food into new born, $\varepsilon_M B_V$ (the parameter $B_V$ is the constant mean biomass of one individual as we explain latter, but could be integrated into a general constant parameter since $B_V$ is a constant) in [n.d.] (non dimensional, so with $B_V$ in [kg], in this formulation $\varepsilon_M$ is in [kg$ {-1}$], which is a rate by kg), and the mortality rate, $m_M$ in [day$^{-1}$], in equation \eqref{eq:simpleSystem} is not straightforward since we have to take into account the mortality rate due to starvation.
 % 
@@
 
-Therefore, if we set  the food-independent mortality rate $m_M = m_{M, -\Phi}$ and the conversion efficiency of ingested food $\varepsilon_M=\alpha ((1 - \kappa) \varepsilon' - \sigma  \kappa ) $ , we get back to the equation provided in model \eqref{eq:simpleSystem} but with a clearer definition of $\varepsilon_M$.
+Therefore, if we set  the food-independent mortality rate $m_M = m_{M, -\Phi}$ and the conversion efficiency of ingested food $\varepsilon_M B_V=\alpha ((1 - \kappa) \varepsilon' - \sigma  \kappa ) $ , we get back to the equation provided in model \eqref{eq:simpleSystem} but with a clearer definition of $\varepsilon_M B_V$.
 
-This is why we parameterized $m_M$ in [day$^{-1}$] as the inverse of life expectancy in [day$^{-1}$], and $\varepsilon_M$ in [n.d.] as an unclear parameter derived from the null model (i.e., without AR) to reflect observed population dynamics.
+This is why we parameterized $m_M$ in [day$^{-1}$] as the inverse of life expectancy in [day$^{-1}$], and $\varepsilon_M B_V$ in [n.d.] as an unclear parameter derived from the null model (i.e., without AR) to reflect observed population dynamics.
 
-In any case, the dimension of the equation $\dfrac{dM}{dt}  = \varepsilon_M \Phi_{V,M}(V)M - m_M M - \Phi_{M,F}(V,M)F$ is well defined, since $\varepsilon_M \Phi_{V,M}(V)$ is in [day$^{-1}$].
+In any case, the dimension of the equation
+\begin{equation}
+\dfrac{dM}{dt}  = \varepsilon_M B_V \Phi_{V,M}(V)M - m_M M - \Phi_{M,F}(V,M)F
+\label{eq:dynamic_M}
+\end{equation}
+%
+is well defined, since $\varepsilon_M B_V \Phi_{V,M}(V)$ is in [day$^{-1}$].
 
@@
 %
-The growth function of the prey population is a function $g_X(X, C_X)$ in [day$^{-1}$] depending on the population density $X$ and the concentration of the contaminant $C_X$.
+The growth function of the prey population is a function $g_X(X, \zeta_X)$ in [day$^{-1}$] depending on the population density $X$ and the dose of the contaminant $\zeta_X$, either the internal concentration $C_X$ or the dose eaten $D_X$.
 %
-The dose-response curve $\mu_X(C_X)$ is defined by a log-normal cumulative distribution function as in \cite{Loos2010}.
+The dose-response curve $\mu_X(\zeta_X)$ is defined by a log-normal cumulative distribution function as in \cite{Loos2010}.
 
@@
 %
-
-\paragraph{Variable for the toxicant dynamics}
-%
-For the dynamic of toxicants, we consider an homogeneous population.
-
-We set a new variable $w_X = C_X B_X X$ which is in [$\mu$g g$^{-1}$] $\times$ [g ind$^{-1}$] $\times$ [ind ha$^{-1}$], so that:
-%
-$w_X$ is in [mg ha$^{-1}$] or [mg population$^{-1}$], which is the quantity of contaminant within the population.
-
-The dynamic of the new variable is defined with:
-%
-\begin{equation}
-w_X = C_X B_X X = C_X B_{T,X} \quad \Rightarrow \quad \frac{dw_X}{dt} = B_X \frac{dX C_X}{dt} = B_X \left( X \frac{dC_X}{dt} + C_X \frac{dX}{dt} \right)
-\end{equation}
-%
-
-As a consequence, $\dfrac{dw_X}{dt}$ is in [mg ha$^{-1}$ day$^{-1}$].
 
@@
 
+\subsubsection{The dose in dose-response}
+
+The dose of the dose-response function can be either based on the exposure in the environment, the food ingested or the internal concentration.
+While data used in this paper are considering internal concentration for voles, this is the ingested concentration used for mustelids and predators.
+
+For the dynamic of dose, we consider an homogeneous population, where the dose eaten, or the internal concentration are equal to all individuals within a population. We denote $\zeta_X$ such a dose.
+
+\paragraph{Dynamic of the dose eaten: $\zeta_X$}
+%
+Since predators are only exposed to the contaminant through the functional response, we introduce $D_{V,M}$, $D_{V,F}$ and $D_{M,F}$  which are respectively the doses of contaminant include in food of individuals in [$\mu$g g$^{-1}$ day$^{-1}$ ind$^{-1}$].
+%
+For simplification of writing, we write $D_M = D_{V,M}$ and $D_F = D_{V,F}+D_{M,F}$
+%
+Since $D_X$ it is by kg of the predator $X$, we have to take into account variation of the biomass of this predator already defined by $B_{T,X}$.
+Therefore we first define the total amount of contaminant eaten by mustelids $w_{exp,M}$ in [mg pop$_M^{-1}$ or mg ha$^{-1}$] and derive the dose.
+%
+\begin{equation}
+w_{exp,X} = D_X B_X X = D_X B_{T,X} \quad \Rightarrow \quad \frac{dw_{exp,X}}{dt} = B_X \frac{dX D_X}{dt} = B_X \left( X \frac{dD_X}{dt} + D_X \frac{dX}{dt} \right)
+\end{equation}
+%
+Therefore, $\dfrac{dw_{exp,X}}{dt}$ is in [mg pop$^{-1}$ day$^{-1}$] or, since we consider a close population within one ha, it is in  [mg ha$^{-1}$ day$^{-1}$].
+
+Also:
+%
+\begin{equation}
+\frac{dD_X}{dt} = \frac{dw_{exp,X}/B_{T,X}}{dt} = \dfrac{1}{B^2_{T,X}} \left( B_{T,X} \frac{dw_{exp,X}}{dt} + w_{exp,X} \frac{dB_{T,X}}{dt} \right)
+\end{equation}
+%
+
+\paragraph{Dynamic of the internal concentration}
+%
+As for the dose provided, we set a new variable $w_X = C_X B_X X$ which is in [$\mu$g g$^{-1}$] $\times$ [g ind$^{-1}$] $\times$ [ind ha$^{-1}$], so that:
+%
+$w_X$ is in [mg ha$^{-1}$] or [mg population$^{-1}$], which is the quantity of contaminant within the population.
+%
+The dynamic of the new variable is defined with:
+%
+\begin{equation}
+w_X = C_X B_X X = C_X B_{T,X} \quad \Rightarrow \quad \frac{dw_X}{dt} = B_X \frac{dX C_X}{dt} = B_X \left( X \frac{dC_X}{dt} + C_X \frac{dX}{dt} \right)
+\end{equation}
+%
+and $\dfrac{dw_X}{dt}$ is in [mg ha$^{-1}$ day$^{-1}$].
+%
+
 \subsubsection{Dynamics of contaminant within populations}
 
-With the previous assumption, the general model is described as follow:
+With all the assumptions made in the manuscript, especially the assumption that individuals of a same population have the same biomass and the same amount of contaminant in their bodies, the model for the transfer of a contaminant  is given by:
+%
 
-\paragraph{Population dynamics}
+\begin{subequations}
+	\begin{empheq}[left={\empheqlbrace\,}]{align}
+	% OUTSIDE toxic
+	\dfrac{dw_{ext}}{dt} & =  T_{Broma}(V) - k_0w_{ext} - \kappa(w_{ext})B_{T,V} 
+	\label{eq:General_Mod:a} \\[3mm]
+	% VOLE Dynamic
+	\dfrac{dB_{T,V}}{dt} & = 
+	B_{T,V} r_V\left( 1- \dfrac{V}{K_V} \right)	- \mu_V(\zeta_V)B_{T,V} -  B_V (\Phi_{V,M}(V) M - \Phi_{V,F}(V,M) F)
+	\label{eq:General_Mod:b} \\[3mm]
+	% VOLE toxic
+	\dfrac{dw_{V}}{dt} & = \kappa(w_{ext}) B_{T,V} - w_V(k_{out, V} + \mu(\zeta_V)) - \notag\\[3mm]
+	&  C_V B_V ( \Phi_{V,M}(V)M + \Phi_{V,F}(V,M)F )
+	\label{eq:General_Mod:c} \\[3mm]
+	%
+	% MUSTELID dynamic
+	\dfrac{dB_{T,M}}{dt} & = 
+	B_{T,M} (\varepsilon_MB_V  \Phi_{V,M}(V) - m_M - \mu_M(\zeta_M) ) - B_M \Phi_{M,F}(V, M) F
+	\label{eq:General_Mod:d} \\[3mm]
+	% MUSTELID toxic
+	\dfrac{dw_{M}}{dt} & =  - k_{out, M} w_M - w_M (m_M + \mu_M(\zeta_M))  - C_M B_M \Phi_{M,F}(V,M)F   + \notag \\[3mm]
+	&  \eta_M C_V B_V \Phi_{V,M}(V)M 
+	\label{eq:General_Mod:e} \\[3mm]
+	%
+	% FOX dynamic
+	\dfrac{dB_{T,F}}{dt} & = 
+	B_{T,F} r_F\left( 1- \dfrac{F}{K_F} \right)	- \mu_F(\zeta_F) B_{T,F}
+	\label{eq:General_Mod:f}  \\[3mm]
+	% FOX toxic
+	\dfrac{dw_{F}}{dt} & =  - k_{out, F} w_F - w_F \mu_F(\zeta_F) + \notag \\[3mm]
+	& \eta_F \left( C_V B_V \Phi_{V,F}(V,M) F +  C_M B_M \Phi_{M,F}(V,M) F\right)  
+	\label{eq:General_Mod:g}
+	\end{empheq}
+	\label{eq:General_Mod}
+\end{subequations}
 
-From the equation \eqref{eq:biomTOpop}, we have: $\dfrac{d B_{T,X}}{dt} = B_X \dfrac{d X}{dt}$
+In the general model \eqref{eq:General_Mod}, the equations \eqref{eq:General_Mod:b}, \eqref{eq:General_Mod:d} and \eqref{eq:General_Mod:f} present the dynamic of the total biomass of voles, mustelids and foxes respectively.
+%
+The equations \eqref{eq:General_Mod:a}, \eqref{eq:General_Mod:c}, \eqref{eq:General_Mod:e} and \eqref{eq:General_Mod:g} stand for the dynamics of contaminant concentrations in the environment and within individuals for respectively the voles, mustelids and foxes.
+The parameter $k_{out,X}$ is the excretion rate of the contaminant and the death of individuals implies a decrease of amounts of the contaminant, respectively $ -w_X m_X$ and $-w_X \mu_X(\zeta_X)$ with $\zeta_X$ being either the internal concentration $C_X$ or the dose eaten $D_X$.
+For predators, data provide a dose-response mortality function resulting from a the ingestion dose and not the internal concentration.
+%
+The transfer of the contaminant through the predation is express by the term $ C_X B_X \Phi(X)Y$, which is removed in prey contaminant dynamic $w_X$ and added in the predator contaminant dynamic $w_Y$ (with a factor of transfer efficiency $\eta_Y$).
+%
 
-\paragraph{Conversion of $w_V$ into $C_V$}
+While we could consider the population dynamic in term of biomasses, it is more convenient to use density of population as suggested by the change of variables which are as follows:
 
-\begin{equation}
-\begin{array}{c l}
-\dfrac{d C_V}{dt} & = \dfrac{d w_V / B_{T,V}}{dt} = \dfrac{1}{B_{T,V}^2} \left( \dfrac{d w_V}{dt} B_{T,V} - w_V \dfrac{d B_{T,V}}{dt}\right) \\[3mm]
-%
-& \\[3mm]
-\end{array}
-\end{equation}
+\paragraph{Internal concentration}~\\
+
+\begin{enumerate}
+	\item Conversion of the dynamics of biomass of population, $B_{T,X}$ into dynamics of the density, $X$. From the equation \eqref{eq:biomTOpop}, we have: $\dfrac{d B_{T,X}}{dt} = B_X \dfrac{d X}{dt}$
+	\item Conversion of amount of contaminant, $w_X$, into within population concentration, $C_X$:	
+	$$\dfrac{d C_X}{dt}  = \dfrac{d w_X / B_{T,X}}{dt} = \dfrac{1}{B_{T,X}^2} \left( \dfrac{d w_X}{dt} B_{T,X} - w_X \dfrac{d B_{T,X}}{dt}\right)$$
+\end{enumerate}
+
+Using these properties, we can convert part of system \eqref{eq:General_Mod}:
 
@@
 \left\lbrace
-\begin{array}{c l}
-% OUTSIDE toxic
-\dfrac{dw_{ext}}{dt} & = T_{Broma}(V) - k_0w_{ext} - \kappa(w_{ext})B_{T,V} \\[3mm]
-% VOLE Dynamic
-\dfrac{dB_{T,V}}{dt} & =
-B_{T,V} g_V(V, C_V)	- \mu_V(C_V)B_{T,V} - B_V \Phi_{V,M}(V) M - B_V \Phi_{V,F}(V,M) F\\[3mm]
-% VOLE toxic
-\dfrac{dw_{V}}{dt} & = \kappa(w_{ext}) B_{T,V} - k_{out, V} w_V - w_V \mu(C_V) - C_V B_V \Phi_{V,M}(V)M + \gamma_V w_V g_V(V, C_V)\\[3mm]
+\begin{array}{c c l}
+\dfrac{dC_V}{dt} & = & \dfrac{d w_V / B_{T,V}}{dt} = \dfrac{1}{B_{T,V}^2} \left( \dfrac{d w_V}{dt} B_{T,V} - w_V\dfrac{d B_{T,V}}{d t}\right) \\[3mm]
 %
-% DEAD VOLE Dynamic
-\dfrac{dB_{T,V_d}}{dt} & = \mu_V(C_V)B_{T,V}\\[3mm]
-% DEAD VOLE toxic
-\dfrac{dw_{V_d}}{dt} & = \dfrac{dw_{V}}{dt}\\[3mm]
+& = & \dfrac{1}{B_{T,V}} \left(\kappa(w_{ext})B_{T,V} - k_{out,V}w_V - w_V \mu_V(C_V) - C_V B_V(\Phi_{V,M}(V)M + \Phi_{V,F}(V,M)F ) \right) - \\[3mm]
+& & \dfrac{w_V}{B_{T,V}^2} \left( B_{T,V} g_V{V,C_V} - \mu_V(C_V) B_{T,V} - B_V(\Phi_{V,M}(V)M + \Phi_{V,F}(V,M)F \right)  \\[3mm]
 %
-% MUSTELID dynamic
-\dfrac{dB_{T,M}}{dt} & =
-B_{T,M} (\varepsilon_MB_V  \Phi_{V,M}(V)M - m_M - \mu_M(C_M) ) - B_M \Phi_{M,F}(V, M) F\\[3mm]
-% MUSTELID toxic
-\dfrac{dw_{M}}{dt} & = - k_{out, M} w_M - w_M (m_M + \mu(C_M)) + \eta_M C_M B_M \Phi_{V,M}(V)M - \\[3mm]
-& C_M B_M \Phi_{M,F}(V,M)F + \gamma_M w_M g_M(M, C_V)\\[3mm]
-%
-% FOX dynamic
-\dfrac{dB_{T,F}}{dt} & =
-B_{T,F} g_F(F, C_F)	- \mu_F(C_F) \\[3mm]
-% FOX toxic
-\dfrac{dw_{F}}{dt} & = - k_{out, F} w_F - w_F \mu(C_F) + \eta_F C_M B_M \Phi_{M,V} F  + \eta_F C_V B_V \Phi_{V,V} F + \gamma_F w_F g_F(F, C_F)\\[3mm]
+& = & \kappa(w_{ext}) - C_V \left( k_{out,V} + r_V\left( 1- \dfrac{V}{K_V} \right)\right)
 \end{array}
@@
 %
+\begin{equation}
+\left\lbrace
+\begin{array}{c c l}
+\dfrac{dC_M}{dt} & = & \dfrac{d w_M / B_{T,M}}{dt} = \dfrac{1}{B_{T,M}^2} \left( \frac{d w_M}{d t} B_{T,M} - w_M \frac{dB_{T,M}}{dt} \right)\\[3mm]
+%
+& = & \dfrac{B_{T,M}}{B_{T,M}^2} \left[ - k_{out,M} w_M - w_M(m_M + \mu_M(D_M)) + \eta_M C_V B_V \Phi_{V,M}(V)  - C_M B_M \Phi_{M,F}(V,M)F \right] - \\[3mm]
+& &\dfrac{w_M}{B_{T,M}^2]} \left[ B_{T,M} (\varepsilon_M B_V \Phi_{V,M}(V) - m_M - \mu_M(D_M) ) - B_M \Phi_{M,F}(V,M)F  \right]\\[3mm]
+%
+& = &\dfrac{1}{B_{T,M}} \left[ C_M B_{T,M} (- k_{out,M} - m_M - \mu_M(D_M)- \Phi_{M,F}(V,M)F) + \eta_M C_V B_V \Phi_{V,M}(V)M  \right]  -  \\[3mm]
+&  & C_M(\varepsilon_M B_V \Phi_{V,M}(V) - m_M - \mu_M(D_M)) + \frac{B_M}{B_{T,M}} C_M \Phi_{M,F}(V,M)F\\[3mm]
+%
+& = & C_M(-k_{out,M} - m_M - \mu_M(D_M)) + \dfrac{1}{B_{T,M}} \eta_M C_V B_V \Phi_{V,M}(V)M - \Phi_{M,F}(V,M)F - \\[3mm]
+&  & C_M \varepsilon_M B_V \Phi_{V,M}(V) + C_M(m_M + \mu_M(D_M) + \Phi_{M,F}(F)F \\[3mm]
+%
+& = & \eta_M C_V \dfrac{B_V}{B_M} \Phi_{V,M}(V) - C_M \left( k_{out,M} + \varepsilon_M B_V \Phi_{V,M}(V)\right)
+\end{array}
+\right.
+\end{equation}
 
-Thus, it is straightforward to express the dynamics as:
+\begin{equation}
+\left\lbrace
+\begin{array}{c c l}
+\dfrac{d C_F}{dt} & = & \dfrac{d w_F / B_{T,F}}{dt} = \dfrac{1}{B_{T,F}^2} \left( \dfrac{d w_F}{dt} B_{T,F} - w_F \dfrac{d B_{T;F}}{dt}\right) \\[3mm]
+%
+& = & \dfrac{1}{B_{T,F}} \left( - k_{out,F}w_F - w_F \mu_F(D_F) + \eta_F(C_M B_M \Phi_{M,F}(V,M)F + C_V B_V \Phi_{V,F}(V,M)F) \right) - \\[3mm] 
+& & \dfrac{w_F}{B_{T,F}^2} \left( B_{T,F} r_F\left( 1- \dfrac{F}{K_F} \right) - B_{T,F} \mu_F(D_F) \right) \\[3mm]
+%
+& = & \eta_F \left( C_V \dfrac{B_V}{B_F} \Phi_{V,F}(V,M) + C_M \dfrac{B_M}{B_F} \Phi_{M,F}(V,M)  \right) - C_F \left( k_{out,F} + r_F\left( 1 - \dfrac{F}{K_F} \right)\right) 
+\end{array}
+\right.
+\end{equation}
+
+%
+In addition, for dead voles, we assume the following dynamics:
+
+\begin{equation}
+\dfrac{dB_{T,V_d}}{dt}  = \mu_V(C_V)B_{T,V}
+\end{equation}
+
+and
+
+\begin{equation}
+\dfrac{dw_{V_d}}{dt}  = \dfrac{dw_{V}}{dt}
+\end{equation}
+%
+
+\paragraph{Dose ingested}~\\
+As for internal concentration, we first define the total amount of contaminant eaten by population $X$ $w_{exp,X}$ in [mg pop$_X^{-1}$ day$^{-1}$ or mg ha$^{-1}$ day$^{-1}$] and derive the dose as:
 
@@
 \begin{array}{c  l}
+\dfrac{dD_M}{dt} & =  \dfrac{d w_{exp,M} / B_{T,M}}{dt} = \dfrac{1}{B_{T,M}^2} \left(\dfrac{d w_{exp,M}}{dt} B_{T,M} - \dfrac{d B_{T,M}}{dt} w_{exp,M}\right)
+\\[3mm]
+%
+& =  \dfrac{1}{B_{T,M}} C_V B_V \Phi_{V,M}(V)M - \\[3mm]
+& \dfrac{D_M B_{T,M}}{B_{T,M}^2} \left( 	B_{T,M} (\varepsilon_MB_V  \Phi_{V,M}(V) - m_M - \mu_M(D_M) ) - B_M \Phi_{M,F}(V, M) F \right)	\\[3mm]
+& =  \dfrac{C_V B_V \Phi_{V,M}(V)}{B_M} - D_M	(\varepsilon_MB_V  \Phi_{V,M}(V) - m_M - \mu_M(D_M) ) + \dfrac{D_M \Phi_{M,F}(V, M) F}{M}  	
+\end{array}
+\right.
+\end{equation}
+
+And, with the same approach for foxes:
+
+\begin{equation}
+\left\lbrace
+\begin{array}{c  l}
+\dfrac{dD_F}{dt} & =  \dfrac{d w_{exp,F} / B_{T,F}}{dt} = \dfrac{1}{B_{T,F}^2} \left(\dfrac{d w_{exp,F}}{dt} B_{T,F} - \dfrac{d B_{T,F}}{dt} w_{exp,F}\right)
+\\[3mm]
+%
+& =  \dfrac{1}{F B_F} (C_V B_V \Phi_{V,F}(V,M)F + C_M B_M \Phi_{M,F}(V,M)F ) - \\[3mm]
+& \dfrac{D_F }{B_{T,F}} \left( 	B_{T,F} r_F\left( 1- \dfrac{F}{K_F} \right)	- \mu_F(D_F) B_{T,F}\right)	\\[3mm]
+& =  \dfrac{C_V B_V \Phi_{V,F}(V,M) + C_M B_M \Phi_{M,F}(V,M)}{B_F} -  D_F \left(r_F \left( 1- \dfrac{F}{K_F} \right)	- \mu_F(D_F)  \right)	
+\end{array}
+\right.
+\end{equation}
+
+\paragraph{General model}~\\
+
+As a consequence of all these assumptions, the model as implemented for simulations is given by:
+
+\begin{equation}
+\left\lbrace
+\begin{array}{c c l}
 % VOLE Dynamic
-\dfrac{dV}{dt} & =
-\overbrace{	r_V V (1-V/K_V)		}^{\text{growth rate}} -
+\dfrac{dV}{dt} & = &
+\overbrace{	r_V V \left( 1- \dfrac{V}{K_V} \right)		}^{\text{growth rate}} -
 \overbrace{	\mu_V(C_V)V	}^{\text{poisoning}} -
@@
 % DEAD VOLE Dynamic
-\dfrac{dV_d}{dt} & =
+\dfrac{dV_d}{dt} & = &
 \overbrace{ \mu_V(C_V)V }^{\text{new dead voles}} -
@@
 % MUSTELIDS Dynamic
-\dfrac{dM}{dt} & =
-\overbrace{	(\varepsilon_M\Theta_M(V,V_d)-m_M) M	}^{\text{growth rate}} - 
-\overbrace{\mu_M(C_M)M	}^{\text{poisoning}} -
+\dfrac{dM}{dt} & = &
+\overbrace{	(\varepsilon_M B_V \Theta_M(V,V_d)-m_M) M	}^{\text{growth rate}} - 
+\overbrace{\mu_M(D_M)M	}^{\text{poisoning}} -
 \overbrace{\Phi_{M,F}(V,V_d,M)F }^{\text{predation}} \\[3mm]
@@
 % F0X Dynamic
-\dfrac{dF}{dt} & =
+\dfrac{dF}{dt} & = &
 \overbrace{	r_F F (1-F/K_F)	}^{\text{growth rate}} -
-\overbrace{\mu_F(C_F)F	}^{\text{poisoning}}  \\
+\overbrace{\mu_F(D_F)F	}^{\text{poisoning}}  \\
 %
@@
 % Toxicological dynamics
-\dfrac{dC}{d t}  & = 
+\dfrac{dw_{ext}}{d t}  & = & 
 \overbrace{T_{Broma}(V)}^{\text{treatment}} - 
-\overbrace{k_0 C}^{\text{degradation}} -
-\overbrace{\kappa(C)}^{\text{consumption}} \\[3mm]
+\overbrace{k_0 w_{ext}}^{\text{degradation}} -
+\overbrace{\kappa(w_{ext}) V}^{\text{consumption}} \\[3mm]
 %
 % Toxico-kinetics in Voles
-\dfrac{dC_V}{dt} &= 
+\dfrac{dC_V}{dt} & = &
 \overbrace{\kappa(C)}^{\text{intake}} - 
@@
 % Toxico-kinetics in Mustelids
-\dfrac{dC_M}{d t} & =
+\dfrac{dC_M}{d t} & = &
 \overbrace{\eta_M \dfrac{B_V}{B_M} (\Phi_{V,M}(V,V_d,M)+\Phi_{V_d,M}(V,V_d,M)) C_V}^{\text{intake}} -
-\overbrace{C_M\left(k_{out,M} + \varepsilon_M \Phi_{V,M}(V,V_d,M) \right)}^{\text{excretion}} \\[3mm]
+\overbrace{C_M\left(k_{out,M} + \varepsilon_M B_V \Phi_{V,M}(V,V_d,M) \right)}^{\text{excretion}} \\[3mm]
+%
+\dfrac{dD_M}{d t} & = &
+\overbrace{
+\dfrac{ B_V }{B_M}C_V\Phi_{V,M}(V)}^{\text{exposure}}
+\overbrace{ - D_M \left(\varepsilon_MB_V  \Phi_{V,M}(V)
+ - m_M - \mu_M(D_M) -\Phi_{M,F}(V, M) \dfrac{F}{M} \right)}^{\text{population change}}   \\[3mm]
 %
 % Toxico-kinetics in Foxes
-\dfrac{dC_F}{d t} &=
-\overbrace{ \eta_F \dfrac{B_V}{B_F} \left( (\Phi_{V,M}(V,V_d,M)+\Phi_{V_d,M}(V,V_d,M)) C_V + \Phi_{M,F}(V,V_d,M) C_M \right)}^{\text{intake}} - \\[3mm]
-&
-\overbrace{k_{out,F} + r_F \left(1 - \dfrac{F}{K_F}\right) }^{\text{excretion}} 
+\dfrac{dC_F}{d t} & = &
+\overbrace{ \eta_F  \left( \dfrac{B_V}{B_F} (\Phi_{V,M}(V,V_d,M)+\Phi_{V_d,M}(V,V_d,M)) C_V + \dfrac{B_M}{B_F}\Phi_{M,F}(V,V_d,M) C_M \right)}^{\text{intake}} - \\[3mm]
+& &
+\overbrace{C_F \left( k_{out,F} + r_F \left(1 - \dfrac{F}{K_F}\right) \right) }^{\text{excretion}}  \\[3mm]
+\dfrac{dD_F}{d t} & = &
+\overbrace{
+ \dfrac{C_V B_V \Phi_{V,F}(V,M) + C_M B_M \Phi_{M,F}(V,M)}{B_F} }^{\text{exposure}} 
+\overbrace{ - D_F \left(r_F \left( 1- \dfrac{F}{K_F} \right)	- \mu_F(D_F)  \right)	}^{\text{population change}}
 \end{array}
@@
 
+ \clearpage
+
+\section{Additional results of simulations}
+
+\subsection{Results with Foxes}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/graph_0_0.png}
+		\caption{Variation of population densities and contaminant concentration in soil simulated over 50 years.
+Scenario \textbf{a}: 0 AR.
+Variables are: $D.F$ for fox density, $D.M$ for mustelid density, $D.V$ for vole density, $D.Vd$ for dead vole density and $C.soil$ for AR concentration in the environment.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/graph_500_20.png}
+		\caption{Variation of population densities and contaminant concentration in soil simulated over 50 years.
+Scenario \textbf{b}: threshold 500 in ha$^{-1}$ and 20 kg ha$^-1$ of AR.
+Variables are: $D.F$ for fox density, $D.M$ for mustelid density, $D.V$ for vole density, $D.Vd$ for dead vole density and $C.soil$ for AR concentration in the environment.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/graph_250_20.png}
+		\caption{Variation of population densities and contaminant concentration in soil simulated over 50 years.
+Scenario \textbf{c}: threshold 250 in ha$^{-1}$ and 20 kg ha$^-1$ of AR.
+Variables are: $D.F$ for fox density, $D.M$ for mustelid density, $D.V$ for vole density, $D.Vd$ for dead vole density and $C.soil$ for AR concentration in the environment.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/graph_250_7_5.png}
+		\caption{Variation of population densities and contaminant concentration in soil simulated over 50 years.
+Scenario \textbf{d}: threshold 250 in ha$^{-1}$ and 7.5 kg ha$^-1$ of AR.
+Variables are: $D.F$ for fox density, $D.M$ for mustelid density, $D.V$ for vole density, $D.Vd$ for dead vole density and $C.soil$ for AR concentration in the environment.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/graph_50_7_5.png}
+		\caption{Variation of population densities and contaminant concentration in soil simulated over 50 years.
+Scenario \textbf{e}: threshold 50 in ha$^{-1}$ and 7.5 kg ha$^-1$ of AR.
+Variables are: $D.F$ for fox density, $D.M$ for mustelid density, $D.V$ for vole density, $D.Vd$ for dead vole density and $C.soil$ for AR concentration in the environment.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/plt_sobolIndex_0_0.png}
+		\caption{First-order Sobol's sensitivity index. Scenario \textbf{a}: No AR.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/plt_sobolIndex_500_20.png}
+		\caption{First-order Sobol's sensitivity index. Scenario \textbf{b}: threshold 500 voles in ha$^{-1}$ and 20 kg ha$^-1$ of AR.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/plt_sobolIndex_250_20.png}
+		\caption{First-order Sobol's sensitivity index. Scenario \textbf{c}: threshold 250 voles in ha$^{-1}$ and 20 kg ha$^-1$ of AR.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/plt_sobolIndex_250_7_5.png}
+		\caption{First-order Sobol's sensitivity index. Scenario \textbf{d}: threshold 250 voles in ha$^{-1}$ and 7.5 kg ha$^-1$ of AR.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/plt_sobolIndex_50_7_5.png}
+		\caption{First-order Sobol's sensitivity index. Scenario \textbf{e}: threshold 50 voles in ha$^{-1}$ and 7.5 kg ha$^-1$ of AR.}
+	\end{center}
+\end{figure}
+
+ \clearpage
+
+\subsection{Results without Foxes}
+
+\begin{table}[!htb]
+	\caption{Results on the cost functions for the comparison of treatment protocols.
+Column \textit{"FFR"} (farmer functional response) refers to the protocol denotation (a to e) and in brackets respectively: the vole density threshold, and the amount of AR/treatment in kg.
+The column \textit{"Nbr treat/yr"} refers to the number of treatment in simulation per year after the burn-in phase.
+\textit{"Total AR/yr"} is the total use of anticoagulant rodenticides per year (bromadiolone in kg).
+\textit{"M, 50\% AR death"} is the proportion of time with mortality by AR over 50\% in mustelids population.
+\textit{"M, 50\% Nat.
+death"} is the proportion of time with mortality by natural reason (i.e., all except AR) over 50\% in mustelids population.
+\textit{"V < 50 ha$^{-1}$"} is the proportion of time with less than 50 voles ha$^{-1}$.
+Finally, \textit{"Mean V density"}, \textit{"Mean M density"} and \textit{"Mean F density"} are mean densities of voles, mustelids and foxes respectively.}
+	\label{tab:SMresults}
+	\begin{tabular}{p{2cm} c c c c c c c c}
+		\hline
+		\tabcell{FFR \\ (dose, limit)} &
+		\tabcell{Nbr. \\ treat/yr} &
+		\tabcell{Total  \\ AR/yr} &
+		\tabcell{M, 50\% \\ AR death} &
+		\tabcell{M, 50\% \\ Nat. death} &
+		\tabcell{V < \\ 50 ha$^{-1}$} &
+		\tabcell{Mean V \\ density} & 
+		\tabcell{Mean M \\ density} & 
+		\tabcell{Mean F \\ density} \\
+		\hline
+		%%
+		a (0,0)    & 0& 0& 0& 1& 0.5               & 215 & 0.61   & 0\\
+		% 
+		b (500,20) & 4.4& 88 & 0.03 &0.97 & 0.4     & 185 & 0.57 & 0\\
+		%
+		c (250, 20) & 2.3 & 46 &0.06 & 0.93 & 0.2 & 110 & 0.31 & 0\\
+		%
+		d (250, 7.5)& 4.9 & 37 & 0.04 & 0.96 & 0.3 & 116 & 0.58 & 0\\
+		%
+		e (50,7.5) & 5.6 &  42 & 0 & 1 & 0.9   & 55 & 0.08 & 0 \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/plot_join_0F.png}
+		\caption{Rate of variation of population densities simulated over 50 years (10 years of burn-in period in grey area): (Left) Dynamics of mustelids (grey line) and red foxes (black line), (Right) Dynamics of live (grey line) and dead (black line) voles.
+Letters at the top-center part of each sub-graph corresponds to the results obtained from each farmer functional response.
+Red area and Green area indicate the farmer-regulated (FR) and mustelid-regulated (MR) periods of vole population dynamics respectively.}
+		\label{fig:populations}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/graph_0_0_0F.png}
+		\caption{Variation of population densities and contaminant concentration in soil simulated over 50 years.
+Scenario \textbf{a}: 0 AR.
+Variables are: $D.F$ for fox density, $D.M$ for mustelid density, $D.V$ for vole density, $D.Vd$ for dead vole density and $C.soil$ for AR concentration in the environment.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/graph_500_20_0F.png}
+		\caption{Variation of population densities and contaminant concentration in soil simulated over 50 years.
+Scenario \textbf{b}: threshold 500 in ha$^{-1}$ and 20 kg ha$^-1$ of AR.
+Variables are: $D.F$ for fox density, $D.M$ for mustelid density, $D.V$ for vole density, $D.Vd$ for dead vole density and $C.soil$ for AR concentration in the environment.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/graph_250_20_0F.png}
+		\caption{Variation of population densities and contaminant concentration in soil simulated over 50 years.
+Scenario \textbf{c}: threshold 250 in ha$^{-1}$ and 20 kg ha$^-1$ of AR.
+Variables are: $D.F$ for fox density, $D.M$ for mustelid density, $D.V$ for vole density, $D.Vd$ for dead vole density and $C.soil$ for AR concentration in the environment.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/graph_250_7_5_0F.png}
+		\caption{Variation of population densities and contaminant concentration in soil simulated over 50 years.
+Scenario \textbf{d}: threshold 250 in ha$^{-1}$ and 7.5 kg ha$^-1$ of AR.
+Variables are: $D.F$ for fox density, $D.M$ for mustelid density, $D.V$ for vole density, $D.Vd$ for dead vole density and $C.soil$ for AR concentration in the environment.}
+	\end{center}
+\end{figure}
+
+\begin{figure}[!htb]
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/graph_50_7_5_0F.png}
+		\caption{Variation of population densities and contaminant concentration in soil simulated over 50 years.
+Scenario \textbf{e}: threshold 50 in ha$^{-1}$ and 7.5 kg ha$^-1$ of AR.
+Variables are: $D.F$ for fox density, $D.M$ for mustelid density, $D.V$ for vole density, $D.Vd$ for dead vole density and $C.soil$ for AR concentration in the environment.}
+	\end{center}
+\end{figure}
+
 % BIBLIOGRAPHY
@@
 
-%\bibliographystyle{elsarticle-harv}
-\bibliographystyle{plainnat}
+ \clearpage
+
+\bibliographystyle{elsarticle-harv}
+%\bibliographystyle{plainnat}
 %\bibliographystyle{unsrt}

### 2019-06-03 trophToxNTO/R/events.R
update package

@@ -9,8 +9,6 @@
 #' @export
 #'
 #'
-
-
 condition = function(Vole.Level,Broma.Level){
   root.event = function(Time, State, Pars){
     with(as.list(c(State, Pars)), { D.V = State[(4*n+1):(5*n)]
@@ -28,11 +26,14 @@ condition = function(Vole.Level,Broma.Level){
       D.Vd = State[(5*n+1):(6*n)]
       D.M = State[(6*n+1):(7*n)]
       D.F = State[(7*n+1):(8*n)]
+      Cing.V = State[(8*n+1):(9*n)]
+      Cing.M = State[(9*n+1):(10*n)]
+      Cing.F = State[(10*n+1):(11*n)]
 
       C.soil=(round((D.V - Vole.Level),5)==0)*Broma.Level #rep(Broma.Level,n) # ppm * kg
 
-      C.V=C.V ;  C.M=C.M ; C.F=C.F ;  D.V = D.V ;  D.Vd= D.Vd ;  D.M= D.M ;  D.F = D.F
-      return(c(C.soil,C.V,C.M,C.F,D.V,D.Vd,D.M,D.F))
+      C.V=C.V ;  C.M=C.M ; C.F=C.F ;  D.V = D.V ;  D.Vd= D.Vd ;  D.M= D.M ;  D.F = D.F ; Cing.V = Cing.V ; Cing.M = Cing.M ; Cing.F = Cing.F
+      return(c(C.soil,C.V,C.M,C.F,D.V,D.Vd,D.M,D.F,Cing.V, Cing.M,Cing.F))
     })
   }
 

### 2019-06-03 trophToxNTO/R/ode_solver.R

@@ -57,11 +57,9 @@ dose.resp.mort=function(C,LD.50,H.coef,days.death){
 #' @export
 #'
 #'
-
-intake.V=function(C,max.intake.C,D50.intake.C){
+intake.V=function(C, max.intake.C, D50.intake.C){
   return(max.intake.C*C/(D50.intake.C+C))
 }
-
 
 #' The system of differential equations to integrate with deSolve package
 #'
@@ -76,7 +74,89 @@ intake.V=function(C,max.intake.C,D50.intake.C){
 #' @export
 #'
 #'
+ode.broma.tritrophic_2nd = function(Time, State, Pars) { # The model is a function
+  with(as.list(c(State, Pars)), { # You need State and Pars in the next lines
 
+    C.soil = State[1:n]
+    C.V = State[(n+1):(2*n)]
+    C.M = State[(2*n+1):(3*n)]
+    C.F = State[(3*n+1):(4*n)]
+    D.V = State[(4*n+1):(5*n)]
+    D.Vd = State[(5*n+1):(6*n)]
+    D.M = State[(6*n+1):(7*n)]
+    D.F = State[(7*n+1):(8*n)]
+    Cing.V = State[(8*n+1):(9*n)]
+    Cing.M = State[(9*n+1):(10*n)]
+    Cing.F = State[(10*n+1):(11*n)]
+
+    # Contaminant concentration
+    #C.soil=approx.broma.farmer.input(Time) # See afterward, for external events, we have to approximate for ode integration
+
+    dC.soil = -k.0*C.soil - intake.V(C.soil,max.intake.C,D50.intake.C) * D.V * B.V  #because we work in kg of species, and parameterization is in g
+    # dC.soil =  -k.0*C.soil
+
+    dC.V = intake.V(C.soil,max.intake.C,D50.intake.C) - C.V*(k.out.V+r.V*(1-D.V/K.V))
+
[…]

### 2019-06-03 trophToxNTO/R/scriptTOrun.R

@@ -1,33 +1,0 @@
-library(trophToxNTO)
-## --------------------------- TIME ------------------------
-times <- seq(from=0, to = 50*365, length = 1e4) ##  30*5 = 5 months
-## --------------------------- EVENEMENTS -------------------
-root.event=condition(Vole.Level=500,Broma.Level=20*50)[[1]]
-event.func=condition(Vole.Level=500,Broma.Level=20*50)[[2]]
-## ------------------------- RUN
-params = pars(n=1)
-init.p <- init(params$n)
-# # Set fix parameters
-# range_param = rbind(unlist(fix_pVar_pars(params,-0.1)), unlist(fix_pVar_pars(params,0.1)))
-# ls_params = as.list(as.data.frame(range_param))
-# ls_params_expand <- expand.grid(ls_params[2:ncol(range_param)])
-
-# # Random parameters
-ls_params_expand <- lapply(1:1e3, function(i) rand_pVar_pars(params))
-
-ls_randSens <- lapply(1:length(ls_params_expand), function(i){
-  tryCatch({
-    model_output = ode(y=init.p,
-                       times,
-                       ode.broma.tritrophic,
-                       parms = ls_params_expand[[i]],
-                       events = list(func = event.func, root = TRUE),
-                       rootfun = root.event)
-    df_temp <- tab_sensitivity(model_output, ls_params_expand[[i]])
-    return(df_temp)},
-    error=function(e) NULL)
-})
-# mat_randSens <- do.call("rbind", ls_randSens)
-df_randSens <- dplyr::bind_rows(ls_randSens)
-# save(df_randSens, file = paste0("data/df_randSens_", Sys.time(), "_", round(runif(1,1e10,1e11)), ".rda"))
-save(df_randSens, file = paste0("df_randSens_", Sys.time(), "_", round(runif(1,1e10,1e11)), ".rda"))

### 2019-06-03 trophToxNTO/vignettes/getting_started.R

@@ -16,61 +16,6 @@ scenario = condition(Vole.Level=vole.level.test,Broma.Level=broma.level.test)
 root.event=scenario[[1]]
 event.func=scenario[[2]]
 
-## ----warning=FALSE-------------------------------------------------------
-# Load package deSolve
-# library(deSolve)
-
-# number of parallel simulation
-params=pars(n=1)
-
-# times vector of integration
-times <- seq(from=0, to=15*365, length.out = 1e3) #  30*5 = 5 months 
-
-# Run the model
-
-model.output    <- ode(times = times,
-                       y = init(params$n),
-                       func = ode.broma.tritrophic,
-                       parms  = params,
-                       events = list(func = event.func, root = TRUE),
-                       rootfun = root.event)
-
-
 ## ----warning=FALSE,message=FALSE-----------------------------------------
-graphique(params$n,times,model.output) 
-
-## ----warning=FALSE-------------------------------------------------------
-# number of parallel simulation
-params=pars(n=2)
-
-# Run the model
-model.output    <- ode(times = times,
-                       y = init(params$n),
-                       func = ode.broma.tritrophic,
-                       parms = pars(
-                         n=params$n, # Do not foret to specify this !!!
-                         K.V=seq(200,800,length=params$n)
-                       ),
-                       events = list(func = event.func, root = TRUE),
-                       rootfun = root.event)
-
-# Graphic output
-graphique(params$n,times,model.output) 
-
-## ----warning=FALSE-------------------------------------------------------
-# number of parallel simulation
-params=pars(n=2)
-
-# Run the model
-model.output    <- ode(times = times,
-                       y = init(params$n,
-                                D.F = c(0,0.2)
-                                ),
-                       func = ode.broma.tritrophic,
-                       parms = params,
-                       events = list(func = event.func, root = TRUE),
-                       rootfun = root.event)
-
-# Graphic output
 graphique(params$n,times,model.output) 
 

### 2019-10-24 manuscript/MS_trophToxNTO.tex
up change in files and models

@@
 
+%\begin{center}
+%	\Large\textbf{
+%	Trophic transfer of anticoagulant rodenticides while managing rodent pests: the fine line between predator-prey regulation and pesticide-pest regulation
+%}\par
+%\end{center}
+
 \begin{center}
 	\Large\textbf{
-	Trophic transfer of anticoagulant rodenticides while managing rodent pests: the fine line between predator-prey regulation and pesticide-pest regulation
+		Trophic transfer of pesticides: the fine line between predator-prey regulation and pesticide-pest regulation
 	}\par
@@
 When predators that otherwise regulate pest densities ingest prey contaminated with pesticides, this can suppress predator populations as contaminated prey act as a super-predator, with pesticides controlling both pest and predators.
-It is, however, unknown how species relationships and protocols of rodenticide treatments, i.e.
+It is, however, unknown how species relationships and protocols of anticoagulant rodenticide (AR) treatments, i.e.
 farmer functional responses, interact to affect pest regulation.
 		\item Here we used linked differential equations to model a heuristic non-spatialized system including montane water voles, specialist vole predators (stoats, weasels), and a generalist predator (red fox) which consume voles, mustelids and other prey.
-By considering anticoagulant rodenticide toxicokinetic and toxicodynamic equations, we explored the impact of 5 farmer functional responses (defined by both rodenticide quantity and threshold vole density above which rodenticide spreading is prohibited) on predator-prey interactions, rodenticide transfer across the trophic chain and population effects.
-		\item Spreading rodenticide while maintaining sufficient voles as prey resources led to less rodenticide being applied and extended periods without rodenticide in the environment, benefitting predators while avoiding episodes with high vole density.
+By considering AR toxicokinetic and toxicodynamic, we explored the impact of 5 farmer functional responses (defined by both AR quantity and threshold vole density above which AR spreading is prohibited) on predator-prey interactions, AR transfer across the trophic chain and population effects.
+		\item Keeping vole density below acceptable bounds by spreading AR while maintaining sufficient voles as prey resources led to less AR being applied and extended periods without AR in the environment, benefitting predators while avoiding episodes with high vole density.
 This may meet farm production interests while minimizing the impact on small mustelid and fox populations.
-		Spreading rodenticide to maintain low vole densities suppressed mustelid and fox populations, leading to vole population dynamics being entirely regulated by rodenticide use.
-This vole-eradication treatment regime inhibited predator ecosystem services and promoted pesticide dependence.
-		\item Farmer functional response with high or intermediate densities thresholds triggering treatment produced vole population fluctuations caused by rodenticide that could be followed by periods during which mustelids regulated vole populations.
-These alternative phases of mustelids and farmer regulation highlight the benefit of intraguild relationship where mustelids may rescue foxes from poisoning.
+		Spreading AR to maintain low vole densities suppressed mustelid and fox populations, leading to vole population dynamics being entirely regulated by AR use.
+Such vole-eradication treatment regime inhibited predator ecosystem services and promoted pesticide dependence.
+		\item An alternative population pattern emerges from a compromise between farmer functional response and population sensitivity to AR resulting in vole population fluctuations caused by AR and followed by periods during which mustelids regulated vole populations.
+These alternating phases of mustelids and farmer regulation highlight the benefit of intraguild relationship where mustelids may rescue foxes from poisoning.
+		\item In such relatively simple mechanistic models, parameters encompass a large set of processes from the natural open systems used as reference.
+Both global and wide-range sensitivity analysis illustrated the delicacy to tune an optimal line between predator-prey regulation and pesticide-pest regulation.
 	\end{enumerate}
@@
 That such a pesticide-tri-trophic system may cause a variety of population dynamics responses to pesticide use in agro-ecosystems is a novel insight.
-Our model reveals the need for maintaining refuges with sufficient non-poisoned voles for specialist mustelids, to conserve predator community, given the super-predator role of rodenticides.
+Our model reveals the need for maintaining refuges with sufficient non-poisoned voles for specialist mustelids, to conserve the predator community, given the super-predator role of rodenticides.
 We suggest that long periods without pesticide treatment are essential to maintain predator populations, and that practices of pesticides use that attempt to permanently eradicate a pest over a large scale are counterproductive.
@@
 Thus, under some regimes of pesticides use, pest populations only become regulated by pesticides once predators have collapsed.
-To preserve ecosystem health and the services of predators through regulation of pest densities, we need to assess the feasibility and benefit of pesticide treatment regimes in their ability to control pest species with minimal damage on predators \citep{Halstead2014}.
+To preserve ecosystem health and the services of predators through regulation of pest densities, we need to assess the feasibility and benefit of pesticide treatment regimes in their ability to control pest species with minimal damage to predators \citep{Halstead2014}.
 It is however empirically challenging to assess the overall impact of pesticide treatments on the dynamics of species linked by trophic interactions.
@@
 The smallest mustelids (e.g.
-weasels Mustela nivalis) are specialist vole predators.
+weasels \textit{Mustela nivalis}) are specialist vole predators.
 Their numerical response has been shown theoretically to be necessary for generating predator-prey cycles \citep{Hanski1991}.
-They are said to be responsible for driving 3-to-5-year vole cycles in Fennoscandia (specialist predation hypothesis) \citet{Hanski1991}.
-Generalist resident predators like foxes (Vulpes vulpes) are expected to have regulatory and limiting effects on voles, owing to dietary plasticity that slows down vole population increase at low density \citet{Hanski1991}.
+They are said to be responsible for driving 3-to-5-year vole cycles in Fennoscandia (the specialist predation hypothesis) \citet{Hanski1991}.
+Generalist resident predators like foxes (\textit{Vulpes vulpes}) are expected to have regulatory and limiting effects on voles, owing to dietary plasticity that slows down vole population increase at low density \citet{Hanski1991}.
 Foxes do not show numerical responses to vole abundance \citep{Weber2002} but they influence the food chain through occasional killing and consumption of bite-sized mustelids.
@@
 With the aim of understanding the potentially complex interactions between prey that are perceived as pest predators and farmers spreading rodenticide in response to vole abundance and their functional responses, we studied a simplified system inspired by cyclically fluctuating montane water voles (\textit{Arvicola scherman}), small mustelids (stoats, weasels) that mostly eat voles (specialists), and foxes (generalists), with voles and mustelids as food items.
-We used a process-based model using differential equations to explore 5 FFR types of AR spread, combining population dynamics, predator-prey interactions and rodenticide transfer across the trophic chains.
+We used a process-based model using differential equations to explore 5 FFR types of AR spread, combining population dynamics, predator-prey interactions and rodenticide transfer across the trophic chain.
 Model parameters and FFR were inspired by farming systems in the Jura Mountains, Franche-Comté (France), the region of Comté cheese production.
@@
 Due to recurrent vole outbreaks and damages to grasslands, massive rodenticide treatments were implemented from the early 80s with consequences on non-target wildlife.
-Practices developed technically under pressure from public opinion, farmer unions and farmer technical organizations collaborating with researchers to find treatment regimes with less harmful consequences for biodiversity \citep{Delattre2009}.
+Practices were developed technically under pressure from public opinion, farmer unions and farmer technical organizations collaborating with researchers to find treatment regimes with less harmful consequences for biodiversity \citep{Delattre2009}.
 
 Hence, our main objective was to explore the properties of FFR in relation to varying vole density, to vole outbreak frequencies, to the guild of interacting predators and the effects on the global tri-trophic population dynamics.
-Additionally, we sought to establish under what regime of use the ARs deliver benefits to farmers without imposing damages to the ecosystem.
+Additionally, we sought to establish what regime of use the ARs deliver benefits to farmers without imposing damages to the ecosystem.
 
@@
 
-The vole population followed a logistic growth rate, with $r_V$  the maximal reproduction rate, fixed at $r_V = \ln(2 \times 600)/365$ per day, since montane water vole populations can increase from 0  to  600 individuals ha$^{-1}$ or more \citep{Giraudoux1997} resulting in the equilibrium density being fixed at $K_V = 600$ individuals.
+The vole population followed a logistic growth rate, with $r_V$  the maximal reproduction rate, fixed at $r_V = \ln(2 \times 600)/365$ per day, since montane water vole populations can increase from 0  to  600 individuals ha$^{-1}$ or more \citep{Giraudoux1997}, resulting in the equilibrium density being fixed at $K_V = 600$ individuals.
 %
@@
 
-Figure~\ref{fig:scheme} represents the whole study system.
-Rodenticide is spread in grasslands during treatments, denoted $T_{Broma}(V)$, triggered by vole density $V$ (FFR described in Table~\ref{tab:results}).
+Figure~\ref{fig:scheme} represents the whole study system. Rodenticide is spread in grasslands during treatments, denoted $T_{Broma}(V)$, triggered by vole density $V$.
 %
@@
 %
-AR provided to weasel though vole poisoning, is denoted $D_M$ induces lethal effect at rate $\mu_M(D_M)$, additive to natural mortality rate $m_M$.
-Parameter definition are further detailed in Supporting Information.
+AR contaminating weasel though vole poisoning, is denoted $D_M$, induces lethal effect at rate $\mu_M(D_M)$, additive to natural mortality rate $m_M$.
+Parameter definitions are further detailed in Supporting Information.
 %
@@
 We used log-logistic equations for describing dose-dependent mortality of animals exposed to AR.
-Vole and predator mortality rates due to AR $\mu_X(\zeta_X)$ ($X$ referring to the considered species, and $\zeta_X$ being the dose eaten $D_X$ or the internal concentration $D_X$) were expressed by equation (5):
+Vole and predator mortality rates due to AR $\mu_X(D_X)$ ($X$ referring to the considered species) were expressed by equation (5):
 
 \begin{equation}
-\mu_X(\zeta_X) = \dfrac{1}{\text{experiment duration}} \times \left( 1- \dfrac{1}{1+ (LD_{50}/\zeta_X)^H} \right) 
+\mu_X(\zeta_X) =   1- \dfrac{1}{1+ (LD_{50}/\zeta_X)^H} 
 \label{eq:muX}
@@
 
-where $LD_{50}$ was the daily median lethal dose (50\% of population dying) with a value for voles of 2 mg kg$^{-1}$ as internal concentration, and a dose given as food of  2.1 mg kg$^{-1}$ mustelids \citep{Grolleau1989} and 0.5 mg kg$^{-1}$ for foxes \citep{Sage2010} (see Table~\ref{tab:parameterization}).
-Parameter $H$ is the Hill's coefficient, modulating the curve steepness and was estimated to fit the sparse data we have (Table~\ref{tab:parameterization}).
-Mortality rates considered the duration of AR toxicity from experiments of up to 6 days \citep{Sage2010}.
+where $LD_{50}$ was the daily median lethal dose (50\% of population dying), parameter $H$ is the Hill's coefficient modulating the curve steepness (see Table~\ref{tab:parameterization}).
+%
+Parameterization of $LD_{50}$ is important since it defines the inflexion point of the ecotoxicological effect.
+While it is a classical parameter targeted in experiments, its value can have different order of magnitude  \cite{Grolleau1989,Erickson2004,Sage2010}.
+We explored this parameterization through a wide-range sensitivity analysis detailed here-after (see also Figure\ref{fig:GSA}).
 
@@
 
-We considered a range of realistic FFR spanning treatments during vole outbreaks only as a precautionary approach; in which treatments only takes place at intermediate or low vole density threshold.
-These scenarios are inspired by historic and contemporary protocols of bromadiolone use to control montane water voles in Franche-Comté, but also representative of practice globally (Table~\ref{tab:results}) \citep{Delattre2009}.
-To check the influence of foxes and intraguild predation on the system’s dynamics, we simulated scenarios with and without foxes.
-
+We considered a range of realistic FFR spanning treatments during vole outbreaks only and a precautionary approach in which treatments only takes place at intermediate or low vole density threshold.
+These scenarios are inspired by historic and contemporary protocols of bromadiolone use to control montane water voles in Franche-Comté, but also representative of practice globally \citep{Delattre2009}: (A) a scenario without AR treatment, (B) a scenario with high vole density threshold triggering treatment (500 ind ha$^{-1}$) and high AR amount per treatment (20 kg ha$^{-1}$), (C) a third scenario with intermediate triggering threshold (250 ind ha$^{-1}$) and high AR per treatment (20 kg ha$^{-1}$), (D) scenario with intermediate threshold (250 ind ha$^{-1}$) and low AR amount (7.5 kg ha$^{-1}$) and (E) scenario with low threshold (50 ind ha$^{-1}$) and low treatment (7.5 kg ha$^{-1}$).
+%
+To check the influence of foxes and intraguild predation on the system dynamics, we also simulated scenarios with and without foxes.
+%
 Our simulations tracked the linked vole-mustelid-fox dynamics for 40 years, after a “burn-in” period of 10 years to reduce dependency of results upon initial conditions, to observe several vole cycles and to characterise AR effects on these species population dynamics.
-This burn-in period also had AR treatment triggered at specific vole densities and with a given rodenticide quantity for each FFR (see Table~\ref{tab:results}).
+This burn-in period also had AR treatment triggered at specific vole densities and with a given rodenticide quantity for each FFR.
 The burn-in phase was selected according to a set of simulations with different initial conditions.
 Those simulations showed that in a given FFR (i.e., same threshold of vole density and amount of AR spread), the dynamics of the population were converging toward a similar pattern.
-In order to assess the impact of FFR on both agriculture and conservation interests, we estimated the following cost functions: (i) Number of treatment events per FFR ; (ii) Cumulative amount of AR (kg) ; (iii) Proportion of time when the AR-induced mortality of mustelids higher than 50\% (i.e.
-lethal exposure profile killing 50\% of mustelid population) ; (iv) Proportion of time when the mortality of mustelids was higher than 50\% due to natural mortality; (v) Proportion of time when the vole density was below 50 voles ha$^{-1}$, as a proxy for time when forage grass grows with low herbivore influence; (vi) Mean vole, mustelid and fox densities.
 
@@
 
-All numerical analysis have been done using the Open language R and particularly the ODE solver package \textit{deSolve} \citep{Soetaert2010}.
+All numerical analysis have been done using the open language R and particularly the ODE solver package \textit{deSolve} \citep{Soetaert2010}.
 Model implementation and code to run analysis are available on the Github repository \url{https://github.com/virgile-baudrot/trophToxNTO/}.
+
+We applied a wide-range sensitivity analysis on lethal dose parameters (i.e. $LD_{50}$) for voles, mustelids and foxes, to explore changes in dynamical patterns.
 %
-From simulations, a sensitivity analysis allowed to study \textit{how uncertainties in the model output are apportioned to different sources of uncertainty in the model input} \citep{Saltelli2019}.
+While wide-range variations of each parameter is likely to have impact on the global pattern, the lethal dose parameter is a direct characteristic of the toxicological part of the model.
 %
-We applied a global approach using the first-order Sobol's sensitivity index $S_i$ defined as $S_i = \mathbb{V}(\mathbb{E}_{-i}(y \vert x_i)) / \mathbb{V}(y)$, where $\mathbb{V}(y)$ is the variance of $y$ when all factors are allowed to vary, $\mathbb{E}_{x\sim i}(y \vert x_i))$ the mean of $y$ when one factor is fixed \citep{Sobol1993}.
+Since the $LD_{50}$ parameter of fox appears to have low impact on dynamical pattern in our model, we evaluated only 4 different values (i.e.
+0.5, 1.5, 2 and 7 ppm) which are consistent with a large range covering such a parameter \cite{Erickson2004,Sage2010}.
+%
+For the prey and the specialist predator, this parameter was much more sensitive, so we explored 10 values for both, respectively in the range $[0.7-7]$ ppm and $[1-25]$ ppm following consistent ranges \cite{Grolleau1989,Erickson2004}.
+%
+We classified the subsequent 1600 simulations according to the dynamical pattern of mustelids population: either "periodic regulation by mustelid" or "AR permanent regulation" where mustelids collapse.
+
+Then, we performed a local sensitivity analysis to study changes in model outputs apportioned to small variation around model parameters input \citep{Saltelli2019}.
+%
+We applied a first-order Sobol's sensitivity index $S_i$ defined as $S_i = \mathbb{V}(\mathbb{E}_{-i}(y \vert x_i)) / \mathbb{V}(y)$, where $\mathbb{V}(y)$ is the variance of $y$ when all factors are allowed to vary, $\mathbb{E}_{x\sim i}(y \vert x_i))$ the mean of $y$ when one factor is fixed \citep{Sobol1993}.
 We defined the domain of variation with a beta distribution within an hyperspace of plus or minus 10\% from the original value for all parameters.
+%
+For output variable of local sensitivity analysis, we estimated the following cost functions: (i) Number of treatment events per FFR; (ii) Cumulative amount of AR (kg); (iii) Proportion of time when the AR-induced mortality of mustelids higher than 50\% (i.e.
+lethal exposure profile killing 50\% of mustelid population); (iv) Proportion of time when the mortality of mustelids was higher than 50\% due to natural mortality; (v) Proportion of time when the vole density was below 50 voles ha$^{-1}$, as a proxy for time when forage grass grows with low herbivore influence; (vi) Mean vole, mustelid and fox densities.
 
@@
 
-\subsection{Population dynamics}
-
 Allowing for mortality by predators ingesting AR-poisoned voles changed the outcome of predator-prey dynamics involving vole, mustelid and fox populations.
 Secondary poisoning led to a rich spectrum of emergent dynamics according to the FFR to vole abundance.
-Without AR (FFR a), vole dynamics were regulated by mustelid predation that gave rise to 6-year cycles (Figure~\ref{fig:populations}).
-The fox population remained at its carrying capacity (i.e., 0.03 ind ha$^{-1}$, see Table~\ref{tab:parameterization}).
+Without AR (scenario A), vole dynamics were regulated by mustelid predation that gave rise to a 6-year cycles (Figure~\ref{fig:populations}) typical of prey-specialist dynamics, while the generalist fox population remained at its carrying capacity (i.e., 0.03 ind ha$^{-1}$, see Table~\ref{tab:parameterization}).
+%
+Based on this null model, we explored 4 scenarios of treatment described earlier and denoted respectively B, C, D and E, defined by the vole density threshold triggering treatment (respectively: 500, 250, 250 and 50 ind ha$^{-1}$) and the AR amount per treatment (respectively: 20, 20, 7.5 and 7.5 kg ha$^{-1}$).
 
-Under FFR b (high vole density threshold triggering treatment and high AR amount per treatment) and FFR d (intermediate threshold and low AR amount), vole dynamics were sequentially regulated by either AR treatments, which we refer to as farmer-regulated phase (red color sections in Figure~\ref{fig:populations}), or by mustelids, mustelids-regulated phase (blue color sections in Figure~\ref{fig:populations}).
-Farmer-regulated periods started when densities of living voles triggered treatments.
-This produced sudden declines of live voles followed by increases of dead voles.
-However, the vole population re-grew quickly which triggered frequent further treatments and pulses of availability of contaminated (both live and dead) voles.
-In FFR d (intermediate threshold and low AR), vole declines were not as deep as when pulses of AR amount were high (FFR b), owing to the reduced rodenticide amount per treatment.
-Mustelids and foxes also experienced AR-induced declines during this period.
-Under FFRs b (high threshold density, high AR) and d (intermediate threshold, low AR), mustelid-regulated periods started when mustelid numbers grew slowly to a peak, which depressed vole density, precluding rodenticide treatments and releasing the fox population from secondary poisoning, such that its abundance rebounded.
-Vole depletion by mustelids and subsequent mustelid declines allowed the vole population to grow again up to threshold densities and initiated a new period of regulation by farmers.
+\subsection{Sensitivity analysis of model parameters}
+
+The exploration of ecotoxicological uncertainties on population dynamics has been done through the lethal dose parameter $LD_{50}$ for voles, mustelids and foxes (see Figure~\ref{fig:GSA}).
 %
-Vole dynamics were permanently regulated by AR treatment under FFR c (intermediate threshold, high AR, Figure~\ref{fig:populations}-c).
-Populations of live and dead voles experienced high frequency fluctuations (around 2 peaks every year) driven by AR.
-As AR treatments were frequent, being triggered by voles peaks, contaminated dead voles were always abundant (peaks at 90 ind ha$^{-1}$), and mustelid and fox densities remained low (see Supporting Information: mustelids <0.150 ind ha$^{-1}$ and foxes <0.01 ind ha$^{-1}$) but with a slowly increasing trend on the long term period (Figure~\ref{fig:populations}-c).
-This pattern of an increasing trend appears to be very sensitive to parameterization (i.e.
-a slight change in parameter may lead to a long-term decreasing pattern).
+From the 1600 patterns explored in this wide-range sensitivity analysis, we observed two typical patterns.
+On one hand (i) vole dynamics were sequentially regulated by either AR treatments, which we refer to as farmer-regulated phase (white sections in Figure~\ref{fig:populations}), or by mustelids, mustelids-regulated phase (grey sections in Figure~\ref{fig:populations}).
+This pattern of succession of farmer-regulated and mustelids-regulated phases are denoted "Mustelid periodic regulation" in Figure~\ref{fig:GSA} (black pixels).
+On the other hand, (ii) vole dynamics were permanently regulated by AR treatments, because of crash of the mustelid population, what is referred as "Permanent AR regulation" in Figure~\ref{fig:GSA} (grey pixels).
 %
-With FFR e (low threshold, low AR), vole populations were maintained by farmers at around 50 voles ha$^{-1}$ (See Figure~\ref{fig:populations} and Supporting Information).
-The population of live voles was buffered by treatments.
-Additionally, whenever voles reached densities triggering treatment, predator populations experienced strong declines.
-However, fox densities were higher compared to FFR c (intermediate threshold, high AR) (>0.01 ind ha$^{-1}$, see Supporting Information), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities.
-The maximum numbers of dead voles under this FFR e was relatively low (highest around 15 ind ha$^{-1}$) but, due to frequent treatments, there was a steady replenishment of contaminated dead voles.
-This, in turn, induced mustelid and fox mortality and population declines (Supporting Information).
-Additionally, low availability of live voles triggered small mustelids mortality through starvation down to abundances similar to those resulting from AR use (Supporting Information).
+We see that  $LD_{50}$ for foxes does not change the general pattern of dynamics.
+Both scenario B and D exhibited the highest part of "Mustelid periodic regulation" compared to other scenarios C and E.
+This illustrated the fine compromise between resource and AR impact to benefit from predator regulation periods.
 
-\subsection{Comparison between farmer functional responses in terms of cost functions}
-
-As expected, spreading AR generally reduced vole densities (Table~\ref{tab:results}) but the extent to which it also affected predators and the costs and benefits for conservation and farmer’s interests varied widely.
-The FFR b (high threshold, high AR) and d (intermediate threshold, low AR) implied the lowest number of treatments (0.5 and 1.8 year$^{-1}$ respectively) and the lowest amount of AR used (mean 10 and 13.5 kg ha$^{-1}$ year$^{-1}$ resp.).
-The FFR e (low threshold, low AR) had the lowest mustelid mortality due to AR (no instance in 40 years) but maximised starvation-induced mortality of mustelids (Table~\ref{tab:results}).
-Additionally, FFR e (low threshold, low AR) had the highest proportion of time with < 50 voles' ind ha$^{-1}$, (91\% of study period).
-When considering the effectiveness in reducing mean vole densities over the long-term period, FFR e (low threshold, low AR) ranked highest (only 55 ind ha$^{-1}$ in means, and >118 ind ha$^{-1}$ for other scenarios).
-The FFR b (high threshold, high AR) and d (intermediate threshold, low AR) allowed the highest mean densities of mustelids and foxes.
+For other parameters, we performed a local sensitivity analysis where we computed a first order sensitivity index \citep{Sobol1993, Saltelli2019} given the contribution of each parameter to the variance (in Figure~\ref{fig:sobol_sensitivity} for scenario D - intermediate amount of 7.5kg AR at threshold density of 250 voles).
 %
-Comparison of these protocols indicates that FFR b and d, inducing successive farmer-regulated and predator-regulated phases, were the best scenarios for conservation interests (Table~\ref{tab:results}).
-
-\subsection{Influence of intra-guild predation}
-
-Figure~\ref{fig:intraguild} shows the system dynamics under FFR d, where successive farmer-regulated and mustelids-regulated phases occured.
-This simulation shows that the removal of foxes did not eliminate the successions of mustelids-regulated and farmer-regulated phases.
-However, the mustelid regulated period allowed short-term peaks of voles, suggesting the emergence of a classical one-predator - one-prey cycles interrupted by a farmer-regulated period.
-Without foxes, population dynamics of mustelids presented a more chaotic behaviour, while it presented regular cyclic pattern with fox occurrence.
-Therefore, this simple model suggests a stabilizing role of a generalist predator (foxes in this system) during the mustelid-regulated period.
-At the end of the mustelid-regulated period, foxes strongly contributed to vole mortality and, to a lesser degree, mustelid mortality.
-Consequently, the removal of foxes implied less predation over voles during the mustelid regulated period, and short-term vole releases from mustelid predation.
-The 2-year rolling mean of vole density (blue lines in Figure~\ref{fig:intraguild}) illustrates the change of regime from farmer-regulated to mustelids-regulated phases.
-Indeed, the amplitude of averaged vole densities (i.e., the amplitude of vole cycles for the 2-year rolling mean) was relatively stable at the beginning of farmer regulation period and then suddenly decreased to become minimal before sharply increasing, announcing the regime change.
-These changes in density amplitude may be used as an early-warning signal of the regime transition.
-
-\subsection{Sensitivity of model under this parameterization}
-
+Parameters were moved within +/- 10\% interval around the fixed value as defined in scenario D (see Table~\ref{tab:parameterization}).
 %
-Figure~\ref{fig:sobol_sensitivity} represents the global sensitivity where we computed a first order sensitivity index \citep{Sobol1993, Saltelli2019} given the contribution of each parameter to the variance of the model d (i.e., 7.5kg of AR at threshold density of 250 voles).
+Crossing all cost functions to test the sensitivity of model parameters (Figure~\ref{fig:sobol_sensitivity}), the influence of each parameter appeared quite homogeneous.
 %
-Parameters are moved within +/- 10\% interval around the fixed value as defined in scenario d (see Table~\ref{tab:parameterization}).
+There was no redundant or conversely any parameters with a dominant influence on model behaviour.
 %
-Crossing all cost functions to test the sensitivity of model parameters, the influence of each parameter appears quite homogeneous.
-%
-There is no redundant or conversely any parameters with a dominant influence on model behaviour.
-%
-Taking cost functions separately, the sensitivity to parameters present some differences.
-%
-This is particularly true for densities of predators on which we focus our attention.  
-For mustelids and foxes variable.
+The homogeneity of the sensitivity to parameters was particularly true for densities of predators to which we focused our attention.
 
@@
 
+\subsection{Population dynamics}
+
+While both "Mustelid periodic regulation" and "Permanent AR regulation" pattern  are possible in each of the 4 scenarios with AR (B, C, D and E) (see Figure~\ref{fig:GSA}), we selected for Figure~\ref{fig:populations} the most contrasting population dynamics for each of them in order to better describe their underlying mechanism.
+The contrasting border emerge for vole $LD_{50}$ around 1.5 to 2 ppm and in the range 2 to 2.25 ppm for mustelid $LD_{50}$.
+
+From Figure~\ref{fig:populations}, we see that scenario B (high vole density threshold triggering treatment and high AR amount per treatment) and scenario D (intermediate threshold and low AR amount) present the most of dynamics with sequentially farmer-regulated phase and mustelids-regulated phase.
+%
+Farmer-regulated periods started when densities of living voles triggered treatments.
+This produced sudden declines of live voles followed by increases of dead voles.
+However, the vole population re-grew quickly which triggered frequent further treatments and pulses of availability of contaminated (both live and dead) voles.
+%
+Also, in scenario D, vole declines were not as deep as when pulses of AR amount were high like in scenario B (see curves in Supporting Information), owing to the reduced AR amount per treatment. 
+%
+Mustelids and sometimes foxes also experienced AR-induced declines during this period (Figure~\ref{fig:populations}-c,f,g,h and Figure~\ref{fig:intraguild}).
+For the predators, mustelid-regulated periods started when mustelid numbers grew slowly to a peak, which depressed vole density, precluding AR treatments and releasing the fox population from secondary poisoning, such that its abundance rebounded.
+Vole depletion by mustelids and subsequent mustelid declines allowed the vole population to grow again up to threshold densities and initiated a new period of regulation by farmers.
+%
+Sometimes, mustelids population has a slowly increasing trend (Figure~\ref{fig:populations}-g)). Under this regime, this increasing trend appears to be very sensitive to parameterization.
+
+For scenario C (intermediate threshold and high AR) vole dynamics were mostly regulated by AR treatment only (e.g.
+Figure~\ref{fig:populations}-e).
+Populations of live and dead voles experienced high frequency fluctuations driven by AR.
+As AR treatments were frequent, being triggered by vole peaks, contaminated dead voles were always abundant (see Supporting Information: peaks at 90 ind ha$^{-1}$).
+%
+With scenario E (low threshold, low AR) vole populations were maintained by farmers at around 50 voles ha$^{-1}$ (See Figure~\ref{fig:populations} and Supporting Information).
+%
+The population of live voles was buffered by treatments. 
+%
+For most of simulations (except when vole $LD_{50}$ with high, Figure~\ref{fig:GSA}), whenever voles reached densities triggering treatment, predator populations experienced strong declines.
+%
+However, fox densities were higher compared to scenario C (intermediate threshold, high AR) (see Supporting Information), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities.
+%
+The maximum numbers of dead voles under this scenario E was relatively low (highest around 15 ind ha$^{-1}$) but, due to frequent treatments, there was a steady replenishment of contaminated dead voles.
+This, in turn, induced mustelid and fox mortality and population declines (Supporting Information).
+%
+Additionally, low availability of live voles triggered small mustelids mortality through starvation, down to abundances similar to those resulting from AR use (Supporting Information).
+
+%\subsection{Comparison between farmer functional responses in terms of cost functions}
+
+%As expected, spreading AR generally reduced vole densities (Table~\ref{tab:results}) but the extent to which it also affected predators and the costs and benefits for conservation and farmer’s interests varied widely.
+%The FFR b (high threshold, high AR) and d (intermediate threshold, low AR) implied the lowest number of treatments (0.5 and 1.8 year$^{-1}$ respectively) and the lowest amount of AR used (mean 10 and 13.5 kg ha$^{-1}$ year$^{-1}$ resp.).
+The FFR e also had the lowest mustelid mortality due to AR (no instance in 40 years) but maximised starvation-induced mortality of mustelids (Table~\ref{tab:results}).
+FFR e (low threshold, low AR) had the highest proportion of time with < 50 voles' ind ha$^{-1}$, (91\% of study period).
+When considering the effectiveness in reducing mean vole densities over the long-term period, FFR e (low threshold, low AR) ranked highest (only mean 55 ind ha$^{-1}$ vs >118 ind ha$^{-1}$ for other scenarios).
+The FFR b (high threshold, high AR) and d (intermediate threshold, low AR) allowed the highest mean densities of mustelids and foxes.
+%
+%Comparison of these protocols indicates that FFR b and d, inducing successive farmer-regulated and predator-regulated phases, were the best scenarios for conservation interests (Table~\ref{tab:results}).
+
+\subsection{Influence of intra-guild predation}
+
+Figure~\ref{fig:intraguild} shows the system dynamics under a scenario D, where successive farmer-regulated and mustelids-regulated phases occurred.
+%
+This simulation shows that the removal of foxes did not eliminate the successions of mustelids-regulated and farmer-regulated phases.
+However, the mustelid regulated period allowed short-term peaks of voles, suggesting the emergence of a classical one-predator - one-prey cycles interrupted by a farmer-regulated period.
+Without foxes, population dynamics of mustelids presented a more chaotic behaviour, while it presented regular cyclic pattern with fox occurrence.
+%
+Therefore, this simple model suggests a stabilizing role of a generalist predator (foxes in this system) during the mustelid-regulated period.
+At the end of the mustelid-regulated period, foxes strongly contributed to vole mortality and, to a lesser degree, mustelid mortality.
+Consequently, the removal of foxes implied less predation over voles during the mustelid regulated period, and short-term vole releases from mustelid predation.
+The 2-year rolling mean of vole density (blue lines in Figure~\ref{fig:intraguild}) illustrates the change of regime from farmer-regulated to mustelids-regulated alternating phases.
+%
+Indeed, the amplitude of averaged vole densities (i.e., the amplitude of vole cycles for the 2-year rolling mean) was relatively stable at the beginning of farmer regulation periods and then suddenly decreased to become minimal before sharply increasing, announcing the regime change.
+These changes in density amplitude may be used as an early-warning signal of the regime transition.
+
 \section{Discussion}
 
-Considering that rodenticide kills not only voles but also their predators through secondary poisoning, our models show that AR profoundly changes the outcome of predator-prey dynamics involving vole, mustelid and fox populations beyond what mere intuition could elucidate.
-Our study reveals how the dual influences of the amount of pesticide spread and the vole density threshold triggering AR spread drives (i) pesticide spreading frequency, (ii) predation ecosystem service, and subsequently (iii) the control of pest outbreaks.
+Considering that rodenticide kills not only voles but also their predators through secondary poisoning, our models show that AR profoundly changes the modeled outcome of predator-prey dynamics involving vole, mustelid and fox populations beyond what mere intuition could elucidate.
+Our study reveals how the dual influences of the amount of pesticide spread and the vole density threshold triggering AR spread drive (i) pesticide spreading frequency, (ii) predation ecosystem service, and subsequently (iii) the control of pest outbreaks.
 Two types of a rich spectrum of emergent dynamics, including farmer or mustelid regulation deviating from classical predator prey dynamics arose because poisoned voles acted as “super-predators”.
@@
 In the ecosystem our models depict, farmers spreading rodenticide not only depleted vole prey exploited by specialist and generalist predators but also created pulses of lethally or sub-lethally poisoned voles that became super-predators by poisoning their predators.
+%
 Arguably this set of ecological interactions has similarities with circumstances where a pathogen affecting prey species also infects predators, as in the case with the flea vectored plague (\textit{Yersinia pestis}) infecting prairie dogs (\textit{Cynomys} spp.) and black footed ferrets (\textit{Mustela nigripes}) in central US \citep{Matchett2010}.
@@
 This is despite the obvious relevance to the management of the globally widespread circumstances where keystone small mammals are poisoned and may secondarily poison their predators \citep{Delibes-Mateos2011}.
-Under the “reference” scenario without AR spreading (FFR a), we assumed a predator-prey cycle which is a plausible pattern thoroughly explored theoretically \citep{Hanski2001}.
-There is no controversy on the role of small mustelids tracking vole dynamics, though it is not yet well understood whether predation may drive steep declines \citep{King2006}.
+
+Under the "reference" scenario without AR spreading (A), we assumed a predator-prey cycle which is a plausible pattern thoroughly explored theoretically \citep{Hanski2001}, though with debated empirical support \cite{Lambin2018}.
+%
+There is no controversy on the role of small mustelids tracking vole dynamics, though it is not yet well understood whether there is sufficient lag between predators and prey for predation to drive steep declines \citep{King2006}.
 Parameters of the reference scenario for our predator-prey model were realistic and tuned to generate population fluctuations similar to those observed in the studied cyclic system \citep{Delattre2009}.
@@
 
-An interesting model behaviour was seen with FFR b (high vole density threshold and high AR) and FFR d (intermediate vole density threshold and low AR) with farmer- and mustelid-regulated phases alternating with low frequency.
+An interesting model behaviour  with farmer- and mustelid-regulated phases alternating with low frequency was seen with scenarios B (high vole threshold, high AR) and D (intermediate vole threshold, intermediate AR), and to a lesser extend scenario C (intermediate vole threshold, high AR) and really few with scenario E where vole density threshold in the lowest.
+%
 Such flipping between alternative states in population dynamics has been previously described in predator-prey model where weasels rely on a primary prey and entrain the dynamics of secondary prey \citep{Hanski1996} but not for the kind of indirect interaction we explore here.
 It further demonstrates that adding biological realistic complexity to simple models may drastically change the emergent properties of trophic interactions.
+%
 From these scenarios, we understand that the emergence of successive farmer and mustelid phases is neither driven by vole density threshold alone nor by AR amount, but instead by a subtle combination of both.
-The modelling description of these patterns uncovered the dual key roles of mustelids on fox dynamics, as intraguild competitors and as a vector for poisoning.
+%
+Also, under sensitive fox scenario, the modelling description of these patterns uncovered the dual key roles of mustelids on fox dynamics, as intraguild competitors and as a vector for poisoning.
 This led to a surprising form of facilitation for foxes: mustelids protect foxes from collapses.
@@
 Secondly, fox predation on mustelids is reduced, and with an intermediate AR amount, this allows mustelids to slowly recover.
-Vole outbreaks and subsequently farmer treatments are then gradually delayed, benefitting mustelids recovery.
+Vole outbreaks and subsequently farmer AR treatments are then gradually delayed, benefitting mustelids recovery.
 This is the point of transition from farmer to mustelids regulation, starting the third step: mustelids increase faster, suppressing vole densities and precluding the need for AR treatments, and eventually indirectly allowing fox population growth.
-Our finding that complexities in trophic interaction, induced by the poisoning of predator by poisoned prey, may cause the system to flip between alternative states is novel and robust.
-However, given we only explored deterministic versions of our models, any inference on the frequency of flipping between states should be cautious given the inherent stochastic nature of natural and farmland environments.
+
+Our findings that complexities in trophic interaction, induced by the poisoning of predator by poisoned prey, may cause the system to flip between alternative states is novel and robust.
+However, given we only explored deterministic and spatially homogeneous versions of our model, any inference on the frequency of flipping between states should be cautious given the inherent stochastic nature of natural and farmland environments.
+%
 If such dynamics occur within real farming systems, flipping between states is unlikely to emerge with regularity where many other factors impact population dynamics.
@@
 
-Under FFR c (intermediate vole density threshold and high AR) and FFR e (low vole density threshold and low AR), the whole system was solely driven by farmer regulation, whereby the chronic use of AR completely suppressed the pest-regulation ecosystem service of predators.
+In many situations, notably scenario E and C and when vole and mustelids are highly sensitive to AR, the whole system was solely driven by farmer regulation, whereby the chronic use of AR completely suppressed the pest-regulation ecosystem service of predators.
+%
 It has previously been shown empirically that repeated rodenticide treatments are highly detrimental to the populations of predators and reduce their densities  
-\citep{Jacquot2013}.
-Secondary poisoning of predators is an established reality \citep{Berny2007}.
+\citep{Jacquot2013}. Secondary poisoning of predators is an established reality \citep{Berny2007}.
+%
 Through modelling, we formalised the insight that some poison deployment protocols, including those presently used in the empirical system which motivated our study, are counterproductive if employed on a large scale, suppressing natural predator regulation of pest rodents.
-It has been long known that poisoning rodents with AR permeates the food chain at peak abundance and achieves little in terms of protecting crops and may have strong deleterious impact (FFR b and c) \citep{Olea2009}.
+It has been long known that poisoning rodents with AR permeates the food chain at peak abundance, achieves little in terms of protecting crops and may have strong deleterious impact \citep{Olea2009}.
 In Franche-Comté, a change in treatment protocols, from controlling voles at high densities to low-intermediate densities, has reduced the mortality of non-target species (including foxes) \citep{Jacquot2013}.
-Nevertheless, deployment regimes of pesticides that can contaminate the food chain should also enable periods of time which allow predator populations to rebound and avoid extirpation from the ecosystem.
-We have shown that, over time, farmers who strictly maintain voles at low density thresholds would suppress predation services provided by vole predators and instigate pesticide dependence.
+%
+Nevertheless, deployment regimes of pesticides that can contaminate the food chain should also include periods of time which permit predator populations to rebound and avoid extirpation from the ecosystem.
+%
+We have shown that, over time, farmers who strictly maintain voles at low density thresholds would likely suppress predation services provided by vole predators and instigate pesticide dependence.
 In addition, small mammals like voles certainly have ecosystem functions.
@@
 
-Presently, in Franche-Comté, farmers relying on bromadiolone alone can only treat pre-emptively when voles are at low densities (FFR e) whereas those using alternative methods without pesticide are allowed to spread AR in low quantity up to intermediate vole densities (FFR d).
+Presently, in Franche-Comté, farmers relying on bromadiolone alone can only treat pre-emptively when voles are at low densities (scenario E) whereas those also using alternative methods (i.e., mechanical, not pesticide-based) are allowed to spread AR in low quantity up to intermediate vole densities (scenario D).
+%
 Spreading AR in low quantity seems superficially desirable, but our heuristic model, assuming an idealised homogenous landscape, show this is associated with frequent treatments.
 Consequently, it would induce a near permanent availability of a small number of intoxicated voles which, combined with low availability of non-contaminated voles, would reduce predator populations.
-Therefore, the extreme situation of using a low vole density threshold (FFR e) at a large scale is undesirable because it depletes the prey resources of foxes and mustelids and their populations.
-Triggering treatment at intermediate vole density with a low amount of AR (FFR d) allowed for temporal refuges, i.e.
+%
+Therefore, the extreme situation of using a low vole density threshold (scenario E) at a large scale is undesirable because it depletes the prey resources of foxes and mustelids and their populations.
+%
+Triggering treatment at intermediate vole density with a low amount of AR (scenario D) allowed for temporal refuges, i.e.
 longer periods free of rodenticide necessary for predator densities to rebound while simultaneously avoiding episodes with high vole density, as required by farm production interests.
-Under a landscape management approach, such temporal refuges could be spatial refuges, with parts of the landscape free of pesticides where predator populations can recover.
+Under a landscape management approach, such temporal refuges could arguably be substituted for by spatial refuges, with parts of the landscape free of pesticides where predator populations can recover.
 
 Our key result and the basis for management prescriptions is that allowing for refuges where voles are not poisoned and allowed to persist at medium-high densities such that they can be exploited by mustelids is crucial for predator population recovery and preserving the ecosystem services mustelids deliver.
+%
 Treatment regimes allowing refuges seems compatible with both conservation and farming interests.
@@
 Our process-based model revealed pesticides that permeate the food chain upward can lead to diverse population dynamics with alternative states regulated by predators and farmers.
-It also shows that the practice currently promoted to use low-dose AR treatments at low vole density could have limited by the undesirable side-effects of leading to chronic application of AR on a large scale, in the absence of refuges, and the depletion of the vole predator community.
+It also shows that the practice currently promoted to use low-dose AR treatments at low vole density could have the undesirable side-effects of leading to chronic application of AR on a large scale, in the absence of refuges, and the depletion of the vole predator community.
+%
 This emerging question would benefit from a landscape modelling approach to characterize spatial refuges.
@@
 XL conceived the initial idea; all authors developed the concept; VB and JF developed the models and led manuscript writing; VB implemented the model and ran simulations.
-GC contributed treatment protocols; JF, VB and MC explored model parameters; XL, PG and MC contributed critically to drafts; all authors gave final approval for publication.
+GC contributed treatment protocols; VB, MC, JF explored model parameters; XL, PG and MC contributed critically to drafts; all authors gave final approval for publication.
 
@@
 %
-$LD_{50,V}$ & Lethal Dose for 50\% of voles & ppm & $2$ \\
-$H_V$ & Hill's coefficient dose-response curve &  n.d. & $4$ \\
+$H_V$, $H_M$, $H_F$ & Hill's coefficient dose-response curve &  n.d. & $4$ \\
 %
-$LD_{50,M}$ & Lethal Dose for 50\% of mustelids & ppm & $2.1$ \\
-$H_M$ & Hill's coefficient dose-response curve & n.d. & $4$  \\
+$LD_{50,X}$ & Lethal Dose for 50\% of individuals of X & ppm & see Text \\
 %
-$LD_{50,F}$ & Lethal Dose for 50\% of foxes & ppm & $0.5$\\
-$H_F$ & Hill's coefficient dose-response curve & n.d. & $2$ \\
 %
@@
 
-\begin{table}
-	\caption{Results on the cost functions for the comparison of treatment protocols.
-Column \textit{"FFR"} (farmer functional response) refers to the protocol denotation (a to e) and in brackets respectively: the vole density threshold, and the amount of AR/treatment in kg.
-The column \textit{"Nbr treat/yr"} refers to the number of treatment in simulation per year after the burn-in phase.
-\textit{"Total AR/yr"} is the total use of anticoagulant rodenticides per year (bromadiolone in kg).
-\textit{"M, 50\% AR death"} is the proportion of time with mortality by AR over 50\% in mustelids population.
-\textit{"M, 50\% Nat.
-death"} is the proportion of time with mortality by natural reason (i.e., all except AR) over 50\% in mustelids population.
-\textit{"V < 50 ha$^{-1}$"} is the proportion of time with less than 50 voles ha$^{-1}$.
-Finally, \textit{"Mean V density"}, \textit{"Mean M density"} and \textit{"Mean F density"} are mean densities of voles, mustelids and foxes respectively.}
-	\label{tab:results}
-	\begin{tabular}{p{2cm} c c c c c c c c}
-		\hline
-		\tabcell{FFR \\ (dose, limit)} & \tabcell{Nbr.
-\\ treat/yr} & \tabcell{Total  \\ AR/yr} & \tabcell{M, 50\% \\ AR death} & \tabcell{M, 50\% \\ Nat.
-death} & \tabcell{V < \\ 50 ha$^{-1}$} & \tabcell{Mean V \\ density} & \tabcell{Mean M \\ density} & \tabcell{Mean F \\ density} \\ \hline
-		%%
-		a (0,0)                   & 0                          & 0                         & 0                             & 1                               & 0.46                          & 220                         & 0.62                        & 0.03                        \\
-		% 
-		b (500,20)                & 0.5                        & 10\texttt{}                       & 0.02                          & 0.98                            & 0.39                          & 198                         & 0.52                        & 0.02                        \\
-		%
-		c (250, 20)                & 2.8                        & 56                        & 0.09                          & 0.91                            & 0.15                          & 118                         & 0.08                        & <0.01                       \\
-		%
-		d (250, 7.5)               & 1.8                        & 13.5                        & 0.04                          & 0.96                            & 0.29                          & 126                         & 0.55                        & 0.02                        \\
-		%
-		e (50,7.5)                 & 5.1                        & 38                        & 0                             & 1                               & 0.91                          & 55                          & 0.07                        & 0.02                        \\ \hline
-	\end{tabular}
-\end{table}
-
 % GRAPHIC SCHEME
@@
 
+% GLOBAL SENSITIVITY ANALYSIS
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\linewidth]{img/plt_GSA.jpg}
+		\caption{Global sensitivity analysis of Lethal Dose for voles, mustelids and predators on the population dynamics of mustelids: "Mustelids periodic regulation" (see Figure \ref{fig:populations}-c,f,g,h) or "Permanent AR regulation" (see Figure \ref{fig:populations}-b,d,e,i,j).
+Points are simulations in Figure~\ref{fig:populations}.}
+		\label{fig:GSA}
+	\end{center}
+\end{figure}
+
 % GRAPHIC POPULATION DYANMICS
@@
 		\includegraphics[width=\linewidth]{img/plot_join.png}
-		\caption{Rate of variation of population densities simulated over 50 years (10 years of burn-in period in grey area): (Left) Dynamics of mustelids (grey line) and red foxes (black line), (Right) Dynamics of live (grey line) and dead (black line) voles.
-Letters at the top-center part of each sub-graph corresponds to the results obtained from each farmer functional response.
-Colored areas indicate the farmer-regulated (red area) and mustelid-regulated (blue area) periods of vole population dynamics respectively.}
+		\caption{Variation of population densities simulated over 50 years (10 years of burn-in period in grey area).
+Capital letter corresponds to farmer functional response: A the null model (no AR), B (vole threshold: 500in/ha, AR 20kg/ha), C(250,20), D(250, 7.5) and E(50,7.5).
+Numbers correspond to points in Figure~\ref{fig:GSA}.}
 		\label{fig:populations}
@@
 \caption{Effects of intraguild predation.
-Top graphics: daily densities of voles (in black) and 2-years rolling means of densities (in blue) for the farmer functional response d (i.e., 7.5kg of rodenticide at threshold density of 250 voles) with and without foxes (left and right respectively).
+Top panels: daily densities of voles (in black) and 2-years rolling means of densities (in blue) for the farmer functional response d (i.e., 7.5kg of rodenticide at threshold density of 250 voles) with and without foxes (left and right respectively).
 Bottom graphics: stacked charts of ingestion and mortality proportion.}
@@
 		\includegraphics[width=\linewidth]{img/plt_sobolIndex_draw.jpg}
-		\caption{First-order Sobol's sensitivity index (denoted $S_p$ for parameter $p$) providing the global sensitivity of cost functions to parameters of the model \citep{Sobol1993} for the model d (i.e., 7.5kg of AR at threshold density of 250 voles).
+		\caption{Local sensitivity analysis: First-order Sobol's sensitivity index (denoted $S_p$ for parameter $p$) providing the global sensitivity of cost functions to parameters of the model \citep{Sobol1993} for the model d (i.e., 7.5kg of AR at threshold density of 250 voles).
 Analysis is based on 4200 simulations.
-Values reflect the expected fractional reduction in the variance for each cost functions that would be achieved if a specific parameter is fixed.
-If $S_p = 1$, the sensitivity is not influenced by the parameter $p$.}
+Values reflect the expected fractional reduction in the variance for each cost functions that would be achieved if a specific parameter is fixed.}
 		\label{fig:sobol_sensitivity}

### 2019-10-24 trophToxNTO/R/ode_solver.R

@@ -105,12 +105,12 @@ ode.broma.tritrophic_2nd = function(Time, State, Pars) { # The model is a functi
 
     ### vole dynamics
     dV = r.V*D.V*(1-D.V/K.V) -
-      dose.resp.mort(C.V, LD.50.V,H.V,days.to.die.V)*D.V - # broma mortality # 6 days to die
+      dose.resp.mort(Cing.V, LD.50.V,H.V,days.to.die.V)*D.V - # broma mortality # 6 days to die
       a.M*D.V/(a.M*D.V+a.Md*D.Vd)*(a.M*D.V+a.Md*D.Vd)/(1+h.M*(a.M*D.V+a.Md*D.Vd))*D.M - #mustelid functional response
       a.F*D.V/(a.F*D.V+a.Fd*D.Vd+a.FM*D.M)*(a.F*D.V+a.Fd*D.Vd+a.FM*D.M)^2/(1+h.F*(a.F*D.V+a.Fd*D.Vd+a.FM*D.M)^2)*D.F #fox functional response
 
     ### dead Vole dynamics:
-    dVd = dose.resp.mort(C.V, LD.50.V,H.V,days.to.die.V)*D.V - d*D.Vd -# add vole dead by bromadiolone
+    dVd = dose.resp.mort(Cing.V, LD.50.V,H.V,days.to.die.V)*D.V - d*D.Vd -# add vole dead by bromadiolone
       a.M*D.Vd/(a.M*D.V+a.Md*D.Vd)*(a.M*D.V+a.Md*D.Vd)/(1+h.M*(a.M*D.V+a.Md*D.Vd))*D.M - #mustelid functional response
       a.F*D.Vd/(a.F*D.V+a.Fd*D.Vd+a.FM*D.M)*(a.F*D.V+a.Fd*D.Vd+a.FM*D.M)^2/(1+h.F*(a.F*D.V+a.Fd*D.Vd+a.FM*D.M)^2)*D.F
 
